---
title: "R Notebook"
output: rmarkdown::github_document
---

# Further data preparation in R 
testing

## Introduction

In this notebook, we will prepare data for analysis.

Specifically, we will:

-   Format the data
-   Add a disease outcome using a combination of data sources
-   Perform exclusions for accelerometer data quality, prior disease, and missing data
-   Recode data where appropriate

## About this notebook

This notebook is converted from a notebook (2_Further_Pref_in_R.ipynb) that runs the R code from a jupyter notebook.

Note that this notebook uses the following conventions:

-   base R syntax (you might be more familiar with either 'tidyverse' or 'data.table' syntax)
-   data frames containing UKB data will be called 'dat_X'
-   output tables will be called 'tab_X'
-   date variables will be called 'date_X'
-   indicator variables will be called 'ind_X'

## How to run this notebook

This notebook should be run in a *RStudio* session. It does *not* require a Spark cluster. See how to set it up [here](https://dnanexus.gitbook.io/uk-biobank-rap/working-on-the-research-analysis-platform/using-rstudio-on-the-research-analysis-platform).

## Set up the session

We load packages we'll use:

```{r, include=FALSE}
# First we need to install packages that aren't already present
pkgs <- c("data.table", "plyr", "dplyr") # packages we need
pkgs_inst <- pkgs[!{pkgs %in% rownames(installed.packages())}] # check which are not present 
install.packages(pkgs_inst, repos = "https://www.stats.bris.ac.uk/R/") # install

# Load packages
lapply(pkgs, library, character.only = TRUE) # using lapply just allows us to load several packages on one line - this could be replaced with several calls to library()

source("rounding_functions.R")
source("cut_by_quantile.R")
source("table_lookup.R")
```

dx download - downloads stored permenant storage RAP data into my temporary session
Needs bash to run 
```{bash}
cd ~/rap_wearables

dx download /users/mcgaghd/Data -r
dx download /shared_data/data_clean/stepcount2.1.5_20230615.csv -r
dx download /shared_data/data_clean/stepcount3.1.0+5.gc84834d_82k_20230926.csv -r
```


We load data:

```{r}
# Note that in order to be able to load the data from these locations, you will need to start a new Jupyter Lab session relative to the previous notebook.
# Alternatively you could replace these locations with appropriate file locations from the current session's storage, or you could try remounting (see: 
# community.dnanexus.com/s/question/0D5t000003vAWYxCAO/it-seems-that-the-recently-dx-uploaded-files-does-not-show-up-on-mntproject-until-i-restart-the-whole-jupyter-lab-vm)
dat <- fread("Data/participant_wacc_data.csv", data.table = FALSE) # fread is a function from the data.table package for fast reading of large data
dat_hes <- fread("Data/hes_wacc_data.csv", data.table = FALSE)
dat_death <- fread("Data/death_wacc_data.csv", data.table = FALSE)
dat_death_cause <- fread("Data/death_cause_wacc_data.csv", data.table = FALSE)
dat_charlson <- fread("Data/charlson_data_10yr.csv", data.table = FALSE)
```



## Basic R formatting

Rename the derived accelerometer variables from Walmsley 
```{r}
colnames(dat)[colnames(dat) == "Light - Overall average | Instance 0"] <- "light_overall_average"
colnames(dat)[colnames(dat) == "Sedentary - Overall average | Instance 0"] <- "sedentary_overall_average"
colnames(dat)[colnames(dat) == "Moderate-Vigorous - Overall average | Instance 0"] <- "MVPA_overall_average"
colnames(dat)[colnames(dat) == "Sleep - Overall average | Instance 0"] <- "sleep_overall_average"
```


Rename the columns in Harper's Charlson groupings so as not to clash with derived continuous CCI groups from Walmsey + Small
```{r}
colnames(dat_charlson)[colnames(dat_charlson) == "date_end_accel_raw"] <- "date_end_accel_raw_CH"
colnames(dat_charlson)[colnames(dat_charlson) == "major_coronary_disease_charlson"] <- "major_coronary_disease_charlson_CH"
colnames(dat_charlson)[colnames(dat_charlson) == "peripheral_vascular_disease_charlson"] <- "peripheral_vascular_disease_charlson_CH"
colnames(dat_charlson)[colnames(dat_charlson) == "cerebrovascular disease_charlson"] <- "cerebrovascular_disease_charlson_CH"
colnames(dat_charlson)[colnames(dat_charlson) == "congestive_heart_failure_charlson"] <- "congestive_heart_failure_charlson_CH"
colnames(dat_charlson)[colnames(dat_charlson) == "cancer_charlson"] <- "cancer_charlson_CH"
colnames(dat_charlson)[colnames(dat_charlson) == "connective_tissue_disorder_charlson"] <- "connective_tissue_disorder_charlson_CH"
colnames(dat_charlson)[colnames(dat_charlson) == "copd_charlson"] <- "copd_charlson_CH"
colnames(dat_charlson)[colnames(dat_charlson) == "dementia_charlson"] <- "dementia_charlson_CH"
colnames(dat_charlson)[colnames(dat_charlson) == "diabetes_charlson"] <- "diabetes_charlson_CH"
colnames(dat_charlson)[colnames(dat_charlson) == "paraplegia_charlson"] <- "paraplegia_charlson_CH"
colnames(dat_charlson)[colnames(dat_charlson) == "hiv_charlson"] <- "hiv_charlson_CH"
colnames(dat_charlson)[colnames(dat_charlson) == "liver_disease_charlson"] <- "liver_disease_charlson_CH"
colnames(dat_charlson)[colnames(dat_charlson) == "peptic_ulcer_charlson"] <- "peptic_ulcer_charlson_CH"
colnames(dat_charlson)[colnames(dat_charlson) == "renal_disease_charlson"] <- "renal_disease_charlson_CH"
```

Combine CH charlson groups to dat 
```{r}
dat <-merge(dat, dat_charlson, by = "eid", all.x = TRUE)
```

We start by keeping a subset of columns which we are likely to use:

```{r}
cols_dat <- c("eid", "sex", "year_birth", "month_birth", "ethnicity_raw",
              "ukb_assess_cent",  "date_baseline", "date_inst_1", "date_inst_2", "date_lost_followup",
              "tdi_raw", "age_education_raw", "qualif_raw", "alcohol_raw", "smoking_raw", "BMI_raw",
              "self_report_cvd_baseline", "self_report_cvd_inst_1", "self_report_cvd_inst_2", 
              "date_end_accel", "quality_good_wear_time", "Wear duration overall", 
              "quality_good_calibration", "clips_before_cal", "clips_after_cal", "total_reads","conditions_at_baseline_2006","sedentary_overall_average", "light_overall_average","MVPA_overall_average","sleep_overall_average",
              "overall_activity", colnames(dat)[grepl("acceleration", colnames(dat))])
cols_dat_hes <- c("eid","dnx_hesin_id", "dnx_hesin_diag_id", 
                  "dateepiimp",  "ins_index", "arr_index", "level",
                  "diag_icd9", "diag_icd9_nb", "diag_icd10", "diag_icd10_nb")

dat <- dat[, cols_dat]
dat_hes <- dat_hes[, cols_dat_hes]
```


We inspect the data structure to check all columns are the types we expect:

```{r}
for (data in list(dat, dat_hes, dat_death, dat_death_cause)){
    str(data, vec.len = 0) # vec.len = 0 avoids accidentally printing data
}
```

Mostly this looks sensible, but there are some things to address. For example, why is "age_education_raw" formatted as character? It turns out it's because there are some special values for [age completed full time education](https://biobank.ndph.ox.ac.uk/showcase/field.cgi?id=845), which need to be removed before it can be coerced to numeric. Let's reformat it appropriately.

```{r}
dat$age_education_revalued_raw <- plyr::revalue(dat$age_education_raw, 
                                                c("Do not know" =  NA, 
                                                  "Prefer not to answer" = NA, 
                                                  "Never went to school" = 0))
dat$age_education_numeric_raw <- as.numeric(dat$age_education_revalued_raw)
```

We also do some simple formatting of date columns:

```{r}
# Tabular participant data
dat$date_lost_followup <- as.Date(dat$date_lost_followup, format = "%Y-%m-%d")
dat$date_end_accel <- as.Date(dat$date_end_accel, format = "%Y-%m-%d")
for (suffix in c("baseline", "inst_1", "inst_2")){
 dat[, paste0("date_", suffix)] <- as.Date(dat[, paste0("date_", suffix)], 
                                           format = "%Y-%m-%d") 
}

# Hospital data
dat_hes$date_hes <- as.Date(dat_hes$dateepiimp, format = "%Y-%m-%d")

# Death data
dat_death$date_death <-
  as.Date(dat_death$date_of_death, format = "%Y-%m-%d")
# A very small number of participants have duplicate records in death data (e.g. perhaps from a second death certificate after post-mortem)
# In this dataset we keep just one record per participant: they should have the same date, and we will use the death_cause dataset for any 
# other records related to death. It also only affects a very small number of participants.
dat_death <-
  dat_death[dat_death$ins_index == 0, ]
```

We'll do more involved processing later on, but this just ensures we have a sensibly coded dataset to work with.

## Find the first occurrence in hospital record data

First study is investigating cross-sectional step count between HC, psoriasis and psoriatic arthritis. 

This will aim to create columnns of participants with prevalent disease at time of accelerometer wear (2013-2015)


We briefly introduced the hospital data in the last notebook. A bit more detail on the structure of the data:

-   Diagnoses in hospital data mostly use [ICD-10 codes](https://icd.who.int/browse10/2010/en). Some records from the 1990s and earlier use its precursor, ICD-9.
-   Recall that there can be multiple records per participant (which is why we will look to identify the earliest record).
-   The dataset we have extracted from UK Biobank is *inpatient* episodes only. That means that it only includes hospital visits where the participant occupied a bed (not necessarily overnight). This means Emergency Department attendances and outpatient appointments which did not result in admission will not be present in the data. This is more of an issue for some conditions than others.
-   While operations and procedures codes are also available in UK Biobank hospital inpatient data, we won't use them here.
-   There are some papers looking at the features of different data sources for different health outcomes e.g. [this paper on stroke](https://n.neurology.org/content/95/6/e697). [This paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3264740/) discusses contexts in which different aspects of accurate classification are important.
-   If you want to learn more about hospital data in UK Biobank, see [this resource](https://biobank.ndph.ox.ac.uk/showcase/ukb/docs/HospitalEpisodeStatistics.pdf).

Following [Ramakrishnan et al. (2021)](https://journals.plos.org/plosmedicine/article?id=10.1371/journal.pmed.1003487), here we will define a record for 'cardiovascular disease' as an appearance of ICD-10 codes I20-I25 (ischaemic heart disease) or I60-I69 (cerebrovascular disease). We will also consider ICD-9 codes 410-414 (ischaemic heart disease) and 430-438 (cerebrovascular disease):


```{r}
PsA_icd10_codes <- "^M07[^4|^5|^6].*$|^L405$|^M090.*$"
PsA_icd9_codes <- "^6960$"
PsO_icd10_codes <- "^L40[^5].*$" 
PsO_icd9_codes <- "^6961$"

# 1 Self reported
dat$self_reported_PsO <- grepl("psoriasis", dat$conditions_at_baseline_2006)
dat$self_reported_PsA <- grepl("psoriatic arthropathy", dat$conditions_at_baseline_2006)

# 2 Death record
dat <- merge(
  dat,
  death_disease_lookup(dat_death_cause, dat, paste(PsA_icd10_codes),
                       "Death_PsA"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, PsA_icd10_codes, PsA_icd9_codes,
                     "prevalent", "hes_prevalent_PsA"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, PsA_icd10_codes, PsA_icd9_codes,
                     "incident", "hes_incident_PsA"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

dat$PsA_incident <- dat$hes_incident_PsA & (!dat$hes_prevalent_PsA |
                                            !dat$Death_PsA | 
                                            !dat$self_reported_PsA)

dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, PsA_icd10_codes, PsA_icd9_codes,
                     "date", "PsA_incident_date"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

dat$PsA_incident_date[!dat$PsA_incident] <- NA
dat$PsA_incident_date <- as.Date(dat$PsA_incident_date, format = "%Y-%m-%d")

# Then we look-up those with psoriasis 
dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, PsO_icd10_codes, PsO_icd9_codes,
                     "prevalent", "hes_prevalent_PsO"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

##create new total psoriasis 
dat$total_prevalent_PsO <- dat$self_reported_PsO | dat$hes_prevalent_PsO

dat$total_prevalent_PsA <- dat$self_reported_PsA | dat$hes_prevalent_PsA

##Create new variable which is psoriasis ONLY
dat$Prevalent_PsO_lessPsA <- dat$total_prevalent_PsO & !dat$total_prevalent_PsA
```

Remove duplicate columns 
```{r}
dat <- dat[, !duplicated(names(dat))]
```


Depression 
An individual is labelled as prevalent depression if: 1. They have been to the doctor (GP) for nerves, anxiety, tension or depression 2. They have been to a psychiatrist for nerves, anxiety, tension or depression 3. They have a hospital episode labelled for depression or 4. They self-reported at baseline as having depression

```{r}
# Look up ICD9 and 10 codes for depression in the HES data
dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, "^F3(2|3).*$", "^311.*$",
                     "prevalent", "hes_depression"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

# Look up depression in the self reported illnesses
dat$self_reported_depression <- grepl('depression', dat$conditions_at_baseline_2006)

# Combine all into one column
dat$depression <- (   dat$hes_depression | 
                     dat$self_reported_depression) 

table(dat$depression)                   
```


```{r}
# The lists of ICD codes we will consider
RA_icd10_codes <- "^M05.*$|^M06.*$|^M08.0$|^M08.2$|^M08.3$|^M08.4$"
RA_icd9_codes <- "^714.*$"
AS_icd10_codes <- "^M45$|^M46.9$" 
AS_icd9_codes <- "^720$"

# 1 Self reported
dat$self_reported_RA <- grepl("rheumatoid arthritis", dat$conditions_at_baseline_2006)
dat$self_reported_AS <- grepl("ankylosing spondylitis", dat$conditions_at_baseline_2006)


# 3 Hospital episode
# We look to exclude all individuals with any records (before or after study)
# with inflammatory joint condition (that is Ank Spond, RA and Psoriatic arthritis)
#dat <- merge(
 # dat,
  #hes_disease_lookup(dat_hes, dat, exclusionPsO_icd10_codes, exclusionPsO_icd9_codes,
                    # "prevalent", "disease_exclude"),
  #by = "eid",
  #all.x = TRUE,
  #suffixes = c("", "dup")
#)

# Then we look-up those with rheumatoid arthritis
dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, RA_icd10_codes, RA_icd9_codes,
                     "prevalent", "hes_prevalent_RA"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

# Then we look-up those with Ankylosing Spondylitis 
dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, AS_icd10_codes, AS_icd9_codes,
                     "prevalent", "hes_prevalent_AS"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

##create new total RA 
dat$total_prevalent_RA <- dat$self_reported_RA | dat$hes_prevalent_RA

##create new total Ank Spond arthritis 
dat$total_prevalent_AS <- dat$self_reported_AS | dat$hes_prevalent_AS

duplicate_cols <- duplicated(names(dat))
dat <- dat[, !duplicate_cols]
```

## Variables

### Age

Age at accelerometer wear:

```{r}
# Add date of birth
dat$approx_dob <-
  as.Date(paste(dat$year_birth, dat$month_birth, "15", sep = "-"),
          "%Y-%B-%d") # UK Biobank doesn't contain day of birth as it would be unnecessary identifying information, so we roughly impute it as the 15th of the birth month.

# Add age at entry in days
dat$age_entry_days <-
  difftime(dat$date_end_accel,
           dat$approx_dob,
           units = "days")

# Convert to age at entry in years
dat$age_entry_years <- as.double(dat$age_entry_days)/365.25
```

### Sex

Male, female

Should need no additional prep

### Age

40-44 $$note this is really 43-44$$; 45-49; 75-79

```{r}
# Add age groups
dat$age_gp <-
  cut(
    dat$age_entry_years,
    breaks = c(40, 45, 50, 55, 60, 65, 70, 75, 80),
    right = FALSE
  )
```

### Ethnicity

White, non-white

```{r}
# Ethnicity
dat$ethnicity <-
  plyr::revalue(
    dat$ethnicity_raw,
    c(
      "British" = "White",
      "Any other white background" = "White",
      "Irish" = "White",
      "White and Asian" = "Nonwhite",
      "Caribbean" = "Nonwhite",
      "Chinese"   = "Nonwhite",
      "Pakistani"  = "Nonwhite",
      "White and Black African" = "Nonwhite",
      "Other ethnic group"  = "Nonwhite",
      "Any other mixed background" = "Nonwhite",
      "African"    = "Nonwhite",
      "White and Black Caribbean" = "Nonwhite",
      "Prefer not to answer" = NA,
      "Indian"  = "Nonwhite",
      "White" = "White",
      "Do not know" = NA,
      "Any other Black background" = "Nonwhite",
      "Any other Asian background"  = "Nonwhite",
      "Bangladeshi"  = "Nonwhite",
      "Mixed"  = "Nonwhite",
      "Asian or Asian British"  = "Nonwhite",
      "Black or Black British"  = "Nonwhite"
    )
  )
```

### BMI

```{r}
# BMI
dat$BMI <- dat$BMI_raw

dat$BMI_cats <-
  cut(dat$BMI,
      breaks = c(0, 18.5, 25, 30, 10000),
      labels = c("<18.5", "18.5-24.9", "25.0-29.9", "30.0+"),
      right = FALSE)
```

### Smoking status

Never, Former, Current

```{r}
# Smoking
dat$smoking <-
  plyr::revalue(dat$smoking_raw, replace = c("Prefer not to answer" = NA))
```

### Alcohol consumption

```{r}
# Alcohol
dat$alcohol <-
  plyr::revalue(
    dat$alcohol_raw,
    replace = c(
      "Prefer not to answer" = NA,
      "Three or four times a week" = "3+ times/week",
      "Special occasions only" = "<3 times/week",
      "One to three times a month" = "<3 times/week",
      "Daily or almost daily" = "3+ times/week",
      "Once or twice a week" = "<3 times/week"
    )
  )
```

### TDI

By quarter in population, based on UK census

```{r}
dat$tdi_cats <- cut(
    dat$tdi_raw,
    breaks = c(-6.3627, -2.4167, -0.4373, 1.7863, 4.7426, 13.5881),
    labels = c("Least Deprived", "2nd Quintile", "3rd Quintile", "4th Quintile", "Most Deprived"),
    right = FALSE
  )
```

```{r}
dat$overall_activity_quarters <- qtile_cut(dat$overall_activity, labels = c("Quarter 1", "Quarter 2", "Quarter 3", "Quarter 4"))
```

### Wear season

Spring, Summer, Autumn, Winter

```{r}
dat$month_wear <- month(dat$date_end_accel)
dat$season_wear <- plyr::mapvalues(dat$month_wear,
                                   c(12, 1:11),
                                   c(
                                     rep("Winter", 3),
                                     rep("Spring", 3),
                                     rep("Summer", 3),
                                     rep("Autumn", 3)
                                   ))
table(dat$month_wear, dat$season_wear) # showing Dec-Feb assigned to winter (based on end time of accelerometer wear) and so on
```

Pulling in steps data from shared data file 

```{r}
dat_steps_ox <- fread("stepcount2.1.5_20230615.csv", data.table = FALSE)
names(dat_steps_ox)[names(dat_steps_ox) == 'StepsDayMedAdjusted'] <-
  'med_steps'

dat <- merge(dat, dat_steps_ox, by = "eid", all.x = TRUE)
```

```{r}
step_cat_bounds <- c(0, 5000, 7500, 10000, 12500, 15000, Inf)
dat$step_cats <- cut(dat$med_steps, breaks = step_cat_bounds, right = FALSE)
```

Read in steps data with timing of steps in day + then create new table with this splitting off from main analysis "dat" as there are currently only ~80,000 people with timing of steps calculated. 
Need to generate an eid column 

```{r}
dat_steps_ox_time <-fread("stepcount3.1.0+5.gc84834d_82k_20230926.csv", data.table=FALSE)

dat_steps_ox_time <- dat_steps_ox_time %>%
  mutate(eid = sub('.*(group[0-9]+)/', '', Filename),
  eid = sub('_.*', '', eid))
```

We will do seperate removals for non-wear down below - as I have done with Scott's steps. Once final steps is updated then merge this work together. 
```{r}
datTime <-merge(dat, dat_steps_ox_time, by = "eid", all.x = TRUE) %>%
  filter(!is.na(StepsQ1DayAvgAdjustedAt))
```

```{r}
tab_exc <- data.frame("Exclusion" = "Starting cohort", "Number_excluded" = NA, "Number_remaining" = nrow(datTime))
```

```{r}
nb <- nrow(datTime)
datTime <- datTime[!is.na(datTime$med_steps), ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "No step data",
      "Number_excluded" = nb - nrow(datTime),
      "Number_remaining" = nrow(datTime)
    )
  )
```

-   Exclude participants whose device could not be calibrated:

```{r}
nb <- nrow(datTime)
datTime <- datTime[datTime$quality_good_calibration == "Yes", ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Poor calibration",
      "Number_excluded" = nb - nrow(datTime),
      "Number_remaining" = nrow(datTime)
    )
  )
```

-   Exclude participants for whom \>1% of values were clipped (fell outside the sensor's range) before or after calibration:

```{r}
nb <- nrow(datTime)
datTime <- datTime[(datTime$clips_before_cal < 0.01*datTime$total_reads) & (datTime$clips_after_cal < 0.01*datTime$total_reads) , ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Too many clips",
      "Number_excluded" = nb - nrow(datTime),
      "Number_remaining" = nrow(datTime)
    )
  )
```

-   Exclude participants who had \<3 days wear or did not have wear in each hour of the 24 hour day:

```{r}
nb <- nrow(datTime)
datTime <- datTime[datTime$quality_good_wear_time == "Yes", ] # Note that this has already been calculated in UKB, 
# we don't need to manually calculate it: https://biobank.ndph.ox.ac.uk/showcase/field.cgi?id=90015
# But we might actually use the values from the new data processing

```
Missing data in adjustment variables:

```{r}
for (
  cov in c(
    "age_entry_years",
    "sex",
    "BMI",
    "ethnicity",
    "tdi_raw",
    "smoking",
    "alcohol"
  )
){
  nb <- nrow(datTime)
  print(cov)
  missing_cov <- is.na(datTime[, cov])|(as.character(datTime[, cov]) == "") |(as.character(datTime[, cov]) == "Missing") # for safety coerce to character for second check as can return NA on some classes e.g. Date
  datTime <- datTime[!missing_cov,]
  tab_exc <- rbind(
    tab_exc,
    data.frame(
      "Exclusion" = paste0("Missing ", cov),
      "Number_excluded" = nb - nrow(datTime),
      "Number_remaining" = nrow(datTime)
    )
  )
}
```


Charlson 
```{r}
charlson_codelist <- list(
  "acute_mi"                   = list("I21|I22|I23|I252|I258", 5),
  "cerebral_vascular_accident" = list("G450|G451|G452|G454|G458|G459|G46|I6", 11),
  "heart_failure"              = list("I50", 13),
  "connective_tissue_disorder" = list("M05|M060|M063|M069|M32|M332|M34|M353", 4),
  "dementia"                   = list("F00|F01|F02|F03|F051", 14),
  "diabetes"                   = list("E101|E105|E106|E108|E109|E111|E115|E116|E118|E119|E131|E135|E136|E138|E139|E141|E145|E146|E148|E149", 3),
  
  "liver_disease"              = list("K702|K703|K717|K73|K74", 8),
  "peptic_ulcer"               = list("K25|K26|K27|K28", 9),
  "peripheral_vascular_disease"= list("I71|I739|I790|R02|Z958|Z959" , 6),
  "pulmonary_disease"          = list("J40|J41|J42|J43|J44|J45|J46|J47|J60|J61|J62|J63|J64|J65|J66|J67", 4),
  "cancer"                     = list("C0|C1|C2|C3|C4|C5|C6|C70|C71|C72|C73|C74|C75|C76|C81|C82|C83|C84|C85|C86|C87|C88|C89|C90|C91|C92|C93|C94|C95|C96|C97", 8),
  
  "diab_complications"         = list("E102|E103|E104|E107|E112|E113|E114|E117|E132|E133|E134|E137|E142|E143|E144|E147", -1),
  "paraplegia"                 = list("G041|G81|G820|G821|G822", 1),
  "renal_disease"              = list("I12|I13|N01|N03|N052|N053|N054|N055|N056|N072|N073|N074|N18|N19|N25", 10),
  "metastatic_cancer"          = list("C77|C78|C79|C80", 14),
  "severe_liver_disease"       = list("K721|K729|K766|K767", 18),
  "hiv"                        = list("B20|B21|B22|B23|B24|O987", 2)
)
```


```{r}
# SUBSET DATA TO THE 5Y PRIOR TO ACC WEAR ================================================================
dat_hes_w_acc <- merge(dat_hes, dat[, c("eid", "date_end_accel")], all.x = TRUE)
dat_hes_w_acc$time_rel_to_acc <- difftime(dat_hes_w_acc$date_hes, dat_hes_w_acc$date_end_accel, units = "days")
dat_hes_5y_lookback <- dat_hes_w_acc[(dat_hes_w_acc$time_rel_to_acc < 0) & (dat_hes_w_acc$time_rel_to_acc > -365.25*5), ]


for (disease in names(charlson_codelist)){
  # Prep
  details <- charlson_codelist[[disease]]
  string <- details[[1]]
  weight <- details[[2]]

  # Restrict to relevant ids
  dat_hes_5y_lookback_current_code_ids <- unique(dat_hes_5y_lookback$eid[grepl(string, dat_hes_5y_lookback$diag_icd10)])
  
  # Record
  dat[, paste0(disease, "_charlson")] <- ifelse(dat$eid %in% dat_hes_5y_lookback_current_code_ids, weight, 0)

  # Spit out progress info
  print(disease)
  print(string)
  print(table(dat[, paste0(disease, "_charlson") ]))
  
  # Tidy
  rm(string, weight, details)
}

# Total score
dat$cci <- apply(dat[, paste0(names(charlson_codelist), "_charlson")] , 1, sum)

# Truncate scores on [0-50]
dat$cci[dat$cci < 0] <- 0 
dat$cci[dat$cci > 50] <- 50
```
Create some CCI categorical variables Kim et al
Low CCI: 0-2, High CCI: greater than or equal to 2  https://www.nature.com/articles/s41598-021-98026-4 
```{r}
dat <- dat %>%
  mutate(CCI_kim = ifelse(cci == 0, "None", ifelse(cci >= 1 & cci <= 2, "Low", "High")))
```

Mild, Moderate, severe: Huang et al
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3891119/
0: none, 1-2: mild, moderate 3-4, severe >/ 5
```{r}
dat <- dat %>%
  mutate(cci_huang = case_when(
    cci == 0 ~ "None",
    cci >= 1 & cci <= 4 ~ "Mild to Moderate",
    cci >= 5 ~ "Severe",
    TRUE ~ NA_character_  # Handle other cases as NA
  ))
```





## Writing out the data for reuse

As previously, we need to write out the data so we can reuse it, and upload it to the RAP system.

```{r}
write.csv(dat, "prepped_steps_before_exclude.csv", row.names = FALSE)
```


## Now create dataset which excludes those with poor wear time etc 


We will record how many participants are excluded at each of the steps, for a flow diagram:
## Only run step below if new working environment
```{r}
dat <- fread("prepped_steps_before_exclude.csv", data.table = FALSE)
```


```{r}
tab_exc <- data.frame("Exclusion" = "Starting cohort", "Number_excluded" = NA, "Number_remaining" = nrow(dat))
```

We do the accelerometer data quality exclusions:

-   Exclude participants without step data:

```{r}
nb <- nrow(dat)
dat <- dat[!is.na(dat$med_steps), ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "No step data",
      "Number_excluded" = nb - nrow(dat),
      "Number_remaining" = nrow(dat)
    )
  )
```

-   Exclude participants whose device could not be calibrated:

```{r}
nb <- nrow(dat)
dat <- dat[dat$quality_good_calibration == "Yes", ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Poor calibration",
      "Number_excluded" = nb - nrow(dat),
      "Number_remaining" = nrow(dat)
    )
  )
```

-   Exclude participants for whom \>1% of values were clipped (fell outside the sensor's range) before or after calibration:

```{r}
nb <- nrow(dat)
dat <- dat[(dat$clips_before_cal < 0.01*dat$total_reads) & (dat$clips_after_cal < 0.01*dat$total_reads) , ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Too many clips",
      "Number_excluded" = nb - nrow(dat),
      "Number_remaining" = nrow(dat)
    )
  )
```

-   Exclude participants who had \<3 days wear or did not have wear in each hour of the 24 hour day:

```{r}
nb <- nrow(dat)
dat <- dat[dat$quality_good_wear_time == "Yes", ] # Note that this has already been calculated in UKB, 
# we don't need to manually calculate it: https://biobank.ndph.ox.ac.uk/showcase/field.cgi?id=90015
# But we might actually use the values from the new data processing

# 2023_01_12 - Now using quality.goodWearTime, which is calcualted from the new data processing
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Poor wear time",
      "Number_excluded" = nb - nrow(dat),
      "Number_remaining" = nrow(dat)
    )
  )
```

-   Exclude participants with unrealistically high overall activity values:

```{r}
nb <- nrow(dat)
dat <- dat[dat$overall_activity < 100, ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Very high overall activity",
      "Number_excluded" = nb - nrow(dat),
      "Number_remaining" = nrow(dat)
    )
  )
```

-   Exclude people lost to follow up before accelerometer wear:

THIS ONE IS REMOVING ALL PARTICIPANTS 
```{r}
nb <- nrow(dat)
dat <- dat[!(dat$date_cens < dat$date_end_accel), ]
tab_exc <- rbind(
  tab_exc,
  data.frame(
    "Exclusion" = "Lost to linked health record follow-up before accelerometer study entry",
    "Number_excluded" = nb - nrow(dat),
    "Number_remaining" = nrow(dat)
  )
)
```


Missing data in adjustment variables:

```{r}
for (
  cov in c(
    "age_entry_years",
    "sex",
    "BMI",
    "ethnicity",
    "tdi_raw",
    "smoking",
    "alcohol"
  )
){
  nb <- nrow(dat)
  print(cov)
  missing_cov <- is.na(dat[, cov])|(as.character(dat[, cov]) == "") |(as.character(dat[, cov]) == "Missing") # for safety coerce to character for second check as can return NA on some classes e.g. Date
  dat <- dat[!missing_cov,]
  tab_exc <- rbind(
    tab_exc,
    data.frame(
      "Exclusion" = paste0("Missing ", cov),
      "Number_excluded" = nb - nrow(dat),
      "Number_remaining" = nrow(dat)
    )
  )
}
```

## Write out

```{r}
write.csv(dat, "prepped_steps.csv")
```


```{r}
print(tab_exc)
```

```{r}
write.csv(tab_exc, "tab_exc.csv")
```



In terminal, run the following code: `dx upload prepped_steps.csv --dest Accelerometry_RAP_Demo/prepped_data.csv`

