---
title: "R Notebook"
output: rmarkdown::github_document
editor_options: 
  markdown: 
    wrap: 72
---

# Data Analysis of UKB Accelerometer Data

In this notebook, we will replicate [this research
paper](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0169649).
This was the paper that first introduced the Accelerometry variables
available in the Showcase in UK Biobank, and is a good general
introduction to the dataset. We briefly review a few key points here,
but that paper is well worth a read!

There is one important difference between that paper and what we'll do
here. Because we've already tidied our data to just work with the subset
of participants without prior cardiovascular disease (as we'll use in
the next notebook), we'll consider just that subset here as well.

This particular notebook has been rewritten from python to R, for using
R studio on the RAP system.

## UK Biobank accelerometry data: a very brief introduction

**What is the UK Biobank accelerometer study?**

Between 2013 and 2015, approximately 100,000 UK Biobank participants
wore a *physical activity monitor*. This *physical activity monitor* was
an [Axivity AX3 research-grade
accelerometer](https://axivity.com/product/ax3), and did not give
feedback to participants.

**Who participated in the UK Biobank accelerometer study?**

Roughly 100,000 participants from the original UK Biobank study.

Recruitment (from [Doherty et
al](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0169649)):

> *Between February 2013 and December 2015, participants who had
> provided a valid email address were sent an email invitation to wear
> an accelerometer for seven days. The participant email addresses were
> chosen randomly, with the exception of the North West region which was
> excluded for much of the project due to participant burden concerns,
> as this area had been used to trial new projects. From June 2013,
> participants were sent devices in order of acceptance.*

**What is an accelerometer?**

An accelerometer is a device to measure acceleration. When the
accelerometer is attached to the body (e.g. worn around the wrist as in
UK Biobank), these acceleration measurements can be used to understand
human movement.

Accelerometers are used in both research-grade and consumer activity
monitors.

**What was the measurement protocol in UK Biobank?**

Please see [Doherty et
al](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0169649).

Participants were posted an Axivity AX3 accelerometer to wear on the
dominant wrist. It was set up to start at 10am two working days after
postal dispatch, and to capture triaxial acceleration data over 7 days
at 100 Hz with a dynamic range of +-8 *g*.

From [Doherty et
al](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0169649):

> *Participants were informed in the invitation email and device
> mail-out letter that the accelerometer should be worn continuously and
> that they should carry on with their normal activities. Participants
> were asked to start wearing the accelerometer immediately after
> receiving it in the post and to wear the monitor on their dominant
> wrist. They were also informed that the device was configured to
> automatically turn itself on soon after its arrival and off seven days
> later. Finally, participants were asked to mail the device back to the
> co-ordinating centre, in a pre-paid envelope, after the seven day
> monitoring period.*

**Did the device record other modalities?**

The device did not record other modalities which may be available in
commercially-available devices (such as heart rate).

The device does record temperature and light, although to our knowledge
these have not been directly used in health applications (temperature is
used in device calibration).

**What can UK Biobank accelerometer data be used for?**

UK Biobank accelerometer data can be used to study many different
phenotypes in conjunction with health and disease outcomes as well as
with genetics:

-   overall physical activity
-   step count
-   physical activity behaviours/ movement behaviours, including sleep,
    sedentary behaviour, light physical activity, moderate-to-vigorous
    physical activity, walking
-   behavioural pattern
-   energy expenditure
-   circadian rhythm
-   sleep quality

**What phenotypes are available?**

Here we will consider only the original accelerometer-based phenotypes
i.e. those that appeared in the initial [Doherty et
al](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0169649)
paper. These are available through the Data Showcase (standard tabular
UK Biobank data). They include metrics relating to data quality (e.g.
calibration and wear time), metrics of overall activity (average
acceleration), metrics of time spent above different acceleration
thresholds (which may be used for so-called 'cut-point' based
definitions of time in different intensities of activity), and metrics
by time of day/day of week. Basic properties of accelerometer wear, such
as the date, are also recorded. Meanings of the different variables are
extensively described both within Showcase and in the paper.

Since then, studies have developed different phenotypes and have applied
their definitions to the UKB accelerometer data. Some of these
phenotypes can now be accessed via 'returns' to the UK Biobank
resources. For example:

-   machine-learning models have been developed to classify movement
    behaviours. See
    [[1](https://www.nature.com/articles/s41598-018-26174-1)];
    [[2](https://www.nature.com/articles/s41467-018-07743-4)]; and
    [[3](https://bjsm.bmj.com/content/early/2022/08/02/bjsports-2021-104050),
    to be released soon].
-   algorithms for [different sleep
    metrics](https://www.nature.com/articles/s41467-019-09576-1)

This probably only represents a small snapshot of what's been done. Let
us know about other resources by adding an issue on GitHub or emailing
us and we will add them to this list.

**I'm interested in a different phenotype. Can I access the raw
accelerometer files to generate it myself?**

Yes. Bulk files for the accelerometer data are available via the RAP
platform. You can also access 'intermediate' files, such as a
[summarisation of the signal each 5
seconds](https://biobank.ndph.ox.ac.uk/showcase/field.cgi?id=90004)
(smaller in size than the raw data, which is a 100 Hz signal i.e. has
100 measurements/second).

You might also be interested in resources available to help process
accelerometer data, such as:

-   <https://github.com/OxWearables/biobankAccelerometerAnalysis>
-   <https://cran.r-project.org/web/packages/GGIR/vignettes/GGIR.html>

Again, get in touch with additions to this list!

If you have other questions about the UKB accelerometer study, feel free
to post them as an issue on GitHub.

We'll now dive into having a look at the data.

## About this notebook

This notebook is written in R.

Most of the material in this notebook is due to Junayed Naushad and
Rosemary Walmsley.

## Set up the session

```{r, include=FALSE}
# First we need to install packages that aren't already present
pkgs <- c("data.table", "plyr", "ggplot2", "reshape2", "MatchIt") # packages we need
pkgs_inst <- pkgs[!{pkgs %in% rownames(installed.packages())}] # check which are not present 
install.packages(pkgs_inst, repos = "https://www.stats.bris.ac.uk/R/") # install

# Load packages
lapply(pkgs, library, character.only = TRUE) # using lapply just allows us to load several packages on one line - this could be replaced with several calls to library()
library(dplyr)

rm(pkgs, pkgs_inst)
```


## Load data

```{r}
dat <- fread("/mnt/project/users/mcgaghd/Data/prepped_steps.csv", data.table = FALSE) # If running in the same session as the previous notebook
```

```{r}
dat <- fread("prepped_steps.csv", data.table = FALSE) # version if running in same session as notebook 2
#dat <- fread("/mnt/project/ssers/mcgaghd/prepped_steps.csv", data.table = FALSE) # version if running in different session
## Summarise and prepare the data
```
Prepped Steps RA- read in the prepped steps from the RA analysis and then remove prevalent RA

```{r}
dat <- dat %>%
  filter(!total_prevalent_RA)
```

We want to isolate those with positive RF 
```{r}
dat <- dat %>%
  filter(!is.na(RF_binary)) %>%
  mutate(RF_binary = ifelse(RF_binary == "Positive", TRUE, FALSE))
```

case-control based on RF positivity to explore whether any prodromal differences in activity
```{r}
caliper <- 0.1


# Perform matching with a caliper for age_entry_years
matched_data <- matchit(RF_binary ~ age_entry_years + season_wear + sex, data = dat, method = "nearest", caliper = caliper,ratio = 2, replace=FALSE)#

summary(matched_data)

# Extract the matched dataset (cases and their matched controls)
datMatch_RheumFactor <- match.data(matched_data)

```

EMS
```{r}
caliper <- 0.1


# Perform matching with a caliper for age_entry_years
matched_data <- matchit(gp_EMS ~ age_entry_years + season_wear + sex, data = dat, method = "nearest", caliper = caliper,ratio = 4, replace=FALSE)#

summary(matched_data)

# Extract the matched dataset (cases and their matched controls)
datMatch_EMS <- match.data(matched_data)

```


PsA cohort 
```{r}
caliper <- 0.1


# Perform matching with a caliper for age_entry_years
matched_data <- matchit(prevalent_PsA_HES_GP ~ age_entry_years + season_wear + sex, data = dat, method = "nearest", caliper = caliper,ratio = 4, replace=FALSE)#

summary(matched_data)

# Extract the matched dataset (cases and their matched controls)
datMatchPsA <- match.data(matched_data)

```

```{r}
datOA <- dat %>% 
  filter(!total_prevalent_AS & !total_prevalent_PsA & !total_prevalent_RA) 

```

```{r}
caliper <- 0.1


# Perform matching with a caliper for age_entry_years
matched_data <- matchit(total_prevalent_OA ~ age_entry_years + season_wear + sex, data =
                          datOA, method = "nearest", caliper = caliper,ratio = 2,
                        replace=FALSE)#

summary(matched_data)

# Extract the matched dataset (cases and their matched controls)
datMatch_OA <- match.data(matched_data)

```


```{r}
dat$PsO_prevalent_coded <- (dat$hes_prevalent_PsO | 
                     dat$PrevalentPsO_gp) 
```

PsO cohort 
```{r}
caliper <- 0.1


# Perform matching with a caliper for age_entry_years
matched_data <- matchit(PrevalentPsO_gp ~ age_entry_years + season_wear + sex, data = dat, method = "nearest", caliper = caliper,ratio = 4, replace=FALSE)#

summary(matched_data)

# Extract the matched dataset (cases and their matched controls)
datMatchPsO <- match.data(matched_data)

```


RA matched cohort 
```{r}
caliper <- 0.1


# Perform matching with a caliper for age_entry_years
matched_data <- matchit(total_prevalent_RA ~ age_entry_years + season_wear + sex, data = dat, method = "nearest", caliper = caliper,ratio = 4, replace=FALSE)#

summary(matched_data)

# Extract the matched dataset (cases and their matched controls)
datMatchRA <- match.data(matched_data)

```

AS matched cohort 
```{r}
caliper <- 0.1


# Perform matching with a caliper for age_entry_years
matched_data <- matchit(total_prevalent_AS ~ age_entry_years + season_wear + sex, data = dat, method = "nearest", caliper = caliper,ratio = 4, replace=FALSE)#

summary(matched_data)

# Extract the matched dataset (cases and their matched controls)
datMatchAS <- match.data(matched_data)

```

```{r}
dat <- datMatchPsA 
```

Following the paper that we're following here, we'll only analyse those
participants aged 45 years and older:

```{r}
print(nrow(dat)) # print number of rows
dat <- dat[dat$age_entry_years >= 45,]
print(nrow(dat)) # print number of rows
```

You might note that there are slight differences in numbers relative to
the published paper. We're using a different cohort, and it looks like
we might have generated the age variable differently (different
precision).

## Convert Fraction of Week to Hours

```{r}
# Column Variables
colvars = names(dat)

# Get the first fraction acceleration column ID
start_loc = match('Fraction acceleration <= 1 milli-gravities',colvars)

# Get the last fraction acceleration column ID
end_loc = match('Fraction acceleration <= 300 milli-gravities',colvars)

# Convert all values in column range to hours
dat[, start_loc:end_loc] = dat[, start_loc:end_loc]*24*7
```

## Wear Duration

Looking at wear duration shows the majority of included participants
have very high wear time:

```{r}
wear_duration <- dat[, 'Wear duration overall']

wear_percentiles <- quantile(wear_duration, c(.25, .5, .75))

cat('25th percentile wear duration: ', round(wear_percentiles[1], digits = 3), 'days\n')
cat('Median wear duration: ', round(wear_percentiles[2], digits = 3), 'days\n')
cat('75th percentile wear duration: ', round(wear_percentiles[3], digits = 3), 'days\n')
```

## Acceleration by Age Group

```{r}
# First we divide participants into age categories
dat$age_cat <- cut(dat$age_entry_years,
                   breaks = c(45, 55, 65, 75, 80),
                   labels = c('45-54', '55-64', '65-74', '75-79'))


p<-ggplot(dat, aes(age_cat, overall_activity)) +
  ggtitle("Acceleration by Age Group") +
  theme(plot.title = element_text(hjust=0.5)) +
  stat_boxplot(geom='errorbar') +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(limits = c(0, quantile(dat$overall_activity, 0.995))) +
  xlab("Age Group [years]") + 
  ylab("Acceleration [mg]")
p
```

## Acceleration by Sex
```{r}
p<-ggplot(dat, aes(sex, overall_activity)) +
  ggtitle("Acceleration by Sex") +
  theme(plot.title = element_text(hjust=0.5)) +
  stat_boxplot(geom='errorbar') +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(limits = c(0, quantile(dat$overall_activity, 0.995))) +
  xlab("Sex") + 
  ylab("Acceleration [mg]")
p
```

```{r}
p<-ggplot(datMatchPsA, aes(total_prevalent_PsA, overall_activity)) +
  ggtitle("Acceleration by PsA group") +
  theme(plot.title = element_text(hjust=0.5)) +
  stat_boxplot(geom='errorbar') +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(limits = c(0, quantile(dat$overall_activity, 0.995))) +
  xlab("Sex") + 
  ylab("Acceleration [mg]")
p
```

## Acceleration by Age Group and Sex
```{r}
p<-ggplot(dat, aes(x=age_cat, y=overall_activity, color=sex)) +
  ggtitle("Acceleration by Age Group and Sex") +
  theme(plot.title = element_text(hjust=0.5)) +
  stat_boxplot(geom='errorbar') +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(limits = c(0, quantile(dat$overall_activity, 0.995))) +
  xlab("Age Group [years]") + 
  ylab("Acceleration [mg]")
p
```
## Acceleration by Day

```{r}
days <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday')

acc_by_day <- dat[, paste(days, 'average acceleration')]
colnames(acc_by_day) <- days

p<-ggplot(melt(acc_by_day), aes(x=variable, y=value)) +
  ggtitle("Acceleration by Day") +
  theme(plot.title = element_text(hjust=0.5)) +
  stat_boxplot(geom='errorbar') +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(limits = c(0, quantile(dat$overall_activity, 0.995))) +
  xlab("Day of Week") + 
  ylab("Acceleration [mg]")
p
```
## Acceleration Weekday/Weekend
```{r}
acc_by_day$Weekday <- rowMeans(acc_by_day[, c("Monday", "Tuesday", "Wednesday", 
                                             "Thursday", "Friday")])
acc_by_day$Weekend <- rowMeans(acc_by_day[, c("Saturday", "Sunday")])

week_vs_weekend = acc_by_day[, c('Weekday', 'Weekend')]

p<-ggplot(melt(week_vs_weekend), aes(x=variable, y=value)) +
  ggtitle("Acceleration by Weekday/Weekend") +
  theme(plot.title = element_text(hjust=0.5)) +
  stat_boxplot(geom='errorbar') +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(limits = c(0, quantile(dat$overall_activity, 0.995))) +
  xlab("") + 
  ylab("Acceleration [mg]")
p
```

## Acceelation by Season
```{r}
dat$end_month <- factor(strftime(dat$date_end_accel, "%m"))

# Define seasons based on months.
season.levels <- list(
  Spring = c('03', '04', '05'),
  Summer = c('06', '07', '08'),
  Autumn = c('09', '10', '11'),
  Winter = c('12', '01', '02'))

dat$season <- `levels<-`(dat$end_month, season.levels)

p<-ggplot(dat, aes(season, overall_activity)) +
  ggtitle("Acceleration by Season") +
  theme(plot.title = element_text(hjust=0.5)) +
  stat_boxplot(geom='errorbar') +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(limits = c(0, quantile(dat$overall_activity, 0.995))) +
  xlab("Season") + 
  ylab("Acceleration [mg]")
p
```



```{r}
# set up tick marks
hours <- c('12am', paste0(1:11,'am'), '12pm', paste0(1:11,'pm'))

# Column Variables
colvars <- names(dat)

# Get the first fraction acceleration column ID
start_loc <- match('Average acceleration 00:00 - 00:59', colvars)

# Get the last fraction acceleration column ID
end_loc <- match('Average acceleration 23:00 - 23:59', colvars)

dat_hour <- dat[, c(1, start_loc:end_loc)]
names(dat_hour)[-1] <- hours

dat_hour <- merge(
  melt(dat_hour, id.vars = 'eid'),
  dat[, c('eid', 'AnalysisLabel')],
  by = 'eid',
  all.x = TRUE
)

# Group by function for dataframe in R using pipe operator 
dat_hour_means <- dat_hour %>% 
  group_by(variable, AnalysisLabel) %>% 
  summarise_at(vars(value), funs(mean(., na.rm = TRUE)))

p <- ggplot(dat_hour_means, aes(x = variable, y = value, color = AnalysisLabel)) +
  theme_bw() + 
  ggtitle("Acceleration by Time of Day") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_line(aes(group=AnalysisLabel)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        text = element_text(size = 10)) +
  xlab("Hour of Day") + 
  ylab("Acceleration [mg]")

p

```

EMS
```{r}
# set up tick marks
#sandbox

 
datMatch_EMS <- datMatch_EMS %>%
  mutate(EMS_binary = ifelse(gp_EMS, "EMS", "Controls"))

hours <- c('12am', paste0(1:11,'am'), '12pm', paste0(1:11,'pm'))

# Column Variables
colvars = names(datMatch_EMS)

# Get the first fraction acceleration column ID
start_loc = match('Average acceleration 00:00 - 00:59',colvars)

# Get the last fraction acceleration column ID
end_loc = match('Average acceleration 23:00 - 23:59',colvars)

dat_hour <- datMatch_EMS[, c(1, start_loc:end_loc)]
dat_hour <- as.data.table(dat_hour)
names(dat_hour)[-1] <- hours


dat_hour <- melt(dat_hour, id.vars = 'eid')
         
subset_datMatch_EMS <- datMatch_EMS %>%
  select(eid, EMS_binary)  

#Now join up dat_hour and subset by eid
dat_hour <- dat_hour %>%
  left_join(subset_datMatch_EMS, by = "eid") %>%
  group_by(eid) %>%
  mutate(EMS_binary = first(EMS_binary))

dat_hour_means <- dat_hour %>% 
  group_by(variable, EMS_binary) %>% 
  summarise_at(vars(value), funs(mean(., na.rm = TRUE)))

EMS_timing <- ggplot(dat_hour_means, aes(x = variable, y = value, color =
                                           EMS_binary)) +  theme_bw() + 
  ggtitle("Acceleration by Time of Day") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_line(aes(group=EMS_binary)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        text = element_text(size = 10)) +
  xlab("Hour of Day") + 
  ylab("Acceleration [mg]")

EMS_timing


```

```{r}
# set up tick marks
#sandbox


datMatch_RheumFactor <- datMatch_RheumFactor %>%
  mutate(RF_binary = ifelse(RF_binary, "RF+ve", "RF-ve"))

hours <- c('12am', paste0(1:11,'am'), '12pm', paste0(1:11,'pm'))

# Column Variables
colvars = names(datMatch_RheumFactor)

# Get the first fraction acceleration column ID
start_loc = match('Average acceleration 00:00 - 00:59',colvars)

# Get the last fraction acceleration column ID
end_loc = match('Average acceleration 23:00 - 23:59',colvars)

dat_hour <- datMatch_RheumFactor[, c(1, start_loc:end_loc)]
names(dat_hour)[-1] <- hours


dat_hour <- melt(dat_hour, id.vars = 'eid')
         
subset_datMatch_RheumFactor <- datMatch_RheumFactor %>%
  select(eid, RF_binary)  

#Now join up dat_hour and subset by eid
dat_hour <- dat_hour %>%
  left_join(subset_datMatch_RheumFactor, by = "eid") %>%
  group_by(eid) %>%
  mutate(RF_binary = first(RF_binary))

dat_hour_means <- dat_hour %>% 
  group_by(variable, RF_binary) %>% 
  summarise_at(vars(value), funs(mean(., na.rm = TRUE)))

RheumFactor_timing <- ggplot(dat_hour_means, aes(x = variable, y = value, color =
                                           RF_binary)) +  theme_bw() + 
  ggtitle("Acceleration by Time of Day") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_line(aes(group=RF_binary)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        text = element_text(size = 10)) +
  xlab("Hour of Day") + 
  ylab("Acceleration [mg]")


```


```{r}
# set up tick marks
#sandbox


datMatch_OA <- datMatch_OA %>%
  mutate(total_prevalent_OA = ifelse(total_prevalent_OA, "OA", "Control"))

hours <- c('12am', paste0(1:11,'am'), '12pm', paste0(1:11,'pm'))

# Column Variables
colvars = names(datMatch_OA)

# Get the first fraction acceleration column ID
start_loc = match('Average acceleration 00:00 - 00:59',colvars)

# Get the last fraction acceleration column ID
end_loc = match('Average acceleration 23:00 - 23:59',colvars)

dat_hour <- datMatch_OA[, c(2, start_loc:end_loc)]
names(dat_hour)[-1] <- hours


dat_hour <- melt(dat_hour, id.vars = 'eid')
         
subset_datMatch_OA <- datMatch_OA %>%
  select(eid, total_prevalent_OA)  

#Now join up dat_hour and subset by eid
dat_hour <- dat_hour %>%
  left_join(subset_datMatch_OA, by = "eid") %>%
  group_by(eid) %>%
  mutate(total_prevalent_OA = first(total_prevalent_OA))

dat_hour_means <- dat_hour %>% 
  group_by(variable, total_prevalent_OA) %>% 
  summarise_at(vars(value), funs(mean(., na.rm = TRUE)))

OA_timing <- ggplot(dat_hour_means, aes(x = variable, y = value, color =
                                           total_prevalent_OA)) +  theme_bw() + 
  ggtitle("Acceleration by Time of Day") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_line(aes(group=total_prevalent_OA)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        text = element_text(size = 10)) +
  xlab("Hour of Day") + 
  ylab("Acceleration [mg]")
OA_timing

 
```





Case control analysis acceleration PsA vs controls 
```{r}
# set up tick marks
#sandbox


datMatchPsA <- datMatchPsA %>%
  mutate(prevalent_PsA_HES_GP = ifelse(prevalent_PsA_HES_GP, "PsA", "Control"))

hours <- c('12am', paste0(1:11,'am'), '12pm', paste0(1:11,'pm'))

# Column Variables
colvars = names(datMatchPsA)

# Get the first fraction acceleration column ID
start_loc = match('Average acceleration 00:00 - 00:59',colvars)

# Get the last fraction acceleration column ID
end_loc = match('Average acceleration 23:00 - 23:59',colvars)

dat_hour <- datMatchPsA[, c(1, start_loc:end_loc)]
names(dat_hour)[-1] <- hours


dat_hour <- melt(dat_hour, id.vars = 'eid')
         
subset_datMatchPsA <- datMatchPsA %>%
  select(eid, prevalent_PsA_HES_GP)  

#Now join up dat_hour and subset by eid
dat_hour <- dat_hour %>%
  left_join(subset_datMatchPsA, by = "eid") %>%
  group_by(eid) %>%
  mutate(prevalent_PsA_HES_GP = first(prevalent_PsA_HES_GP))

dat_hour_means <- dat_hour %>% 
  group_by(variable, prevalent_PsA_HES_GP) %>% 
  summarise_at(vars(value), funs(mean(., na.rm = TRUE)))

PsA_timing <- ggplot(dat_hour_means, aes(x = variable, y = value, color =
                                           prevalent_PsA_HES_GP)) +  theme_bw() + 
  ggtitle("Acceleration by Time of Day") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_line(aes(group=prevalent_PsA_HES_GP)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        text = element_text(size = 10)) +
  xlab("Hour of Day") + 
  ylab("Acceleration [mg]")

PsA_timing
```

```{r}
# set up tick marks
#sandbox


datMatchPsO <- datMatchPsO %>%
  mutate(PrevalentPsO_gp = ifelse(PrevalentPsO_gp, "Psoriasis", "Control"))

hours <- c('12am', paste0(1:11,'am'), '12pm', paste0(1:11,'pm'))

# Column Variables
colvars = names(datMatchPsO)

# Get the first fraction acceleration column ID
start_loc = match('Average acceleration 00:00 - 00:59',colvars)

# Get the last fraction acceleration column ID
end_loc = match('Average acceleration 23:00 - 23:59',colvars)

dat_hour <- datMatchPsO[, c(1, start_loc:end_loc)]
names(dat_hour)[-1] <- hours


dat_hour <- melt(dat_hour, id.vars = 'eid')
         
subset_datMatchPsO <- datMatchPsO %>%
  select(eid, PrevalentPsO_gp)  

#Now join up dat_hour and subset by eid
dat_hour <- dat_hour %>%
  left_join(subset_datMatchPsO, by = "eid") %>%
  group_by(eid) %>%
  mutate(PrevalentPsO_gp = first(PrevalentPsO_gp))

dat_hour_means <- dat_hour %>% 
  group_by(variable, PrevalentPsO_gp) %>% 
  summarise_at(vars(value), funs(mean(., na.rm = TRUE)))

PsO_timing <- ggplot(dat_hour_means, aes(x = variable, y = value, color =
                                           PrevalentPsO_gp)) +  theme_bw() + 
  ggtitle("Acceleration by Time of Day") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_line(aes(group=PrevalentPsO_gp)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        text = element_text(size = 10)) +
  xlab("Hour of Day") + 
  ylab("Acceleration [mg]")


```


```{r}
dir.create('outputs', showWarnings = FALSE)  
```

```{r}
# Save ggplot object as SVG in the "/outputs/" folder
output_folder <- here::here("outputs")
svg_file <- file.path(output_folder, paste0("PsA_timing_acceleration.svg"))
ggsave(filename = svg_file, plot = PsA_timing, device = "svg", width = 8, height = 5)
```

case control for RA 

```{r}
# set up tick marks
#sandbox

datMatchAS <- datMatchAS %>%
  mutate(total_prevalent_AS = ifelse(total_prevalent_AS, "AS", "Control"))

hours <- c('12am', paste0(1:11,'am'), '12pm', paste0(1:11,'pm'))

# Column Variables
colvars = names(datMatchAS)

# Get the first fraction acceleration column ID
start_loc = match('Average acceleration 00:00 - 00:59',colvars)

# Get the last fraction acceleration column ID
end_loc = match('Average acceleration 23:00 - 23:59',colvars)

dat_hour <- datMatchAS[, c(2, start_loc:end_loc)]
names(dat_hour)[-1] <- hours


dat_hour <- melt(dat_hour, id.vars = 'eid')
         
subset_datMatchAS <- datMatchAS %>%
  select(eid, total_prevalent_AS)  

#Now join up dat_hour and subset by eid
dat_hour <- dat_hour %>%
  left_join(subset_datMatchAS, by = "eid") %>%
  group_by(eid) %>%
  mutate(total_prevalent_AS = first(total_prevalent_AS))

dat_hour_means <- dat_hour %>% 
  group_by(variable, total_prevalent_AS) %>% 
  summarise_at(vars(value), funs(mean(., na.rm = TRUE)))

timing_AS <- ggplot(dat_hour_means, aes(x = variable, y = value, color = total_prevalent_AS)) +
  theme_bw() + 
  ggtitle("Acceleration by Time of Day") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_line(aes(group=total_prevalent_AS)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        text = element_text(size = 10)) +
  xlab("Hour of Day") + 
  ylab("Acceleration [mg]")

p
```

```{r}
# set up tick marks
#sandbox

datMatchRA <- datMatchRA %>%
  mutate(total_prevalent_RA = ifelse(total_prevalent_RA, "RA", "Control"))

hours <- c('12am', paste0(1:11,'am'), '12pm', paste0(1:11,'pm'))

# Column Variables
colvars = names(datMatchRA)

# Get the first fraction acceleration column ID
start_loc = match('Average acceleration 00:00 - 00:59',colvars)

# Get the last fraction acceleration column ID
end_loc = match('Average acceleration 23:00 - 23:59',colvars)

dat_hour <- datMatchRA[, c(2, start_loc:end_loc)]
names(dat_hour)[-1] <- hours


dat_hour <- melt(dat_hour, id.vars = 'eid')
         
subset_datMatchRA <- datMatchRA %>%
  select(eid, total_prevalent_RA)  

#Now join up dat_hour and subset by eid
dat_hour <- dat_hour %>%
  left_join(subset_datMatchRA, by = "eid") %>%
  group_by(eid) %>%
  mutate(total_prevalent_RA = first(total_prevalent_RA))

dat_hour_means <- dat_hour %>% 
  group_by(variable, total_prevalent_RA) %>% 
  summarise_at(vars(value), funs(mean(., na.rm = TRUE)))

timing_RA <- ggplot(dat_hour_means, aes(x = variable, y = value, color = total_prevalent_RA)) +
  theme_bw() + 
  ggtitle("Acceleration by Time of Day") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_line(aes(group=total_prevalent_RA)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        text = element_text(size = 10)) +
  xlab("Hour of Day") + 
  ylab("Acceleration [mg]")

timing_RA

```

```{r}
# Save ggplot object as SVG in the "/outputs/" folder
output_folder <- here::here("outputs")
svg_file <- file.path(output_folder, paste0("PsA_timing_acceleration.svg"))
ggsave(filename = svg_file, plot = timing_RA, device = "svg", width = 8, height = 5)
```

PsA vs PsO
```{r}
# set up tick marks
#sandbox

datMatchPsA <- datMatchPsA %>%
  mutate(total_prevalent_PsA = ifelse(total_prevalent_PsA, "PsA", "Psoriasis"))

hours <- c('12am', paste0(1:11,'am'), '12pm', paste0(1:11,'pm'))

# Column Variables
colvars = names(datMatchPsA)

# Get the first fraction acceleration column ID
start_loc = match('Average acceleration 00:00 - 00:59',colvars)

# Get the last fraction acceleration column ID
end_loc = match('Average acceleration 23:00 - 23:59',colvars)

dat_hour <- datMatchPsA[, c(2, start_loc:end_loc)]
names(dat_hour)[-1] <- hours


dat_hour <- melt(dat_hour, id.vars = 'eid')
         
subset_datMatchPsA <- datMatchPsA %>%
  select(eid, total_prevalent_PsA)  

#Now join up dat_hour and subset by eid
dat_hour <- dat_hour %>%
  left_join(subset_datMatchPsA, by = "eid") %>%
  group_by(eid) %>%
  mutate(total_prevalent_PsA = first(total_prevalent_PsA))

dat_hour_means <- dat_hour %>% 
  group_by(variable, total_prevalent_PsA) %>% 
  summarise_at(vars(value), funs(mean(., na.rm = TRUE)))

timing_PsA <- ggplot(dat_hour_means, aes(x = variable, y = value, color = total_prevalent_PsA)) +
  theme_bw() + 
  ggtitle("Acceleration by Time of Day") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_line(aes(group=total_prevalent_PsA)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        text = element_text(size = 10)) +
  xlab("Hour of Day") + 
  ylab("Acceleration [mg]")

timing_PsA

```

```{r}
# Save ggplot object as SVG in the "/outputs/" folder
output_folder <- here::here("outputs")
svg_file <- file.path(output_folder, paste0("PsA_timing_acceleration.svg"))
ggsave(filename = svg_file, plot = timing_RA, device = "svg", width = 8, height = 5)
```

datMatchAS


```{r}
# set up tick marks
hours <- c('12am', paste0(1:11,'am'), '12pm', paste0(1:11,'pm'))

# Column Variables
colvars = names(dat)

# Get the first fraction acceleration column ID
start_loc = match('Average acceleration 00:00 - 00:59',colvars)

# Get the last fraction acceleration column ID
end_loc = match('Average acceleration 23:00 - 23:59',colvars)

dat_hour <- dat[, c(1, start_loc:end_loc)]
names(dat_hour)[-1] <- hours

dat_hour <- merge(melt(dat_hour, id.vars='eid'), 
                  dat[, c('eid', 'total_prevalent_PsA')], 
                  by='eid', all.x=TRUE)

# Group by function for dataframe in R using pipe operator 
dat_hour_means <- dat_hour %>% group_by(variable, total_prevalent_PsA) %>% summarise_at(vars(value),funs(mean(.,na.rm=TRUE)))

p<-ggplot(dat_hour_means[dat_hour_means$total_prevalent_PsA=='TRUE',], aes(x=variable, y=value, color=total_prevalent_PsA)) +
  theme_bw() + 
  ggtitle("Acceleration by Time of Day - PsA") +
  theme(plot.title = element_text(hjust=0.5)) +
  geom_line(aes(group=total_prevalent_PsA)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust=1), 
        text = element_text(size=10)) +
  xlab("Hour of Day") + 
  ylab("Acceleration [mg]")
p
```

```{r}
# set up tick marks
hours <- c('12am', paste0(1:11,'am'), '12pm', paste0(1:11,'pm'))

# Column Variables
colvars = names(dat)

# Get the first fraction acceleration column ID
start_loc = match('Average acceleration 00:00 - 00:59',colvars)

# Get the last fraction acceleration column ID
end_loc = match('Average acceleration 23:00 - 23:59',colvars)

dat_hour <- dat[, c(1, start_loc:end_loc)]
names(dat_hour)[-1] <- hours

dat_hour <- merge(melt(dat_hour, id.vars='eid'), 
                  dat[, c('eid', 'AnalysisLabel', 'sex')], 
                  by='eid', all.x=TRUE)

# Group by function for dataframe in R using pipe operator 
dat_hour_means <- dat_hour %>% group_by(variable, AnalysisLabel) %>% summarise_at(vars(value),funs(mean(.,na.rm=TRUE)))

p<-ggplot(dat_hour_means[dat_hour_means], aes(x=variable, y=value, color=AnalysisLabel)) +
  theme_bw() + 
  ggtitle("Acceleration by Time of Day") +
  theme(plot.title = element_text(hjust=0.5)) +
  geom_line(aes(group=AnalysisLabel)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust=1), 
        text = element_text(size=10)) +
  xlab("Hour of Day") + 
  ylab("Acceleration [mg]")
p
```

## Acceleration by Time of Day
```{r}
# set up tick marks
hours <- c('12am', paste0(1:11,'am'), '12pm', paste0(1:11,'pm'))

# Column Variables
colvars = names(dat)

# Get the first fraction acceleration column ID
start_loc = match('Average acceleration 00:00 - 00:59',colvars)

# Get the last fraction acceleration column ID
end_loc = match('Average acceleration 23:00 - 23:59',colvars)

dat_hour <- dat[, c(1, start_loc:end_loc)]
names(dat_hour)[-1] <- hours

dat_hour <- merge(melt(dat_hour, id.vars='eid'), 
                  dat[, c('eid', 'AnalysisLabel')], 
                  by='eid', all.x=TRUE)

# Group by function for dataframe in R using pipe operator 
dat_hour_means <- dat_hour %>% group_by(variable, AnalysisLabel) %>% summarise_at(vars(value),funs(mean(.,na.rm=TRUE)))

p<-ggplot(dat_hour_means[dat_hour_means$AnalysisLabel=='NonPsoriatic',], aes(x=variable, y=value, color=AnalysisLabel)) +
  theme_bw() + 
  ggtitle("Acceleration by Time of Day - NonPsoriatic") +
  theme(plot.title = element_text(hjust=0.5)) +
  geom_line(aes(group=AnalysisLabel)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust=1), 
        text = element_text(size=10)) +
  xlab("Hour of Day") + 
  ylab("Acceleration [mg]")
p


p<-ggplot(dat_hour_means[dat_hour_means$sex=='Male',], aes(x=variable, y=value, color=age_cat)) +
  theme_bw() + 
  ggtitle("Acceleration by Time of Day - Male") +
  theme(plot.title = element_text(hjust=0.5)) +
  geom_line(aes(group=age_cat)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, hjust=1), 
        text = element_text(size=10)) +
  xlab("Hour of Day") + 
  ylab("Acceleration [mg]")
p
```

In this code, we filter the dat_hour_means dataframe to include only the desired "AnalysisLabel" values ("NonPsoriatic," "PsoriasisOnly," and "PsA"), and then we create a single ggplot to visualize the average acceleration across these groups. The color aesthetic is set to "variable," so each group will have a different color in the plot.
```{r}
desired_labels <- c("NonPsoriatic", "PsoriasisOnly", "PsA")
dat_filtered <- dat %>% filter(AnalysisLabel %in% desired_labels)

# Create a ggplot to visualize the average acceleration
p <- ggplot(dat_filtered, aes(x = variable, y = value, color = AnalysisLabel)) +
  theme_bw() +
  ggtitle("Average Acceleration by AnalysisLabel") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_line() +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        text = element_text(size = 10)) +
  xlab("Hour of Day") +
  ylab("Average Acceleration [mg]")

p



```

## Cumulative time spent at acceleration levels below certain values by sex and age

We'll look at cumulative time spent up to particular acceleration values, split by sex and age. As the acceleration range spanned is large, we'll make two cumulative plots, with different levels of granularity. This allows us to see both the detail at the lower end and the big picture. So we can easily repeat across groups, we'll write some functions to make the plots:
```{r}
# Column Variables
colvars = names(dat)

# Get the first fraction acceleration column ID
start_loc = match('Fraction acceleration <= 1 milli-gravities', colvars)

# Get the last fraction acceleration column ID
end_loc = match('Fraction acceleration <= 300 milli-gravities', colvars)

dat_cummulative <- dat[, c(1, seq(start_loc, end_loc, by=2))]

# pattern is extract value
names(dat_cummulative)[-1] <- as.numeric(gsub("Fraction acceleration <= | milli-gravities", 
                                              "", 
                                       names(dat_cummulative)[-1]))

dat_cummulative <- merge(melt(dat_cummulative, id.vars='eid'), 
                         dat[, c('eid', 'age_cat', 'sex')], 
                         by='eid', all.x=TRUE)

p<-ggplot(dat_cummulative[dat_cummulative$age_cat=='45-54',], 
          aes(x=variable, y=value, color=sex)) +
  ggtitle("Acceleration by Time of Day - 45-54") +
  theme(plot.title = element_text(hjust=0.5)) +
  stat_boxplot(geom='errorbar') +
  geom_boxplot(outlier.shape = NA) +
  xlab("Acceleration level [mg]") + 
  ylab("Cummulative time (hours/week)")
p

p<-ggplot(dat_cummulative[dat_cummulative$age_cat=='55-64',], aes(x=variable, y=value, color=sex)) +
  ggtitle("Acceleration by Time of Day - 55-64") +
  theme(plot.title = element_text(hjust=0.5)) +
  stat_boxplot(geom='errorbar') +
  geom_boxplot(outlier.shape = NA) +
  xlab("Acceleration level [mg]") + 
  ylab("Cummulative time (hours/week)")
p

p<-ggplot(dat_cummulative[dat_cummulative$age_cat=='65-74',], aes(x=variable, y=value, color=sex)) +
  ggtitle("Acceleration by Time of Day - 65-74") +
  theme(plot.title = element_text(hjust=0.5)) +
  stat_boxplot(geom='errorbar') +
  geom_boxplot(outlier.shape = NA) +
  xlab("Acceleration level [mg]") + 
  ylab("Cummulative time (hours/week)")
p

p<-ggplot(dat_cummulative[dat_cummulative$age_cat=='75-79',], aes(x=variable, y=value, color=sex)) +
  ggtitle("Acceleration by Time of Day - 75-79") +
  theme(plot.title = element_text(hjust=0.5)) +
  stat_boxplot(geom='errorbar') +
  geom_boxplot(outlier.shape = NA) +
  xlab("Acceleration level [mg]") + 
  ylab("Cummulative time (hours/week)")
p
```

```{r}
# set up tick marks
hours <- c('12am', paste0(1:11,'am'), '12pm', paste0(1:11,'pm'))

# Column Variables
colvars = names(datMatchPsA)

# Get the first fraction acceleration column ID
start_loc = match('Average acceleration 00:00 - 00:59',colvars)

# Get the last fraction acceleration column ID
end_loc = match('Average acceleration 23:00 - 23:59',colvars)

dat_hour <- datMatchPsA[, c(1, start_loc:end_loc)]
names(dat_hour)[-1] <- hours

dat_hour <- dat_hour %>%
  rename(eid = V1)

dat_hour <- merge(melt(dat_hour, id.vars='eid'), 
                  datMatchPsA[, c('eid', 'total_prevalent_PsA')], 
                  by='eid', all.x=TRUE)

# Group by function for dataframe in R using pipe operator 
dat_hour_means <- dat_hour %>% group_by(variable, total_prevalent_PsA) %>% summarise_at(vars(value),funs(mean(.,na.rm=TRUE)))

p<-ggplot(dat_hour_means[dat_hour_means$total_prevalent_PsA=='TRUE',], aes(x=variable, y=value, color=total_prevalent_PsA)) +
  theme_bw() + 
  ggtitle("Acceleration by Time of Day") +
  theme(plot.title = element_text(hjust=0.5)) +
  geom_line(aes(group=total_prevalent_PsA)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust=1), 
        text = element_text(size=10)) +
  xlab("Hour of Day") + 
  ylab("Acceleration [mg]")
p


p<-ggplot(dat_hour_means[dat_hour_means$sex=='Male',], aes(x=variable, y=value, color=age_cat)) +
  theme_bw() + 
  ggtitle("Acceleration by Time of Day - Male") +
  theme(plot.title = element_text(hjust=0.5)) +
  geom_line(aes(group=age_cat)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, hjust=1), 
        text = element_text(size=10)) +
  xlab("Hour of Day") + 
  ylab("Acceleration [mg]")
p
```
