---
title: "ExtractingGPdata"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "R Notebook"
output: rmarkdown::github_document
---

# Further data preparation in R 
testing

## Introduction

In this notebook, we will prepare data for analysis.

Extract GP data 


## How to run this notebook

This notebook should be run in a *RStudio* session. It does *not* require a Spark cluster. See how to set it up [here](https://dnanexus.gitbook.io/uk-biobank-rap/working-on-the-research-analysis-platform/using-rstudio-on-the-research-analysis-platform).

## Set up the session

We load packages we'll use:

```{r, include=FALSE}
# First we need to install packages that aren't already present
pkgs <- c("data.table", "plyr", "dplyr","stringr") # packages we need
pkgs_inst <- pkgs[!{pkgs %in% rownames(installed.packages())}] # check which are not present 
install.packages(pkgs_inst, repos = "https://www.stats.bris.ac.uk/R/") # install

# Load packages
lapply(pkgs, library, character.only = TRUE) # using lapply just allows us to load several packages on one line - this could be replaced with several calls to library()
```

dx download - downloads stored permanent storage RAP data into my temporary session
Needs bash to run 
```{bash}
cd ~/rap_wearables

dx download /users/mcgaghd/PrimaryCare -r
```

```{r}
dat_GPevent <- fread("~/rap_wearables/PrimaryCare/gp_event_data.csv", data.table = FALSE)
dat_GPmap <- fread("~/rap_wearables/PrimaryCare/gp_r2_map_r3.csv", data.table = FALSE)
```

Filter for PsA - Read codes -> all subtypes of M160, and juvenile codes  

```{r}
psori_table <- dat_GPmap %>%
  filter(grepl("psoriasis", READV2_DESC, ignore.case = TRUE))
```


```{r}
PsA_V3 <- dat_GPmap %>%
  filter(!is.na(READV3_CODE) & str_detect(READV3_CODE, "^(?i)(M160(.*)|M1600|M160z|Nyu13|X701u|X701v|X701w|X7026|X7027|X7028)")) %>%
  arrange(TERMV2_ORDER)
```

READV3 codes for PsA 
```{r}
PsA_V3 <- dat_GPmap %>%
  filter(!is.na(READV3_CODE) & str_detect(READV3_CODE, "^(?i)(M160(.*)|M1600|M160z|Nyu13|X701u|X701v|X701w|X7026|X7027|X7028)")) %>%
  arrange(TERMV2_ORDER)
```

Term3 codes for PsA 
```{r}
#READ V3 codes 

PsA_Read3 <- dat_GPmap %>%
  filter(!is.na(TERMV3_CODE) & str_detect(TERMV3_CODE, "^(Y70Eo|Y70Es|Y70Et|YMAqA|Y70Ex|Y70Er|YMAq6|YMAMW|Y70En)")) %>%
  arrange(TERMV2_ORDER)
```

ReadV2 codes for psoriasis
```{r}
PsO_V2 <- dat_GPmap %>%
  filter(!is.na(READV2_CODE) & str_detect(READV2_CODE, "^(^M16[^0|^2|^3|^4|^5].*$|Myu30|14F2.)")) %>%
  arrange(TERMV2_ORDER)
```

TERMV3_CODE for psoriasis

```{r}
PsO_V3 <- dat_GPmap %>%
  filter(!is.na(READV3_CODE) & str_detect(READV3_CODE, "^M16[^0|^2|^3|^4|^5].*$|Myu30|X506Y|X506Z|X506a|X506b|X506c|X506d|X506e|X506f|X506g|X506h|X506i|X506j|X506k|X506l|X506m|X506n|X506o|X506p|X506q|X506r|X506s|X506t|X506u|X506v|XE1CQ|XaNXa|XaNXt|XaYOx")) %>%
  arrange(TERMV2_ORDER)
```


Now create a Read v2, CTV3 and v3 list for each 

Psoriasis first 
```{r}
PsO_Read2Strings<- "^(^M16[^0|^2|^3|^4|^5].*$|Myu30|14F2.)"

# Create the new column "PsO_v2" based on the matching condition
PsO_Read2 <- dat_GPevent %>%
  filter(grepl(PsO_Read2Strings, read_2, ignore.case = TRUE))

# Show the first few rows of the 'PsO_V2' dataset for verification
```

```{r}
PsO_Read3strings <- "^M16[^0|^2|^3|^4|^5].*$|Myu30|X506Y|X506Z|X506a|X506b|X506c|X506d|X506e|X506f|X506g|X506h|X506i|X506j|X506k|X506l|X506m|X506n|X506o|X506p|X506q|X506r|X506s|X506t|X506u|X506v|XE1CQ|XaNXa|XaNXt|XaYOx"


# Use data.table's fast filter function
PsO_Read3 <- dat_GPevent %>%
  filter(grepl(PsO_Read3strings, read_3, ignore.case = TRUE))
```

```{r}
PsO_Read2 <- PsO_Read2 %>%
  filter(!is.na(event_dt))
PsO_Read2[is.na(PsO_Read2)] <- ""

PsO_Read3 <- PsO_Read3 %>%
  filter(!is.na(event_dt))
PsO_Read3[is.na(PsO_Read3)] <- ""
```


Data cleaning. As we're confident in our codes from data extract script. We can just create a ew column of Psoriasis T/F or PsA TRUE/FALSE

```{r}
PsO_Read3$gp_PsO <- TRUE
PsO_Read2$gp_PsO <- TRUE
```

```{r}
PsO_Read3subset <- PsO_Read3[, c("eid", "data_provider", "event_dt", "gp_PsO")]
PsO_Read2subset <- PsO_Read2[, c("eid", "data_provider", "event_dt", "gp_PsO")]

# Append the two dataframes together
datPsO_gp <- rbind(PsO_Read2subset, PsO_Read3subset)
```

```{r}
datPsO_gp <- datPsO_gp %>%
  filter(event_dt > as.Date("1940-01-01"))
```

```{r}
write.csv(datPsO_gp, "datPsO_gp.csv")
```

```{bash}
dx upload datPsO_gp.csv --dest /users/mcgaghd/PrimaryCare/datPsO_gp.csv
```

PsA Read V2 
```{r} 
PsA_Read2Strings <- "^(?i)(M160(.*)|M1602|N0452|Nyu13)"


# Use data.table's fast filter function
PsA_Read2 <- dat_GPevent %>%
  filter(grepl(PsA_Read2Strings, read_2, ignore.case = TRUE))
```

```{r}
PsA_Read3Strings  <-"^(?i)(M160(.*)|M1600|M160z|Nyu13|X701u|X701v|X701w|X7026|X7027|X7028)"

PsA_Read3 <- dat_GPevent %>%
  filter(grepl(PsA_Read3Strings, read_3, ignore.case = TRUE))
```

```{r}
PsA_Read2 <- PsA_Read2 %>%
  filter(!is.na(event_dt))
PsA_Read2[is.na(PsA_Read2)] <- ""

PsA_Read3 <- PsA_Read3 %>%
  filter(!is.na(event_dt))
PsA_Read3[is.na(PsA_Read3)] <- ""
```


Data cleaning. As we're confident in our codes from data extract script. We can just create a ew column of Psoriasis T/F or PsA TRUE/FALSE

```{r}
PsA_Read3$gp_PsA <- TRUE
PsA_Read2$gp_PsA <- TRUE
```

```{r}
PsA_Read3subset <- PsA_Read3[, c("eid", "data_provider", "event_dt", "gp_PsA")]
PsA_Read2subset <- PsA_Read2[, c("eid", "data_provider", "event_dt", "gp_PsA")]

# Append the two dataframes together
datPsA_gp <- rbind(PsA_Read2subset, PsA_Read3subset)
```

```{r}
datPsA_gp <- datPsA_gp %>%
  filter(event_dt > as.Date("1940-01-01"))
```

```{r}
write.csv(datPsA_gp, "datPsA_gp.csv")
```

```{bash}
dx upload datPsA_gp.csv --dest /users/mcgaghd/PrimaryCare/datPsA_gp.csv
```


#Morning stiffness 


```{r}
MorningStiffness <- dat_GPmap %>%
  filter(!is.na(READV3_CODE) & str_detect(TERMV3_CODE, "^(Ya1np)")) %>%
  arrange(TERMV2_ORDER)
```

```{r}
MorningStiffnessStrings  <-"(^Xa0YV|^Ya1np|^1D17.$)"

MorningStiffness_Read2 <- dat_GPevent %>%
  filter(grepl(MorningStiffnessStrings, read_2, ignore.case = TRUE))
```

```{r}
# Use data.table's fast filter function
MorningStiffness_Read3 <- dat_GPevent %>%
  filter(grepl(MorningStiffnessStrings, read_3, ignore.case = TRUE))
```

```{r}
MorningStiffness_Read2$gp_EMS <- TRUE
MorningStiffness_Read3$gp_EMS <- TRUE
```

```{r}
EMS_read3subset <- MorningStiffness_Read3[, c("eid", "data_provider", "event_dt", "gp_EMS")]
EMS_read2subset <- MorningStiffness_Read2[, c("eid", "data_provider", "event_dt", "gp_EMS")]

# Append the two dataframes together
datEMS_gp <- rbind(EMS_read3subset, EMS_read2subset)
```
 
```{r}
write.csv(datEMS_gp, "datEMS_gp.csv")
```

```{bash}
dx upload datEMS_gp.csv --dest /users/mcgaghd/PrimaryCare/datEMS_gp.csv
```


