---
title: "8_GP_prescriptions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "R Notebook"
output: rmarkdown::github_document
---

# Further data preparation in R 
testing

## Introduction

In this notebook, we will prepare data for analysis.

Extract GP data 


## How to run this notebook

This notebook should be run in a *RStudio* session. It does *not* require a Spark cluster. See how to set it up [here](https://dnanexus.gitbook.io/uk-biobank-rap/working-on-the-research-analysis-platform/using-rstudio-on-the-research-analysis-platform).

## Set up the session

We load packages we'll use:

```{r, include=FALSE}
# First we need to install packages that aren't already present
pkgs <- c("data.table", "plyr", "dplyr","stringr") # packages we need
pkgs_inst <- pkgs[!{pkgs %in% rownames(installed.packages())}] # check which are not present 
install.packages(pkgs_inst, repos = "https://www.stats.bris.ac.uk/R/") # install

# Load packages
lapply(pkgs, library, character.only = TRUE) # using lapply just allows us to load several packages on one line - this could be replaced with several calls to library()
```

dx download - downloads stored permanent storage RAP data into my temporary session
Needs bash to run 
```{bash}
cd ~/rap_wearables

dx download /users/mcgaghd/Data/scripts_wacc_data.csv -r
```

```{r}
dat_scripts <- fread("~/rap_wearables/scripts_wacc_data.csv", data.table = FALSE)
```

###Methotrexate 
```{r}
methotrexate <- dat_scripts %>% 
  filter(grepl("methotrexate", drug_name, ignore.case = TRUE))
```

Metoject
```{r}
metoject <- dat_scripts %>% 
  filter(grepl("Metoject", drug_name, ignore.case = TRUE))
```


Summary total on methotrexate 
```{r}
eid_summary <- methotrexate %>%
  group_by(eid) %>%
  summarise(count = n())
```


###Sulfasalazine
```{r}
sulfasalazine <- dat_scripts %>% 
  filter(grepl("Sulfasalazine|sulphasalazine|salazopyrin", drug_name, ignore.case = TRUE))
```

```{r}
eid_summary <- sulfasalazine %>%
  group_by(eid) %>%
  summarise(count = n())
```
