---
title: "8_GP_prescriptions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "R Notebook"
output: rmarkdown::github_document
---

# Further data preparation in R 
testing

## Introduction

In this notebook, we will prepare data for analysis.

Extract GP data 


## How to run this notebook

This notebook should be run in a *RStudio* session. It does *not* require a Spark cluster. See how to set it up [here](https://dnanexus.gitbook.io/uk-biobank-rap/working-on-the-research-analysis-platform/using-rstudio-on-the-research-analysis-platform).

## Set up the session

We load packages we'll use:

```{r, include=FALSE}
# First we need to install packages that aren't already present
pkgs <- c("data.table", "plyr", "dplyr","stringr") # packages we need
pkgs_inst <- pkgs[!{pkgs %in% rownames(installed.packages())}] # check which are not present 
install.packages(pkgs_inst, repos = "https://www.stats.bris.ac.uk/R/") # install

# Load packages
lapply(pkgs, library, character.only = TRUE) # using lapply just allows us to load several packages on one line - this could be replaced with several calls to library()
```

dx download - downloads stored permanent storage RAP data into my temporary session
Needs bash to run 
```{bash}
cd ~/rap_wearables

dx download /users/mcgaghd/Data/scripts_wacc_data.csv -r
```

```{r}
dat_scripts <- fread("scripts_wacc_data.csv", data.table = FALSE)

```
##for NAFLD analysis - take prepped_steps.csv from end of script 
```{r}
dat <- fread("prepped_steps.csv", data.table=FALSE)

names(dat) <- make.unique(names(dat))

```

```{r}
unique_eids <- dat_GPevent %>% 
  select(eid) %>% 
  distinct()

unique_eids <- unique_eids$eid

```

```{r}
unique_eids %>% 
  mutate(record = TRUE)
```

```{r}
dat_subset <- dat %>%
  filter(eid %in% unique_eids)
```

```{r}
dat_subset <- dat_subset %>%
  mutate(total_prevalent_IA = total_prevalent_AS | total_prevalent_RA|total_prevalent_PsA)

```

##csDMARDs first 

###Methotrexate 
```{r}
methotrexate <- dat_scripts %>% 
  filter(grepl("methotrexate|metoject|Jylamvo|Maxtrex|Methofill|Nordimet|Zlatal|Trexall|Otrexup|Rasuvo|Rheumatrex|Xatmep|reditrex", drug_name, ignore.case = TRUE))
```

```{r}
drug_summary <- methotrexate %>%
  group_by(drug_name) %>%
  summarise(count = n())
```

## Write out
```{r}
write.csv(methotrexate, "methotrexate.csv")
```


In terminal, run the following code: `dx upload methotrexate.csv --dest users/mcgaghd/methotrexate.csv`


Metoject
```{r}
metoject <- dat_scripts %>% 
  filter(grepl("Metoject", drug_name, ignore.case = TRUE))
```


Summary total on methotrexate 
```{r}
eid_summary <- methotrexate %>%
  group_by(eid) %>%
  summarise(count = n())
```


###Sulfasalazine
```{r}
sulfasalazine <- dat_scripts %>% 
  filter(grepl("Sulfasalazine|sulphasalazine|salazopyrin", drug_name, ignore.case = TRUE))
```

```{r}
eid_summary <- sulfasalazine %>%
  group_by(eid) %>%
  summarise(count = n())
```

##Leflunomide
```{r}
leflunomide <- dat_scripts %>% 
  filter(grepl("leflunomide", drug_name, ignore.case = TRUE))
```

###Ciclosporin 
```{r}
ciclosporin <- dat_scripts %>% 
  filter(grepl("Ciclosporin|Capimune|deximune", drug_name, ignore.case = TRUE))
```

###Azathioprine 
```{r}
azathioprine <- dat_scripts %>% 
  filter(grepl("azathioprine|Imuran", drug_name, ignore.case = TRUE))
```

###hydroxychloroquine
```{r}
hydroxychloroquine <- dat_scripts %>% 
  filter(grepl("hydroxychloroquine|Plaquenil", drug_name, ignore.case = TRUE))
```

DMARD
```{r}
DMARD <- rbind(hydroxychloroquine,azathioprine,ciclosporin,leflunomide,sulfasalazine,methotrexate)
```

```{r}
joined_data <- DMARD %>%
  left_join(dat_accel_GP, by = "eid")
```

```{r}
filtered_data <- joined_data %>%
  filter(issue_date <= date_end_accel | is.na(date_end_accel))
```

```{r}
library(lubridate)

enhanced_data <- filtered_data %>%
  mutate(
    date_end_accel = as.Date(date_end_accel), 
    issue_date = as.Date(issue_date),
    DMARD = issue_date >= (date_end_accel %m-% months(60)) & issue_date <= date_end_accel
  )

```

```{r}
final_data <- enhanced_data %>%
  group_by(eid) %>%
  summarise(DMARD = any(DMARD))
```

```{r}
final_data <- final_data %>%
  filter(DMARD == TRUE)
```

```{r}
GP_DMARD <- dat_subset %>%
  left_join(final_data, by = "eid") %>%
  mutate(DMARD = ifelse(is.na(DMARD), FALSE, DMARD))

```

```{r}
GP_DMARD_IA <- GP_DMARD %>%
  subset(total_prevalent_IA ==TRUE)
```

```{r}
model_steps1 <- lm(med_steps ~ DMARD + as.factor(sex)  + as.factor(age_gp) + as.factor(season_wear)+ as.factor(tdi_cats) + as.factor(alcohol) + as.factor(smoking),  data = GP_DMARD_IA)
summary(model_steps1)
  confint(model_steps1)
  
emmeans_model_steps1_np<- emmeans(model_steps1, specs = "DMARD")
emmeans_model_steps1 <- emmeans(model_steps1, specs = pairwise~ DMARD)
summary(emmeans_model_steps1_np)
summary(emmeans_model_steps1)
confint(emmeans_model_steps1)

```


##biologic DMARDs

###anti-TNF agents 
```{r}
antiTNF <- dat_scripts %>% 
  filter(grepl("humira|adalimumab|etanercept|Enbrel|infliximab|golimumab|certolizumab|INFLECTRA|secukinumab|ustekinumab ", drug_name, ignore.case = TRUE))
```

##Corticosteroids

###oral steroids

##https://datacompass.lshtm.ac.uk/id/eprint/2799/1/corticosteroids_aurum_feb21.txt Pulled from LSHTM
```{r}
OralSteroids <- dat_scripts %>% 
  filter(
    grepl("Cortisone|hydrocortisone|prednisolone|Prednisone|Prednesol|Dexamethasone|Budesonide|Betamethasone|Millipred|Dexasone|Orapred|prelone|depopred|decadron|cortisyl|deltastab|triamcinolone|cortelan|deltastab|efcortelan|cortistab|deflazacort|adcortyl|hydrocortone|kenalog|deltacortril|medrone|hydrocortistab|betnesol|lederspan|triamcinolone|lodotra|plenadren|pevanti|dilacort |glensoludex|neofordex|alkindi|hydventia", drug_name, ignore.case = TRUE) &
      !grepl("vial|inj|subcutan|intramusc", drug_name, ignore.case = TRUE) &  #Remove injection
    !grepl("oint|ointment|eye|foam|Enema|suppository|ear|supposi|supp|cream|crm|oin|cre|lotion|topical|spray|Lidocaine|solution|drop|mousse|lot|paste|Calcipotriol|scalp|inhaler|Fludrocortisone|gel|inh|Turbohaler|canister|drp|plaster", drug_name, ignore.case = TRUE) #remove topical
  )
```

```{r}
InjectedSteroids <- dat_scripts %>% 
  filter(
    grepl("Cortisone|hydrocortisone|prednisolone|Prednisone|Prednesol|Dexamethasone|Millipred|Dexasone|prelone|depopred|decadron", drug_name, ignore.case = TRUE) &
      !grepl("tab|tablet|capsule|lozenge", drug_name, ignore.case = TRUE) &  #Remove oral
    !grepl("oint|ointment|eye|foam|Enema|suppository|ear|supposi|supp|cream|crm|oin|cre|lotion|topical|spray|drop|scalp|mousse|lot|inhaler|gel|inh|Turbohaler|canister|drp|plaster", drug_name, ignore.case = TRUE) #remove topical
  ) 
```

SystemicCorticosteroids
```{r}
SystemicCorticosteroids <- rbind(OralSteroids, InjectedSteroids)
```

## Write out
```{r}
write.csv(SystemicCorticosteroids, "systemic_corticosteroids.csv")
```


In terminal, run the following code: `dx upload systemic_corticosteroids.csv --dest users/mcgaghd/systemic_corticosteroids.csv`

```{r}
drug_summary <- InjectedSteroids %>%
  group_by(drug_name) %>%
  summarise(count = n())
```

```{r}
drug_summary <- hydrocortisone %>%
  group_by(drug_name) %>%
  summarise(count = n())
```

```{r}
MildOpiates <- dat_scripts %>% 
  filter(
    grepl("tramadol|codeine|co-codamol|Dextropropoxyphene|Pethidine|Df118|Zydol|Pentazocine|Tramacet|fortral", drug_name, ignore.case = TRUE) &
    !grepl("oint|ointment|injection|eye|foam|Enema|suppository|ear|supposi|supp|cream|crm|oin|cre|lotion|topical|spray", drug_name, ignore.case = TRUE) 
  )
```

## Write out
```{r}
write.csv(MildOpiates, "MildOpiates.csv")
```


In terminal, run the following code: `dx upload MildOpiates.csv --dest users/mcgaghd/MildOpiates.csv`

```{r}
ModerateOpiates <- dat_scripts %>% 
  filter(
    grepl("Morphine|MST contin|Oramorph|Sevredol|Oxycodone|OxyContin|Buprenorphine|BuTrans|Pethidine|Tapentadol|Meptazinol|Meptazinol|Meptid|Transtec|DHC Continus|oxynorm", drug_name, ignore.case = TRUE) &
    !grepl("oint|ointment|eye|foam|Enema|ear|supp|cream|crm|oin|cre|lotion|topical|spray", drug_name, ignore.case = TRUE) 
  )
```

## Write out
```{r}
write.csv(ModerateOpiates, "ModerateOpiates.csv")
```


In terminal, run the following code: `dx upload ModerateOpiates.csv --dest users/mcgaghd/ModerateOpiates.csv`

```{r}
HighOpiates <- dat_scripts %>% 
  filter(
    grepl("Fentanyl|Durogesic|Hydromorphone|Palfium|Dextromoramide|Cyclimorph|Palladone|MXL Capsules|Papaveretum", drug_name, ignore.case = TRUE) &
    !grepl("oint|ointment|eye|foam|Enema|ear|supp|cream|crm|oin|cre|lotion|topical|spray", drug_name, ignore.case = TRUE) 
  )
```

## Write out
```{r}
write.csv(HighOpiates, "HighOpiates.csv")
```


In terminal, run the following code: `dx upload HighOpiates.csv --dest users/mcgaghd/HighOpiates.csv`

```{r}
AnyOpioid <- rbind(MildOpiates,ModerateOpiates,HighOpiates)
```

##NSAIDs

```{r}
NSAIDs <- dat_scripts %>% 
  filter(
    grepl("Ibuprofen|Diclofenac|nurofen|Naproxen|Piroxicam|Indometacin|Mefenamic Acid|Acemetacin|Ketoprofen|Celecoxib|Aceclofenac|Etoricoxib|Meloxicam|Flurbiprofen|Tiaprofenic Acid|Sulindac|Tenoxicam|Nabumetone|Etodolac|Dexketoprofen|Fenoprofen|Diclofenac Potassium|Dexibuprofen", drug_name, ignore.case = TRUE) &
    !grepl("oint|ointment|eye|foam|Enema|ear|supp|cream|crm|oin|cre|lotion|topical|gel|jel|spray", drug_name, ignore.case = TRUE) 
  )
```

## Write out
```{r}
write.csv(NSAIDs, "NSAIDs.csv")
```

In terminal, run the following code: `dx upload NSAIDs.csv --dest users/mcgaghd/NSAIDs.csv`

```{r}
Demerol <- dat_scripts %>% 
  filter(
    grepl("Demerol", drug_name, ignore.case = TRUE) &
    !grepl("oint|injection|ointment|eye|foam|Enema|suppository|ear|supposi|supp|cream|crm|oin|cre|lotion|topical|spray", drug_name, ignore.case = TRUE) 
  )
```
 
```{r}
drug_summary <- NSAIDs %>%
  group_by(drug_name) %>%
  summarise(count = n())
```

Link to all those with prescribing data 
```{r}
unique_eids <- dat_scripts %>% 
  select(eid) %>% 
  distinct()
```

```{r}
unique_eids %>% 
  mutate(script = TRUE)
```


#Subset based on eid and date
```{r}
dat_accel_GP <- dat_subset %>%
  select(eid, date_end_accel)
```

Join
```{r}
joined_data <- SystemicCorticosteroids %>%
  left_join(dat_accel_GP, by = "eid")
```

```{r}
filtered_data <- joined_data %>%
  filter(issue_date <= date_end_accel | is.na(date_end_accel))
```

```{r}
library(lubridate)

enhanced_data <- filtered_data %>%
  mutate(
    date_end_accel = as.Date(date_end_accel), 
    issue_date = as.Date(issue_date),
    SystemicCorticosteroids = issue_date >= (date_end_accel %m-% months(18)) & issue_date <= date_end_accel
  )

```

```{r}
final_data <- enhanced_data %>%
  group_by(eid) %>%
  summarise(SystemicCorticosteroids = any(SystemicCorticosteroids))
```

```{r}
final_data <- final_data %>%
  filter(SystemicCorticosteroids == TRUE)
```

GP data full with oral steroids 
```{r}
GP_steroids <- dat_subset %>%
  left_join(final_data, by = "eid") %>%
  mutate(SystemicCorticosteroids = ifelse(is.na(SystemicCorticosteroids), FALSE, SystemicCorticosteroids))

```

PsA steroids
```{r}
GP_steroids_PsA <- GP_steroids %>%
  filter(total_prevalent_PsA == TRUE)
```

```{r}
statistics <- GP_steroids_PsA %>%
  group_by(OralSteroids) %>%
  summarise(
    mean_med_steps = mean(med_steps, na.rm = TRUE),
    sd_med_steps = sd(med_steps, na.rm = TRUE)
  )
```

```{r}

GP_steroids_PsA <- GP_steroids_PsA %>%
  mutate(sedentary = med_steps < 5000)

```

Join opioids
```{r}
dat_opioid <- AnyOpioid %>%
  left_join(dat_accel_GP, by = "eid")
```


```{r}
dat_opioid <- dat_opioid %>%
  filter(issue_date <= date_end_accel | is.na(date_end_accel))
```

Persistent Opioid within 2 years - 
```{r}
library(purrr)

check_persistent_opioid <- function(dates, end_date) {
  sum(dates >= (end_date %m-% months(24)) & dates <= end_date) >= 3
}

dat_opioid <- dat_opioid %>%
  group_by(eid) %>%
  mutate(PersistentOpioid = check_persistent_opioid(issue_date, date_end_accel)) %>%
  ungroup()

```

```{r}
check_long_term_opioid <- function(dates, end_date) {
  # Standard: 3 or more prescriptions within a 90 day period in the first year
  standard <- any(rollapply(dates, width = 3, FUN = function(x) diff(range(x)) <= 120, align = "right", fill = FALSE))

  # Stringent: 10 or more prescriptions over more than 90 days in the first year
  stringent <- length(dates) >= 10 && diff(range(dates)) > 90

  # Broad: More than 3 prescriptions at monthly intervals in the first 12 months
  broad <- sum(dates >= (end_date %m-% months(24)) & dates <= end_date) > 3

  # Combine the conditions
  standard | stringent | broad
}


```

```{r}
library(zoo)
check_long_term_opioid <- function(dates, end_date) {
  # Filter dates within 3 years from end_date
  dates_within_3_years <- dates[dates >= (end_date %m-% months(36)) & dates <= end_date]

  # Standard: 3 or more prescriptions within a 90 day period in the first year
  standard <- length(dates_within_3_years) >= 3 && any(rollapply(dates_within_3_years, width = 3, FUN = function(x) diff(range(x)) <= 90, align = "right", fill = NA))

  # Stringent: 10 or more prescriptions over more than 90 days in the first year
  stringent <- length(dates_within_3_years) >= 10 && diff(range(dates_within_3_years)) > 90

  # Broad: More than 3 prescriptions at monthly intervals in the first 12 months
  broad_dates <- dates[dates >= (end_date %m-% months(12)) & dates <= end_date]
  broad <- length(broad_dates) > 3

  # Combine the conditions
  standard | stringent | broad
}


# Applying the function to each group (eid) in dat_opioid
dat_opioid <- dat_opioid %>%
  group_by(eid) %>%
  mutate(LongTermOpioid = check_long_term_opioid(issue_date, date_end_accel)) %>%
  ungroup()

# Now dat_opioi
```

```{r}
dat_opioid <- dat_opioid %>%
  group_by(eid) %>%
  summarise(LongTermOpioid = any(LongTermOpioid))
```

```{r}
dat_opioid <- dat_opioid %>%
  filter(LongTermOpioid == TRUE)
```

```{r}
GP_opioids <- dat_subset %>%
  left_join(dat_opioid, by = "eid") %>%
  mutate(LongTermOpioid = ifelse(is.na(LongTermOpioid), FALSE, LongTermOpioid))
```



```{r}
dat_opioid <- dat_opioid %>%
  group_by(eid) %>%
  summarise(PersistentOpioid = any(PersistentOpioid))
```

```{r}
dat_opioid <- dat_opioid %>%
  filter(PersistentOpioid == TRUE)
```

```{r}
GP_opioids <- subset_dat %>%
  left_join(dat_opioid, by = "eid") %>%
  mutate(PersistentOpioid = ifelse(is.na(PersistentOpioid), FALSE, PersistentOpioid))
```

```{r}
GP_opioids_PsA <- GP_opioids %>%
  filter(total_prevalent_PsA == TRUE)
```


```{r}
statistics <- GP_opioids_PsA %>%
  group_by(LongTermOpioid) %>%
  summarise(
    mean_med_steps = mean(med_steps, na.rm = TRUE),
    sd_med_steps = sd(med_steps, na.rm = TRUE)
  )
```

```{r}
model_steps1 <- lm(med_steps ~ SystemicCorticosteroids + as.factor(sex)  + as.factor(age_gp) + as.factor(season_wear)+ as.factor(tdi_cats) + as.factor(alcohol) + as.factor(smoking),  data = GP_steroids_PsA)
summary(model_steps1)
  confint(model_steps1)
  
emmeans_model_steps1_np<- emmeans(model_steps1, specs = "SystemicCorticosteroids")
emmeans_model_steps1 <- emmeans(model_steps1, specs = pairwise~ SystemicCorticosteroids)
summary(emmeans_model_steps1_np)
summary(emmeans_model_steps1)
confint(emmeans_model_steps1)

```

```{r}
GP_steroids_RA <- GP_steroids %>%
  filter(total_prevalent_RA == TRUE | total_prevalent_PsA == TRUE | total_prevalent_AS == TRUE)
```

```{r}
GP_opioids_IA <- GP_opioids %>%
  subset(total_prevalent_IA == TRUE)
```

```{r}
model_steps1 <- lm(med_steps ~ LongTermOpioid + as.factor(sex)  + as.factor(age_gp) + as.factor(season_wear)+ as.factor(tdi_cats) + as.factor(alcohol) + as.factor(smoking),  data = GP_opioids_IA)
summary(model_steps1)
  confint(model_steps1)
  
emmeans_model_steps1_np<- emmeans(model_steps1, specs = "LongTermOpioid")
emmeans_model_steps1 <- emmeans(model_steps1, specs = pairwise~ LongTermOpioid)
summary(emmeans_model_steps1_np)
summary(emmeans_model_steps1)
confint(emmeans_model_steps1)
```

```{r}
GP_opioids_RA <- GP_opioids %>%
  filter(total_prevalent_RA == TRUE | total_prevalent_PsA == TRUE | total_prevalent_AS == TRUE)
```

```{r}
statistics <- GP_opioids_RA %>%
  group_by(PersistentOpioid) %>%
  summarise(
    mean_med_steps = mean(med_steps, na.rm = TRUE),
    sd_med_steps = sd(med_steps, na.rm = TRUE)
  )
```

```{r}
model_steps1 <- lm(med_steps ~ LongTermOpioid + as.factor(sex)  + as.factor(age_gp) + as.factor(season_wear)+ as.factor(tdi_cats) + as.factor(alcohol) + as.factor(smoking),  data = GP_opioids_PsA)
summary(model_steps1)
  confint(model_steps1)
  
emmeans_model_steps1_np<- emmeans(model_steps1, specs = "LongTermOpioid")
emmeans_model_steps1 <- emmeans(model_steps1, specs = pairwise~ LongTermOpioid)
summary(emmeans_model_steps1_np)
summary(emmeans_model_steps1)
confint(emmeans_model_steps1)

```





