---
title: "1sdPlots"
author: "Dylan McGagh"
date: "`r format(Sys.time(), '%a %b %d, %Y %X')`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    toc_depth: 5
    theme: lumen
---
```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This script runs models for 1 SD increase in exposure.

# Set up workspace

## Set-up useful functions and packages
```{r}
#Load packages
pkgs <- c("data.table", "rmarkdown", "plyr", "lubridate", "magrittr", "ggplot2", "data.table", "dplyr", "stringr", "splitstackshape", "scales", "haven", "Epi", "emmeans", "rtables", "table1",  "tableone", "boot", "knitr", "pspline", "survival", "xtable", "here", "forcats", "patchwork") # packages we need

pkgs_inst <- pkgs[!{pkgs %in% rownames(installed.packages())}] # check which are not present 
install.packages(pkgs_inst, repos = "https://www.stats.bris.ac.uk/R/") # install
lapply(pkgs, library, character.only = TRUE) #load all packages
```

## Load data
```{r, warning=FALSE, error=FALSE, message=FALSE, results = 'hide'}
## Outcome dfs
#updated file path 2023 10 03
steps_datLow <- fread("PRSbyGroup/steps_sdmodelresultsLow.csv", data.table = FALSE)
steps_datInter <- fread("PRSbyGroup/steps_sdmodelresultsInter.csv", data.table = FALSE)
steps_datHigh <- fread("PRSbyGroup/steps_sdmodelresultsHigh.csv",data.table = FALSE)

```

```{r}
# Load the forestplot package (if not already loaded)
library(forestplot)
```


```{r}
# Load the forestplot package (if not already loaded)

# Create a list to store forest plots for each model
forest_plots <- list()

# List of models to iterate over
models <- unique(steps_datLow$Model)

# Create a separate forest plot for each model
for (model in models) {
  model_data <- subset(steps_datLow, Model == model)

  # Split HR_CI into HR and CI columns
  model_data$HR <- gsub(".*\\(", "", model_data$HR_CI)
  model_data$HR <- as.numeric(gsub("\\).*", "", model_data$HR))
  model_data$CI <- gsub(".*\\(", "", model_data$HR_CI)
  model_data$CI <- gsub(" - .*", "", model_data$CI)
  model_data$CI <- as.numeric(model_data$CI)

  # Create the forest plot for the current model
  forest_plots[[model]] <- forestplot(
    labeltext = model_data$Outcome,
    mean = model_data$HR,
    lower = model_data$HR - 0.5 * model_data$CI,
    upper = model_data$HR + 0.5 * model_data$CI,
    title = paste("Cox Proportional Hazards Model Results -", model),
    xlab = "Hazard Ratio (HR)",
    is.summary = c(TRUE, FALSE),
    col = fpColors(box = "blue", line = "blue"),
    hrzl_lines = list(
      "Model" = gpar(fontsize = 12),
      "P.trend" = gpar(fontsize = 12),
      "AIC" = gpar(fontsize = 12)
    ),
    new_page = TRUE
  )
}

# Plot the forest plots for each model
for (i in seq_along(models)) {
  print(forest_plots[[i]])
}


```

```{r}

# Load the forestplot package (if not already loaded)
library(forestplot)

# Create a list to store forest plots for each model
forest_plots <- list()

# List of models to iterate over
models <- unique(steps_datLow$Model)

# Create a separate forest plot for each model
for (model in models) {
  model_data <- subset(steps_datLow, Model == model)

  # Extract HR and CI, handling different formats
  model_data$HR <- as.numeric(gsub(".*?([0-9.]+)\\s*\\(.*", "\\1", model_data$HR_CI))
  model_data$CI <- as.numeric(gsub(".*?\\(([0-9.]+)\\s*-\\s*([0-9.]+)\\).*", "\\2", model_data$HR_CI))

  # Handle cases where extraction produces NA
  model_data$HR[is.na(model_data$HR)] <- 1.0
  model_data$CI[is.na(model_data$CI)] <- 0.0

  # Create the forest plot for the current model
  forest_plots[[model]] <- forestplot(
    labeltext = model_data$Outcome,
    mean = model_data$HR,
    lower = model_data$HR - 0.5 * model_data$CI,
    upper = model_data$HR + 0.5 * model_data$CI,
    title = paste("Cox Proportional Hazards Model Results -", model),
    xlab = "Hazard Ratio (HR)",
    is.summary = c(TRUE, FALSE),
    col = fpColors(box = "blue", line = "blue"),
    hrzl_lines = list(
      "Model" = gpar(fontsize = 12),
      "P.trend" = gpar(fontsize = 12),
      "AIC" = gpar(fontsize = 12)
    ),
    new_page = TRUE
  )
}

# Plot the forest plots for each model
for (i in seq_along(models)) {
  print(forest_plots[[i]])
}

```

#### Steps - male - prep to save
Combine results from sex specific models and main model
```{r, warning=FALSE, error=FALSE, message=FALSE, results = 'hide'}
## Prep steps dataset for modeling
steps_Low <- steps_datLow

# Drop rows where HR..95..CI. = "REF"
steps_Low <- steps_Low %>%
  filter(!str_detect(`HR..95..CI.`, "REF"))

# Rename column
steps_Low <- steps_Low %>%
  rename(estimate = `HR..95..CI.`)

# Extract the lower and upper bounds of the estimates from the "estimate" column
steps_Low$lower <- as.numeric(gsub(".*\\((.*) -.*", "\\1", steps_Low$estimate))
steps_Low$upper <- as.numeric(gsub(".*- (.*)\\)", "\\1", steps_Low$estimate))
steps_Low$hr <- as.numeric(gsub(" \\(.*", "", steps_Low$estimate))


# rename
steps_Low <-rename(steps_Low, outcome = Outcome)

```


```{r, warning=FALSE, error=FALSE, message=FALSE, results = 'hide'}
########################Clean for export
dat <- steps_Low
#combine estimate and confidence intervals, keeping 2 digits after decimal including 0s
dat$hr_cis <- sprintf("%.2f (%.2f-%.2f)", dat$hr, dat$lower, dat$upper)
######## minimally adjusted
dat_minimal <- subset(dat, Model == "Unadjusted model")
dat_minimal <- dat_minimal %>%
  rename(hr_cis_minimal = hr_cis)
dat_minimal <- dat_minimal %>%
  select(outcome, hr_cis_minimal)

######## Fully adjusted
dat_full <- subset(dat, Model == "Fully-adjusted")
dat_full <- dat_full %>%
  rename(hr_cis_full = hr_cis)
dat_full <- dat_full %>%
  select(outcome, hr_cis_full)


######## Combine and save
# Merge dat_minimal, dat_full, and dat_full_bmi by outcomes_short
combined_data <- merge(dat_minimal, dat_full, by = "outcome", all = TRUE)

# Reorder columns with outcome_type as the first column
combined_data <- combined_data[, c("outcome", setdiff(names(combined_data), "outcome"))]
#sort alphabetically first by outcometype and then by outcomes_short
#sort alphabetically first by outcometype and then by outcomes_short

```

```

##### Combine and save
```{r, warning=FALSE, error=FALSE, message=FALSE, results = 'hide'}
########################Clean for export
dat <- steps_male
#combine estimate and confidence intervals, keeping 2 digits after decimal including 0s
dat$hr_cis <- sprintf("%.2f (%.2f-%.2f)", dat$hr, dat$lower, dat$upper)
######## minimally adjusted
dat_minimal <- subset(dat, Model == "Unadjusted model")
dat_minimal <- dat_minimal %>%
  rename(hr_cis_minimal = hr_cis)
dat_minimal <- dat_minimal %>%
  select(outcometype, outcome, outcomes_short, casecountmale, hr_cis_minimal)

######## Fully adjusted
dat_full <- subset(dat, Model == "Fully-adjusted")
dat_full <- dat_full %>%
  rename(hr_cis_full = hr_cis)
dat_full <- dat_full %>%
  select(outcomes_short, hr_cis_full)

######## Fully adjusted + bmi
dat_full_bmi <- subset(dat, Model == "Fully adjusted + bmi")
dat_full_bmi <- dat_full_bmi %>%
  rename(hr_cis_full_bmi = hr_cis)
dat_full_bmi <- dat_full_bmi %>%
  select(outcomes_short, hr_cis_full_bmi)

######## Combine and save
# Merge dat_minimal, dat_full, and dat_full_bmi by outcomes_short
combined_data <- merge(dat_minimal, dat_full, by = "outcomes_short", all = TRUE)
combined_data <- merge(combined_data, dat_full_bmi, by = "outcomes_short", all = TRUE)

# Reorder columns with outcome_type as the first column
combined_data <- combined_data[, c("outcometype", setdiff(names(combined_data), "outcometype"))]
#sort alphabetically first by outcometype and then by outcomes_short
#sort alphabetically first by outcometype and then by outcomes_short
combined_data <- combined_data %>%
  arrange(outcometype, outcomes_short)
write.csv(combined_data, file=here::here("data", "results", paste0("step_sd_modelresults_clean_male.csv")), row.names=FALSE)
```



 