---
title: "d_analyse_data"
format: html
editor: visual
---

## Load data

Data:

```{bash}

dx download  /users/mcgaghd/Data/prepped_steps.csv
dx download /users/mcgaghd/PrimaryCare -r

```

```{bash}
dx download /users/mcgaghd/PrimaryCare -r

```

```{r}
dat <- fread("prepped_steps.csv", data.table = FALSE)

dat$V1 <- NULL
# this file location assumes running in same session as previous script
#dat <- fread("Data/prepped_steps_RA2years.csv", data.table = FALSE)

```

Symptoms indicator
```{r}
datArthralgia_gp <-fread("PrimaryCare/datArthralgia_gp.csv", data.table = FALSE) # fread is a function from the data.table package for fast reading of large data
datBackPain_gp <- fread("PrimaryCare/datBackPain_gp.csv", data.table = FALSE)
datBursitis_gp <- fread("PrimaryCare/datBursitis_gp.csv", data.table = FALSE)
datFatigue_gp <- fread("PrimaryCare/datFatigue_gp.csv", data.table = FALSE)
datFingerPain_gp <-fread("PrimaryCare/datFingerPain_gp.csv", data.table = FALSE) # fread is a function from the data.table package for fast reading of large data
datHipPain_gp <- fread("PrimaryCare/datHipPain_gp.csv", data.table = FALSE)
datKneePain_gp <- fread("PrimaryCare/datKneePain_gp.csv", data.table = FALSE)
datMyalgia_gp <- fread("PrimaryCare/datMyalgia_gp.csv", data.table = FALSE)
datStiffneess_gp <- fread("PrimaryCare/datStiffness_gp.csv", data.table = FALSE)
datSwelling_gp <- fread("PrimaryCare/datSwelling_gp.csv", data.table = FALSE)
datTendonitis_gp <- fread("PrimaryCare/datTendonitis_gp.csv", data.table = FALSE)
datWristHandPain_gp <- fread("PrimaryCare/datWristHandPain_gp.csv", data.table = FALSE)
```

```{r}
datArthralgia_gp$event_dt <- as.Date(datArthralgia_gp$event_dt, format = "%Y-%m-%d")
datBackPain_gp$event_dt <- as.Date(datBackPain_gp$event_dt, format = "%Y-%m-%d")
datBursitis_gp$event_dt <-  as.Date(datBursitis_gp$event_dt, format = "%Y-%m-%d")
datFatigue_gp$event_dt <-  as.Date(datFatigue_gp$event_dt, format = "%Y-%m-%d")
datHipPain_gp$event_dt <-  as.Date(datHipPain_gp$event_dt, format = "%Y-%m-%d")
datKneePain_gp$event_dt <- as.Date(datKneePain_gp$event_dt, format = "%Y-%m-%d")
datFingerPain_gp$event_dt <-  as.Date(datFingerPain_gp$event_dt, format = "%Y-%m-%d")
datMyalgia_gp$event_dt <-  as.Date(datMyalgia_gp$event_dt, format = "%Y-%m-%d")
datStiffneess_gp$event_dt <-  as.Date(datStiffneess_gp$event_dt, format = "%Y-%m-%d")
datSwelling_gp$event_dt <-  as.Date(datSwelling_gp$event_dt, format = "%Y-%m-%d")
#datEnthesitis_gp$event_dt <-  as.Date(datEnthesitis_gp$event_dt, format = "%Y-%m-%d")

```

```{bash}
dx download /users/mcgaghd/Data/participant_data_pain.csv -r

```

```{r}
dat_pain <- fread("participant_data_pain.csv", data.table = FALSE)

```

Create a chronic MSK pain indicator 
```{r}

dat_pain <- dat_pain %>%
  mutate(PainMissing = grepl("Prefer not to answer", pain_last_month, ignore.case = TRUE))

```

Pain last month
```{r}
dat_pain <- dat_pain %>%
  mutate(JointPain_1month = grepl("neck or shoulder|back pain|hip pain|knee pain", pain_last_month, ignore.case = TRUE))

#Pain all over body 
dat_pain <- dat_pain %>%
  mutate(PainWidespread_1month = grepl("all over the body", pain_last_month, ignore.case = TRUE))
```


3 month pain 
```{r}
dat_pain <- dat_pain %>%
  mutate(
    pain_3month_Back = ifelse(PainMissing, "Prefer not to answer", ifelse(pain_3month_Back == "", "No", pain_3month_Back)),
    pain_3month_Hip = ifelse(PainMissing, "Prefer not to answer", ifelse(pain_3month_Hip == "", "No", pain_3month_Hip)),
    pain_3month_General = ifelse(PainMissing, "Prefer not to answer", ifelse(pain_3month_General == "", "No", pain_3month_General)),
    pain_3month_Knee = ifelse(PainMissing, "Prefer not to answer", ifelse(pain_3month_Knee == "", "No", pain_3month_Knee)),
    pain_3month_NeckShoulder = ifelse(PainMissing, "Prefer not to answer", ifelse(pain_3month_NeckShoulder == "", "No", pain_3month_NeckShoulder))
  )
```


```{r}
dat_pain <- dat_pain %>%
  mutate(ChronicMSKpain = ifelse(pain_3month_Back == "Yes" | pain_3month_Hip == "Yes" | pain_3month_Knee == "Yes" | pain_3month_NeckShoulder == "Yes", TRUE, FALSE)
  )
```

Drop main columns
```{r}
dat_pain_subset <- dat_pain %>%
  select(eid, ChronicMSKpain, pain_3month_Back, pain_3month_Hip, pain_3month_Knee, pain_3month_NeckShoulder, pain_3month_General, pain_last_month)

# Join the subset onto dat by eid
dat <- dat %>%
  left_join(dat_pain_subset, by = "eid")

```

Bring in the primary care prodromal indicators

Prodromal PsA

Subset dat to keep date_baseline and eid 
```{r}
dat_eid <- dat %>% dplyr::select(eid, date_end_accel)

```

Arthralgia
```{r}
datArthralgia_gp$event_dt <- as.Date(datArthralgia_gp$event_dt)
dat_eid$date_baseline <- as.Date(dat_eid$date_baseline)



# Merge dat_subset with datArthralgia_gp
merged_data <- merge(dat_eid, datArthralgia_gp, by = "eid", all.x = TRUE)

# Create the arthralgia_6months, arthralgia_1year, and arthralgia_5years columns
merged_data <- merged_data %>%
  mutate(
    arthralgia_6months = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(6)), TRUE, FALSE),
    arthralgia_1year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(12)), TRUE, FALSE),
    arthralgia_2year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(24)), TRUE, FALSE),
    arthralgia_5years = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(60)), TRUE, FALSE)
  )

# Summarize to get TRUE/FALSE for each eid
dat_subset <- merged_data %>%
  group_by(eid) %>%
  summarize(
    arthralgia_6months = any(arthralgia_6months, na.rm = TRUE),
    arthralgia_1year = any(arthralgia_1year, na.rm = TRUE),
    arthralgia_2year = any(arthralgia_2year, na.rm = TRUE),
    arthralgia_5years = any(arthralgia_5years, na.rm = TRUE)
  ) %>%
  ungroup()

# Display the resul

dat <- dat %>%
  left_join(dat_subset, by = "eid")
```

Myalgia
```{r}
datMyalgia_gp$event_dt <- as.Date(datMyalgia_gp$event_dt)
dat_eid$date_baseline <- as.Date(dat_eid$date_baseline)



# Merge dat_subset with datArthralgia_gp
merged_data <- merge(dat_eid, datMyalgia_gp, by = "eid", all.x = TRUE)

# Create the arthralgia_6months, arthralgia_1year, and arthralgia_5years columns
merged_data <- merged_data %>%
  mutate(
    myalgia_6months = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(6)), TRUE, FALSE),
    myalgia_1year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(12)), TRUE, FALSE),
    myalgia_2year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(24)), TRUE, FALSE),
    myalgia_5years = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(60)), TRUE, FALSE)
  )

# Summarize to get TRUE/FALSE for each eid
dat_subset <- merged_data %>%
  group_by(eid) %>%
  summarize(
    myalgia_6months = any(myalgia_6months, na.rm = TRUE),
    myalgia_1year = any(myalgia_1year, na.rm = TRUE),
    myalgia_2year = any(myalgia_2year, na.rm = TRUE),
    myalgia_5years = any(myalgia_5years, na.rm = TRUE)
  ) %>%
  ungroup()

# Display the resul

dat <- dat %>%
  left_join(dat_subset, by = "eid")
```


Fatigue
```{r}
datFatigue_gp$event_dt <- as.Date(datFatigue_gp$event_dt)
dat_eid$date_baseline <- as.Date(dat_eid$date_baseline)



# Merge dat_subset with datArthralgia_gp
merged_data <- merge(dat_eid, datFatigue_gp, by = "eid", all.x = TRUE)

# Create the arthralgia_6months, arthralgia_1year, and arthralgia_5years columns
merged_data <- merged_data %>%
  mutate(
    fatigue_6months = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(6)), TRUE, FALSE),
    fatigue_1year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(12)), TRUE, FALSE),
        fatigue_2year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(24)), TRUE, FALSE),
    fatigue_5years = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(60)), TRUE, FALSE)
  )

# Summarize to get TRUE/FALSE for each eid
dat_subset <- merged_data %>%
  group_by(eid) %>%
  summarize(
    fatigue_6months = any(fatigue_6months, na.rm = TRUE),
    fatigue_1year = any(fatigue_1year, na.rm = TRUE),
    fatigue_2year = any(fatigue_2year, na.rm=TRUE),
    fatigue_5years = any(fatigue_5years, na.rm = TRUE)
  ) %>%
  ungroup()

# Display the resul

dat <- dat %>%
  left_join(dat_subset, by = "eid")
```

Swelling
```{r}
datSwelling_gp$event_dt <- as.Date(datSwelling_gp$event_dt)
dat_eid$date_baseline <- as.Date(dat_eid$date_baseline)



# Merge dat_subset with datArthralgia_gp
merged_data <- merge(dat_eid, datSwelling_gp, by = "eid", all.x = TRUE)

# Create the arthralgia_6months, arthralgia_1year, and arthralgia_5years columns
merged_data <- merged_data %>%
  mutate(
    swelling_6months = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(6)), TRUE, FALSE),
    swelling_1year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(12)), TRUE, FALSE),
    swelling_2year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(24)), TRUE, FALSE),
    swelling_5years = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(60)), TRUE, FALSE)
  )

# Summarize to get TRUE/FALSE for each eid
dat_subset <- merged_data %>%
  group_by(eid) %>%
  summarize(
    swelling_6months = any(swelling_6months, na.rm = TRUE),
    swelling_1year = any(swelling_1year, na.rm = TRUE),
    swelling_2year = any(swelling_2year, na.rm = TRUE),
    swelling_5years = any(swelling_5years, na.rm = TRUE)
  ) %>%
  ungroup()

# Display the resul

dat <- dat %>%
  left_join(dat_subset, by = "eid")
```

Back pain 
```{r}
datBackPain_gp$event_dt <- as.Date(datBackPain_gp$event_dt)
dat_eid$date_baseline <- as.Date(dat_eid$date_baseline)


# Merge dat_subset with datArthralgia_gp
merged_data <- merge(dat_eid, datBackPain_gp, by = "eid", all.x = TRUE)

# Create the arthralgia_6months, arthralgia_1year, and arthralgia_5years columns
merged_data <- merged_data %>%
  mutate(
    BackPain_6months = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(6)), TRUE, FALSE),
    BackPain_1year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(12)), TRUE, FALSE),
    BackPain_2year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(24)), TRUE, FALSE),
    BackPain_5years = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(60)), TRUE, FALSE)
  )

# Summarize to get TRUE/FALSE for each eid
dat_subset <- merged_data %>%
  group_by(eid) %>%
  summarize(
    BackPain_6months = any(BackPain_6months, na.rm = TRUE),
    BackPain_1year = any(BackPain_1year, na.rm = TRUE),
    BackPain_2year = any(BackPain_2year, na.rm = TRUE),
    BackPain_5years = any(BackPain_5years, na.rm = TRUE)
  ) %>%
  ungroup()

# Display the resul

dat <- dat %>%
  left_join(dat_subset, by = "eid")


```


Knee pain 
```{r}
datKneePain_gp$event_dt <- as.Date(datKneePain_gp$event_dt)
dat_eid$date_baseline <- as.Date(dat_eid$date_baseline)


# Merge dat_subset with datArthralgia_gp
merged_data <- merge(dat_eid, datKneePain_gp, by = "eid", all.x = TRUE)

# Create the arthralgia_6months, arthralgia_1year, and arthralgia_5years columns
merged_data <- merged_data %>%
  mutate(
    KneePain_6months = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(6)), TRUE, FALSE),
    KneePain_1year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(12)), TRUE, FALSE),
    KneePain_2year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(24)), TRUE, FALSE),
    KneePain_5years = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(60)), TRUE, FALSE)
  )

# Summarize to get TRUE/FALSE for each eid
dat_subset <- merged_data %>%
  group_by(eid) %>%
  summarize(
    KneePain_6months = any(KneePain_6months, na.rm = TRUE),
    KneePain_1year = any(KneePain_1year, na.rm = TRUE),
    KneePain_2year = any(KneePain_2year, na.rm = TRUE),
    KneePain_5years = any(KneePain_5years, na.rm = TRUE)
  ) %>%
  ungroup()

# Display the resul

dat <- dat %>%
  left_join(dat_subset, by = "eid")


```


```{r}
datHipPain_gp$event_dt <- as.Date(datHipPain_gp$event_dt)
dat_eid$date_baseline <- as.Date(dat_eid$date_baseline)


# Merge dat_subset with datArthralgia_gp
merged_data <- merge(dat_eid, datHipPain_gp, by = "eid", all.x = TRUE)

# Create the arthralgia_6months, arthralgia_1year, and arthralgia_5years columns
merged_data <- merged_data %>%
  mutate(
    HipPain_6months = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(6)), TRUE, FALSE),
    HipPain_1year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(12)), TRUE, FALSE),
        HipPain_2year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(24)), TRUE, FALSE),
    HipPain_5years = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(60)), TRUE, FALSE)
  )

# Summarize to get TRUE/FALSE for each eid
dat_subset <- merged_data %>%
  group_by(eid) %>%
  summarize(
    HipPain_6months = any(HipPain_6months, na.rm = TRUE),
    HipPain_1year = any(HipPain_1year, na.rm = TRUE),
        HipPain_2year = any(HipPain_2year, na.rm = TRUE),
    HipPain_5years = any(HipPain_5years, na.rm = TRUE)
  ) %>%
  ungroup()

# Display the resul

dat <- dat %>%
  left_join(dat_subset, by = "eid")

```

Fingerpain
```{r}
datFingerPain_gp$event_dt <- as.Date(datFingerPain_gp$event_dt)
dat_eid$date_baseline <- as.Date(dat_eid$date_baseline)


# Merge dat_subset with datArthralgia_gp
merged_data <- merge(dat_eid, datFingerPain_gp, by = "eid", all.x = TRUE)

# Create the arthralgia_6months, arthralgia_1year, and arthralgia_5years columns
merged_data <- merged_data %>%
  mutate(
    FingerPain_6months = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(6)), TRUE, FALSE),
    FingerPain_1year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(12)), TRUE, FALSE),
    FingerPain_2year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(24)), TRUE, FALSE),
    FingerPain_5years = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(60)), TRUE, FALSE)
  )

# Summarize to get TRUE/FALSE for each eid
dat_subset <- merged_data %>%
  group_by(eid) %>%
  summarize(
    FingerPain_6months = any(FingerPain_6months, na.rm = TRUE),
    FingerPain_1year = any(FingerPain_1year, na.rm = TRUE),
    FingerPain_2year = any(FingerPain_2year, na.rm = TRUE),
    FingerPain_5years = any(FingerPain_5years, na.rm = TRUE)
  ) %>%
  ungroup()

# Display the resul

dat <- dat %>%
  left_join(dat_subset, by = "eid")

```

Bursitis
```{r}
datBursitis_gp$event_dt <- as.Date(datBursitis_gp$event_dt)
dat_eid$date_baseline <- as.Date(dat_eid$date_baseline)


# Merge dat_subset with datArthralgia_gp
merged_data <- merge(dat_eid, datBursitis_gp, by = "eid", all.x = TRUE)

# Create the arthralgia_6months, arthralgia_1year, and arthralgia_5years columns
merged_data <- merged_data %>%
  mutate(
    Bursitis_6months = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(6)), TRUE, FALSE),
    Bursitis_1year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(12)), TRUE, FALSE),
    Bursitis_2year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(24)), TRUE, FALSE),
    Bursitis_5years = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(60)), TRUE, FALSE)
  )

  # Summarize to get TRUE/FALSE for each eid
dat_subset <- merged_data %>%
  group_by(eid) %>%
  summarize(
    Bursitis_6months = any(Bursitis_6months, na.rm = TRUE),
    Bursitis_1year = any(Bursitis_1year, na.rm = TRUE),
    Bursitis_2year = any(Bursitis_2year, na.rm = TRUE),
    Bursitis_5years = any(Bursitis_5years, na.rm = TRUE)
  ) %>%
  ungroup()

# Display the resul

dat <- dat %>%
  left_join(dat_subset, by = "eid")

```


Stiffness
```{r}
datStiffneess_gp$event_dt <- as.Date(datStiffneess_gp$event_dt)
dat_eid$date_baseline <- as.Date(dat_eid$date_baseline)



# Merge dat_subset with datArthralgia_gp
merged_data <- merge(dat_eid, datStiffneess_gp, by = "eid", all.x = TRUE)

# Create the arthralgia_6months, arthralgia_1year, and arthralgia_5years columns
merged_data <- merged_data %>%
  mutate(
    Stiffness_6months = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(6)), TRUE, FALSE),
    Stiffness_1year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(12)), TRUE, FALSE),
    Stiffness_2year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(24)), TRUE, FALSE),
    Stiffness_5years = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(60)), TRUE, FALSE)
  )

# Summarize to get TRUE/FALSE for each eid
dat_subset <- merged_data %>%
  group_by(eid) %>%
  summarize(
    Stiffness_6months = any(Stiffness_6months, na.rm = TRUE),
    Stiffness_1year = any(Stiffness_1year, na.rm = TRUE),
    Stiffness_2year = any(Stiffness_2year, na.rm = TRUE),
    Stiffness_5years = any(Stiffness_5years, na.rm = TRUE)
  ) %>%
  ungroup()

  # Display the resul

dat <- dat %>%
  left_join(dat_subset, by = "eid")
```

Tendonitis
```{r}
datTendonitis_gp$event_dt <- as.Date(datTendonitis_gp$event_dt)
dat_eid$date_baseline <- as.Date(dat_eid$date_baseline)



# Merge dat_subset with datArthralgia_gp
merged_data <- merge(dat_eid, datTendonitis_gp, by = "eid", all.x = TRUE)

# Create the arthralgia_6months, arthralgia_1year, and arthralgia_5years columns
merged_data <- merged_data %>%
  mutate(
    tendonitis_6months = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(6)), TRUE, FALSE),
    tendonitis_1year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(12)), TRUE, FALSE),
    tendonitis_2year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(24)), TRUE, FALSE),
    tendonitis_5years = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(60)), TRUE, FALSE)
  )

# Summarize to get TRUE/FALSE for each eid
dat_subset <- merged_data %>%
  group_by(eid) %>%
  summarize(
    tendonitis_6months = any(tendonitis_6months, na.rm = TRUE),
    tendonitis_1year = any(tendonitis_1year, na.rm = TRUE),
    tendonitis_2year = any(tendonitis_2year, na.rm = TRUE),
    tendonitis_5years = any(tendonitis_5years, na.rm = TRUE)
  ) %>%
  ungroup()

  # Display the resul

dat <- dat %>%
  left_join(dat_subset, by = "eid")
```

Wrist Hand Pain 
```{r}
datWristHandPain_gp$event_dt <- as.Date(datWristHandPain_gp$event_dt)
dat_eid$date_baseline <- as.Date(dat_eid$date_baseline)

# Merge dat_subset with datArthralgia_gp
merged_data <- merge(dat_eid, datWristHandPain_gp, by = "eid", all.x = TRUE)

# Create the arthralgia_6months, arthralgia_1year, and arthralgia_5years columns
merged_data <- merged_data %>%
  mutate(
    WristHandPain_6months = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(6)), TRUE, FALSE),
    WristHandPain_1year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(12)), TRUE, FALSE),
    WristHandPain_2year = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(24)), TRUE, FALSE),
    WristHandPain_5years = ifelse(event_dt < date_baseline & event_dt >= (date_baseline %m-% months(60)), TRUE, FALSE)
  )

# Summarize to get TRUE/FALSE for each eid
dat_subset <- merged_data %>%
  group_by(eid) %>%
  summarize(
    WristHandPain_6months = any(WristHandPain_6months, na.rm = TRUE),
    WristHandPain_1year = any(WristHandPain_1year, na.rm = TRUE),
    WristHandPain_2year = any(WristHandPain_2year, na.rm = TRUE),
    WristHandPain_5years = any(WristHandPain_5years, na.rm = TRUE)
  ) %>%
  ungroup()

  # Display the resul

dat <- dat %>%
  left_join(dat_subset, by = "eid")

```


```{r}

early_cases <- (dat$RA_incident & dat$fu_time <= 1460)

dat <- dat[!(early_cases), ] 
```

```{r}
library(data.table)
library(plyr)
library(dplyr)
library(magrittr)
library(lubridate)
library(readr)
library(survival)
library(Epi)
library(emmeans)
library(ggplot2)
#library(survminer)
#library(ckbplotr)

dir.create("outputs")
```

## Set-up useful functions

```{r}
source("rounding_functions.R")
source("cut_by_quantile.R")
source("cox_functions.R")
```

## Additional data preprocessing

Done in this script not the previous to ensure reference values behave (easy to fix by relevelling otherwise).

```{r}
dat$step_quarters <- qtile_cut(dat$med_steps, probs = seq(0, 1, by = 0.25))
dat$acc_quarters <- qtile_cut(dat$overall_activity, probs = seq(0, 1, by = 0.25), dp_label = 1)
dat$tdi_quarters <- qtile_cut(dat$tdi_raw, probs = seq(0, 1, by = 0.25), dp_label = 1)
dat$cadence30_quarters <- qtile_cut(dat$CadencePeak30Adjusted.steps.min., probs = seq(0,1, by=0.25))
dat$age_gp_crude <- cut(dat$age_entry_years, seq(40, 80, by = 10), right = FALSE, labels = c("40-49", "50-59", "60-69", "70-79"))
dat$BMI_cats <-
  cut(dat$BMI,
      breaks = c(0, 25, 30, 10000),
      labels = c("<24.9", "25.0-29.9", "30.0+"),
      right = FALSE)
```

2 year sens processing
```{r}
dat2yr$step_quarters <- qtile_cut(dat2yr$med_steps, probs = seq(0, 1, by = 0.25))
dat2yr$acc_quarters <- qtile_cut(dat2yr$overall_activity, probs = seq(0, 1, by = 0.25), dp_label = 1)
dat2yr$tdi_quarters <- qtile_cut(dat2yr$tdi_raw, probs = seq(0, 1, by = 0.25), dp_label = 1)
dat2yr$cadence30_quarters <- qtile_cut(dat2yr$CadencePeak30Adjusted.steps.min., probs = seq(0,1, by=0.25))
dat2yr$age_gp_crude <- cut(dat2yr$age_entry_years, seq(40, 80, by = 10), right = FALSE, labels = c("40-49", "50-59", "60-69", "70-79"))
dat2yr$BMI_cats <-
  cut(dat2yr$BMI,
      breaks = c(0, 25, 30, 10000),
      labels = c("<24.9", "25.0-29.9", "30.0+"),
      right = FALSE)
```

```{r}
dat <-dat2yr 
```

```{r}
dat4yr$step_quarters <- qtile_cut(dat4yr$med_steps, probs = seq(0, 1, by = 0.25))
dat4yr$acc_quarters <- qtile_cut(dat4yr$overall_activity, probs = seq(0, 1, by = 0.25), dp_label = 1)
dat4yr$tdi_quarters <- qtile_cut(dat4yr$tdi_raw, probs = seq(0, 1, by = 0.25), dp_label = 1)
dat4yr$cadence30_quarters <- qtile_cut(dat4yr$CadencePeak30Adjusted.steps.min., probs = seq(0,1, by=0.25))
dat4yr$age_gp_crude <- cut(dat4yr$age_entry_years, seq(40, 80, by = 10), right = FALSE, labels = c("40-49", "50-59", "60-69", "70-79"))
dat4yr$BMI_cats <-
  cut(dat4yr$BMI,
      breaks = c(0, 25, 30, 10000),
      labels = c("<24.9", "25.0-29.9", "30.0+"),
      right = FALSE)
```

## Table

```{r}
# MAKE SURE FACTORS NEATLY ORGANISED ================================================
dat$smoking <-
  factor(dat$smoking,
         levels = c("Never", "Previous", "Current"),
         ordered = TRUE)
dat$qualif <-
  factor(
    dat$qualif,
    ordered = TRUE,
    levels = c("School leaver", "Further education", "Higher education")
  )

dat$season_wear <-
  factor(
    dat$season_wear,
    ordered = TRUE,
    levels = c("Spring", "Summer", "Autumn", "Winter")
  )

dat$alcohol <-
  factor(
    dat$alcohol,
    ordered = TRUE,
    levels = c("Never", "<3 times/week", "3+ times/week")
  )

dat <- dat %>%
  mutate(red_processed_total = recode(red_processed_total,
    "[6,1e+05)" = ">6 times per week",
    "[4,6)" = "4-6 timess per week",
    "[2,4)" = "2-4 times per week",
    "[0,2)" = "<2 times per week")) %>%
  mutate(red_processed_total = factor(red_processed_total, levels = c(">6 times per week", "4-6 timess per week","2-4 times per week" , "<2 times per week")))

dat <- dat %>%
  mutate(fruit_veg_total = recode(fruit_veg_total,
    "[11,1e+05)" = ">11 servings",
    "[7,11)" = "7-11 servings",
    "[4,7)" = "4-7 servings",
    "[0,4)" = "<4 servings")) %>%
  mutate(fruit_veg_total = factor(fruit_veg_total, levels = c("<4 servings", "4-7 servings", "7-11 servings", ">11 servings")))


# Set cadence to zero if NA


# FUNCTIONS FOR EXTRACTING NUMBERS======================================================
source("format_for_table.R")

# TABLE 1======================================================
## FIRST ROW
table1 <-
  cbind(data.frame(
    "Characteristic" = c("Overall"),
    "Group" = c(" ")), get_table_numbers(dat, dat)
  )

## LIST AND ITERATE THROUGH VARIABLES 
varlist <-
  list(
    "Sex" = "sex",
    "Age (years)" = "age_gp_crude",
    "Ethnicity" = "ethnicity",
    "Quarter of Townsend Deprivation Index" = "tdi_quarters",
    "Educational Qualifications" = "qualif",
    "Smoking Status" = "smoking",
    "Alcohol Consumption" = "alcohol",
    "Red and processed meat" = "red_processed_total", 
    "Total fruit and vegetables" = "fruit_veg_total",
    "BMI (kg/m2)" = "BMI_cats",
    "Wear Season" = "season_wear"
  )

for (name in names(varlist)) {
  print(name)
 
  variable <- varlist[[name]]
  dat[, variable] <- factor(dat[, variable]) # ensure factors
  factor_levels <- levels(dat[, variable])
  
  for (level in factor_levels) {
    # GET LOCAL DATASET
    dat_loc <- dat[dat[, variable] == level, ]
    
    # FILL OUT VARS FOR DATA FRAME
    char <-
      ifelse(level == factor_levels[1], name, "") # assign characteristic name only for first
    table1_numbers <- get_table_numbers(dat_loc, dat)
    
    # MAKE INTO TABLE
    table1_row <- cbind(
      data.frame(
        "Characteristic" = char,
        "Group" = level), table1_numbers
       
      )
    table1 <- rbind(table1, table1_row)
    
    
    # TIDY
    rm(char,
       dat_loc,
       table1_numbers,
       table1_row)
  }
  # TIDY
  rm(factor_levels, variable)
  
}

write.csv(table1, paste0("outputs/table1_rw_", Sys.Date(), ".csv"), row.names = FALSE)
```


## Modelling

Set up record for models:

```{r}
exposures <- c("step_cats", "step_quarters", "acc_quarters","cadence30_quarters")
outcomes <- c("RA_incident") 

multivar_covs <-
c(
"ethnicity",
"tdi_quarters", 
"qualif",
"smoking",
"alcohol",
"red_processed_total",
"fruit_veg_total"
)

adjustments <-
list(
"model1" = NULL,
"model2" = c(strata("sex")),
"model3" = c(strata("sex"), multivar_covs),
"model4" = c(strata("sex"), multivar_covs, "BMI_cats"),
"model5" = c(strata("sex"), "PRS_RA_RiskGroup"),
"model6" = c(strata("sex"), multivar_covs,"PRS_RA_RiskGroup"), 
"model7" = c(strata("sex"), multivar_covs,"PRS_RA_RiskGroup","BMI_cats"), 
"model8" = c(strata("sex"), multivar_covs,"age_gp_crude") # this is just a crude check that
)

```

```{r}
sd(dat$med_steps) #3822.266
# Generate a standardization variable
dat$med_steps_std = (dat$med_steps - mean(dat$med_steps))/sd(dat$med_steps)
# Check distribution of variable
hist(dat$med_steps_std)
```


```{r}
cadence_1st_percentile <- quantile(dat$CadencePeak30Adjusted.steps.min., 0.01)  # Calculate the 1st percentile

# Filter the data to remove values less than the 1st percentile
dat <- dat %>%
  mutate(cad30_clipped = pmax(CadencePeak30Adjusted.steps.min., cadence_1st_percentile))
sd(dat$cad30_clipped) #3822.266
# Generate a standardization variable
dat$cad30_std = (dat$cad30_clipped - mean(dat$cad30_clipped))/sd(dat$cad30_clipped)
# Check distribution of variable
hist(dat$cad30_std)
```


```{r}

library("multcomp", "foreign", "biostat3")
cox_BMI_cad <- coxph((Surv(age_entry_days, age_exit_days, RA_incident)) ~ 
                           cad30_std*BMI_cats+ strata(sex)+
                           as.factor(ethnicity) + as.factor(tdi_quarters) +as.factor(qualif)+
                           as.factor(season_wear) + as.factor(smoking) + as.factor(alcohol)+as.factor(processed_meat)+as.factor(fresh_fruit)+as.factor(oily_fish),
                         data=dat)

summary(cox_BMI_cad)


subgroup_BMI_cad <- lincom(cox_BMI_cad, c("cad30_std",
                          "cad30_std+cad30_std:BMI_cats25.0-29.9",
                          "cad30_std+cad30_std:BMI_cats30.0+"),
       eform=TRUE)
subgroup_BMI_cad


```


Subgroup analysis in those with high BMI 
```{r}

library("multcomp", "foreign", "biostat3", "car")
cox_BMI <- coxph((Surv(age_entry_days, age_exit_days, RA_incident)) ~ 
                           med_steps_std*BMI_cats+ strata(sex)+
                           as.factor(ethnicity) + as.factor(tdi_quarters) +as.factor(qualif)+
                           as.factor(season_wear) + as.factor(smoking) + as.factor(alcohol)+as.factor(processed_meat)+as.factor(fresh_fruit)+as.factor(oily_fish),
                         data=dat)

summary(cox_BMI)


sub_BMI <- lincom(cox_BMI, c("med_steps_std",
                          "med_steps_std+med_steps_std:BMI_cats25.0-29.9",
                          "med_steps_std+med_steps_std:BMI_cats30.0+"),
       eform=TRUE)
as.data.frame(sub_BMI)

ggplot(sub_BMI, aes(x = Estimate, y = Term)) +
  geom_point() +
  geom_errorbarh(aes(xmin = LowerCI, xmax = UpperCI), height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  theme_minimal() +
  labs(x = "Hazard Ratio (95% CI)", y = "")


```

```{r}
write.csv(sub_BMI, "sub_BMI_output.csv")
str(sub_BMI)

df <- data.frame(
  Estimate = sub_BMI$Estimate,
  LowerCI = sub_BMI$`2.5 %`,
  UpperCI = sub_BMI$`97.5 %`,
  Term = c("med_steps_std", "med_steps_std:BMI_cats25.0-29.9", "med_steps_std:BMI_cats30.0+")
)

sub_BMI$Term <- c("med_steps_std", "med_steps_std:BMI_cats25.0-29.9", "med_steps_std:BMI_cats30.0+")

ggplot(sub_BMI, aes(y = Term, x = Estimate, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_point() +
  geom_errorbarh(height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  theme_minimal() +
  labs(x = "Hazard Ratio (95% CI)", y = "")
```

Rheumatoid Factor positive 
```{r}
library("multcomp", "foreign", "biostat3")

dat$RF_binary <- as.factor(dat$RF_binary)

dat_RF <- dat[!is.na(dat$RF_binary), ]

cox_PRS <- coxph((Surv(age_entry_days, age_exit_days, RA_incident)) ~ 
                           med_steps_std + med_steps_std:PRS_RA_RiskGroup +as.factor(ethnicity) + as.factor(tdi_quarters)
                 +as.factor(qualif)+as.factor(season_wear) + as.factor(smoking) +
                   as.factor(alcohol)+as.factor(processed_meat)+as.factor(fresh_fruit)+as.factor(oily_fish),
                         data=dat)

summary(cox_PRS)

subgroupPRS <- lincom(cox_PRS, c("med_steps_std",
                                "med_steps_std+med_steps_std:PRS_RA_RiskGroupLow",
                               "med_steps_std+med_steps_std:PRS_RA_RiskGroupIntermediate"),
       eform=TRUE)
subgroupPRS
```

```{r}
cox_sex <- coxph((Surv(age_entry_days, age_exit_days, RA_incident)) ~ 
                           med_steps_std + med_steps_std:strata(sex)+as.factor(ethnicity) + as.factor(tdi_quarters)
                 +as.factor(qualif)+as.factor(season_wear) + as.factor(smoking) +
                   as.factor(alcohol)+as.factor(processed_meat)+as.factor(fresh_fruit)+as.factor(oily_fish),
                         data=dat)

#p-value on the interaction term between sex + med steps, if significnat  then suggests heterogeneity betweens males and females -> 0.118
summary(cox_sex)


subgroup_sex <- lincom(cox_sex, c("med_steps_std",
                                  "med_steps_std+med_steps_std:strata(sex)Male"),
                       eform=TRUE)
subgroup_sex

```

```{r}


cox_age <- coxph((Surv(age_entry_days, age_exit_days, RA_incident)) ~ 
                           med_steps_std + med_steps_std:age_gp_crude +as.factor(ethnicity) + as.factor(tdi_quarters)
                 +as.factor(qualif)+as.factor(season_wear) + as.factor(smoking) +
                   as.factor(alcohol)+as.factor(processed_meat)+as.factor(fresh_fruit)+as.factor(oily_fish),
                         data=dat)

#p-value on the interaction term between sex + med steps, if significnat  then suggests heterogeneity betweens males and females -> 0.118
summary(cox_age)


subgroup_age <- lincom(cox_age, c("med_steps_std",
                                  "med_steps_std+med_steps_std:age_gp_crude50-59",
                                  "med_steps_std+med_steps_std:age_gp_crude60-69",
                                  "med_steps_std+med_steps_std:age_gp_crude70-79"),
                                    eform=TRUE)
subgroup_age

```


```{r}
for (factor in c("sex", "BMI_cats","PRS_RA_RiskGroup", multivar_covs, exposures)){
  dat[, factor] <- factor(dat[, factor], ordered = FALSE)
} # note all are factors so this works

```
2 year model
```{r}
for (factor in c("sex", "BMI_cats","PRS_RA_RiskGroup", multivar_covs, exposures)){
  dat2yr[, factor] <- factor(dat2yr[, factor], ordered = FALSE)
} # note all are factors so this works

```

4 year sensitivity
```{r}
for (factor in c("sex", "BMI_cats","PRS_RA_RiskGroup", multivar_covs, exposures)){
  dat4yr[, factor] <- factor(dat4yr[, factor], ordered = FALSE)
} # note all are factors so this works

```


We select the relevant columns for our analysis

```{r}
covariates <- c("ethnicity","qualif","smoking","alcohol","tdi_quarters",
                "red_processed_total","fruit_veg_total","sex")
mediators <- c("BMI_cats")
out_vars <- c("age_entry_days", "age_exit_days", "RA_incident")
primary <- c("med_steps_std")

cols <- c(covariates, mediators, out_vars, primary)

missing_columns <- setdiff(cols, colnames(dat))
if (length(missing_columns) > 0) {
  err_message <- paste("The following columns are not found in the dataset: ", 
                       paste(missing_columns, collapse = ", "))
  stop(err_message)
} else {
  datCheck <- dat[, cols]
}
```

We then test the departure from linearity for each of the relevant covariates

1. Steps
If the p value is < 0.05, it means we cannot use this variable as continuous.
Cadence 30 quarters -  0.1909
Steps_quarters - 0.2856
Acc_quarters -  0.6179
```{r}
# Perform the likelihood ratio test
modelA <- coxph(as.formula(paste("Surv(", paste(out_vars, collapse = ", "), 
                                 ") ~ as.numeric(step_quarters) + factor(sex)")), 
                data = dat)

modelB <- coxph(as.formula(paste("Surv(", paste(out_vars, collapse = ", "), 
                                 ") ~ factor(step_quarters) + factor(sex)")), 
                data = dat)

lr_test <- lmtest::lrtest(modelA, modelB)

print(lr_test)
```

```{r}
dat$std_steps <- standardise_column(dat[[primary]], 1000, TRUE)
```

We confirm the proportional hazards assumption

```{r}
# Proportional Hazard Model
formula <- as.formula(paste("Surv(", paste(out_vars, collapse = ", "), ") ~ med_steps_std +", 
                            paste(c(covariates, mediators), collapse = "+")))

model_ph_test <- coxph(formula, data = dat)

summary(model_ph_test)

hazards_table <- cox.zph(model_ph_test)
print(hazards_table)

# Schoenfeld Residuals Plot
for (i in seq_along(c(covariates, mediators, "std_steps"))) {
  # Create a Schoenfeld Residuals Plot for the i-th variable
  plot(hazards_table, var = i, main = "Schoenfeld Residual Plot")
  abline(h = coef(model_ph_test)[2], lty = 2, col = "red")
}
```

Cadence plotting
```{r}
for (exposure in exposures) {
  # SET UP RECORDING FRAME ======================================================
  results_columns <-   c("Exposure",
                         "Outcome",
                         "Model",
                         "Adjustment",
                         "n",
                         "n_event",
                         as.vector(outer(
                           # excessively complex code to get crossed column names
                           c(
                             "HR",
                             "Lower_CI",
                             "Upper_CI",
                             "floatedlnHR",
                             "floatedSE",
                             "floatedHR",
                             "floatedLower_CI",
                             "floatedUpper_CI",
                             "n",
                             "n_event", 
                             "mean_steps", 
                             "mean_acc",
                             "mean_cad30"
                           ),
                           unique(dat[, exposure]),
                           paste,
                           sep = "_"
                         )))
  results_tab <-
    data.frame(matrix(ncol = length(results_columns) ,
                      nrow = 0))
  colnames(results_tab) <- results_columns
  
  # Do modelling: ===============================================================
  for (i in 1:length(outcomes)) {
    for (j in 1:length(adjustments)) {
      # Pull relevant values-----------------------------
      outcome <- outcomes[i]
      model_name <- names(adjustments)[j]
      covs <- adjustments[[j]]
      
      # Organise for model calculation-------------------
      cov_sum <- paste0(c(exposure, covs), collapse = "+")
      
      # Model--------------------------------------------
      form <-
        paste0("Surv(age_entry_days, age_exit_days, ",
               outcome,
               ") ~ ",
               cov_sum)
      model <-
        coxph(as.formula(form), dat)
      print(cox.zph(model))
      
      # Extract values-----------------------------------
      results_frame <-
        data.frame(
          "Exposure" = exposure,
          "Outcome" = outcome,
          "Model" = model_name,
          "Adjustment" = paste0(covs, collapse = ","),
          "n" = model$n,
          "n_event" = model$nevent
        )
      
      # CYCLE THROUGH EXPOSURE LEVELS EXTRACTING INFO=============================================
      fac_levels <- levels(factor(dat[, exposure]))
      for (cat in fac_levels) {
        
        # HRs=====================================================================================
        if (cat == fac_levels[1]) {
          HRs <- data.frame(matrix(c(1, 1, 1), nrow = 1))
        }
        else {
          HRs <-
            summary(model)$conf.int[paste0(exposure, cat), c("exp(coef)", "lower .95", "upper .95"), drop = FALSE]
          rownames(HRs) <- NULL
        }
        colnames(HRs) <-
          as.vector(outer(c("HR", "Lower_CI", "Upper_CI"), cat, paste, sep = "_"))
        
        # FLOATED HRs etc =======================================================================
        float_model <- float(model)# This is doing floating absolute risks using Epi package
        
        ## Values
        zval <- qnorm(0.975)
        se <- sqrt(float_model$var[cat])
        lnhr <- float_model$coef[cat]
        
        # Collate into frame
        HRs_floated <-
          data.frame(matrix(c(
            lnhr,
            se,
            exp(lnhr),
            exp(lnhr - se[cat] * zval),
            exp(lnhr + se[cat] * zval)
          ), nrow = 1))
        colnames(HRs_floated) <-
          as.vector(outer(
            c(
              "floatedlnHR",
              "floatedSE",
              "floatedHR",
              "floatedLower_CI",
              "floatedUpper_CI"
            ),
            cat,
            paste,
            sep = "_"
          ))
    
        # n and nevent ===================================================================================
        n_group <- nrow(dat[dat[, exposure] == cat, ])
        n_event_group <- nrow(dat[(dat[, exposure] == cat) & (dat[, outcome]), ])
        n_by_group <- data.frame(matrix(c(n_group, n_event_group), nrow = 1))
        colnames(n_by_group) <-
          as.vector(outer(c("n", "n_event"), cat, paste, sep = "_"))
        
        # Mean exposure ==================================================================================
        col_me <- paste0("mean_steps_", cat)
        col_acc <- paste0("mean_acc_", cat)
        col_cad30 <- paste0("mean_cad30_", cat)
        mean_exposure <- data.frame(mean(dat$med_steps[dat[, exposure] == cat]), 
                                    mean(dat$overall_activity[dat[, exposure] == cat]),
                                    mean(dat$CadencePeak30Adjusted.steps.min.[dat[, exposure] == cat]))
        colnames(mean_exposure) <- c(col_me, col_acc,col_cad30)
        
        # Bind together in final data frame ==============================================================
        results_frame <-
          cbind(results_frame, HRs, HRs_floated, n_by_group, mean_exposure)
      }
      
      
      # CHECK THAT n and n events match by different routes
      nsum <- sum(results_frame[,  as.vector(outer(c("n"), fac_levels, paste, sep = "_"))])
      n_eventsum <- sum(results_frame[,  as.vector(outer(c("n_event"), fac_levels, paste, sep = "_"))])
      if ((results_frame$n != nsum)|(results_frame$n_event != n_eventsum)){
        stop("Mismatch between numbers derived from different sources. Recheck model. May be that some rows are being excluded due to missing covariate data.")
      }
      
      
      # BIND INTO FINAL RESULTS FRAME AND DELETE
      results_tab <- rbind(results_tab, results_frame)
      
      rm(results_frame)
      
    }
  }
  
  assign(paste0(exposure, "_results_tab"), results_tab)
  write.csv(results_tab, paste0("outputs/", exposure, "_tab.csv"))
  # print(results_tab)
  rm(results_tab, results_columns)
}
```

Log-log plots to visually explore variables that violate PH assumption 
```{r}

#Log-log for tdi 
tdi_quarters_log <- survfit(Surv(age_entry_days,age_exit_days, RA_incident) ~ as.factor(tdi_quarters), data = dat)
#autoplot(km_home) # just to see the km curves

plot(tdi_quarters_log, fun = "cloglog", xlab = "Time (in days) using log",
     ylab = "log-log survival", main = "log-log curves by tdi_quarters")

#log-log for sex 
sex_log <- survfit(Surv(age_entry_days,age_exit_days, RA_incident) ~ as.factor(sex), data = dat)
#autoplot(km_home) # just to see the km curves

plot(sex_log, fun = "cloglog", xlab = "Time (in days) using log",
     ylab = "log-log survival", main = "log-log curves by sex")

#log-log for BMI cats
BMIcats_log <- survfit(Surv(age_entry_days,age_exit_days, RA_incident) ~ as.factor(BMI_cats), data = dat)
#autoplot(km_home) # just to see the km curves

plot(BMIcats_log, fun = "cloglog", xlab = "Time (in days) using log",
     ylab = "log-log survival", main = "log-log curves by BMI_cats")

```




Modelling - 2 years sensitivity

2year sensitivity
```{r}
for (exposure in exposures) {
  # SET UP RECORDING FRAME ======================================================
  results_columns <-   c("Exposure",
                         "Outcome",
                         "Model",
                         "Adjustment",
                         "n",
                         "n_event",
                         as.vector(outer(
                           # excessively complex code to get crossed column names
                           c(
                             "HR",
                             "Lower_CI",
                             "Upper_CI",
                             "floatedlnHR",
                             "floatedSE",
                             "floatedHR",
                             "floatedLower_CI",
                             "floatedUpper_CI",
                             "n",
                             "n_event", 
                             "mean_steps", 
                             "mean_acc",
                             "mean_cad30"
                           ),
                           unique(dat2yr[, exposure]),
                           paste,
                           sep = "_"
                         )))
  results_tab <-
    data.frame(matrix(ncol = length(results_columns) ,
                      nrow = 0))
  colnames(results_tab) <- results_columns
  
  # Do modelling: ===============================================================
  for (i in 1:length(outcomes)) {
    for (j in 1:length(adjustments)) {
      # Pull relevant values-----------------------------
      outcome <- outcomes[i]
      model_name <- names(adjustments)[j]
      covs <- adjustments[[j]]
      
      # Organise for model calculation-------------------
      cov_sum <- paste0(c(exposure, covs), collapse = "+")
      
      # Model--------------------------------------------
      form <-
        paste0("Surv(age_entry_days, age_exit_days, ",
               outcome,
               ") ~ ",
               cov_sum)
      model <-
        coxph(as.formula(form), dat2yr)
      print(cox.zph(model))
      
      # Extract values-----------------------------------
      results_frame <-
        data.frame(
          "Exposure" = exposure,
          "Outcome" = outcome,
          "Model" = model_name,
          "Adjustment" = paste0(covs, collapse = ","),
          "n" = model$n,
          "n_event" = model$nevent
        )
      
      # CYCLE THROUGH EXPOSURE LEVELS EXTRACTING INFO=============================================
      fac_levels <- levels(factor(dat2yr[, exposure]))
      for (cat in fac_levels) {
        
        # HRs=====================================================================================
        if (cat == fac_levels[1]) {
          HRs <- data.frame(matrix(c(1, 1, 1), nrow = 1))
        }
        else {
          HRs <-
            summary(model)$conf.int[paste0(exposure, cat), c("exp(coef)", "lower .95", "upper .95"), drop = FALSE]
          rownames(HRs) <- NULL
        }
        colnames(HRs) <-
          as.vector(outer(c("HR", "Lower_CI", "Upper_CI"), cat, paste, sep = "_"))
        
        # FLOATED HRs etc =======================================================================
        float_model <- float(model)# This is doing floating absolute risks using Epi package
        
        ## Values
        zval <- qnorm(0.975)
        se <- sqrt(float_model$var[cat])
        lnhr <- float_model$coef[cat]
        
        # Collate into frame
        HRs_floated <-
          data.frame(matrix(c(
            lnhr,
            se,
            exp(lnhr),
            exp(lnhr - se[cat] * zval),
            exp(lnhr + se[cat] * zval)
          ), nrow = 1))
        colnames(HRs_floated) <-
          as.vector(outer(
            c(
              "floatedlnHR",
              "floatedSE",
              "floatedHR",
              "floatedLower_CI",
              "floatedUpper_CI"
            ),
            cat,
            paste,
            sep = "_"
          ))
    
        # n and nevent ===================================================================================
        n_group <- nrow(dat2yr[dat2yr[, exposure] == cat, ])
        n_event_group <- nrow(dat2yr[(dat2yr[, exposure] == cat) & (dat2yr[, outcome]), ])
        n_by_group <- data.frame(matrix(c(n_group, n_event_group), nrow = 1))
        colnames(n_by_group) <-
          as.vector(outer(c("n", "n_event"), cat, paste, sep = "_"))
        
        # Mean exposure ==================================================================================
        col_me <- paste0("mean_steps_", cat)
        col_acc <- paste0("mean_acc_", cat)
        col_cad30 <- paste0("mean_cad30_", cat)
        mean_exposure <- data.frame(mean(dat2yr$med_steps[dat2yr[, exposure] == cat]), 
                                    mean(dat2yr$overall_activity[dat2yr[, exposure] == cat]),
                                    mean(dat2yr$CadencePeak30Adjusted.steps.min.[dat2yr[, exposure] == cat]))
        colnames(mean_exposure) <- c(col_me, col_acc,col_cad30)
        
        # Bind together in final data frame ==============================================================
        results_frame <-
          cbind(results_frame, HRs, HRs_floated, n_by_group, mean_exposure)
      }
      
      
      # CHECK THAT n and n events match by different routes
      nsum <- sum(results_frame[,  as.vector(outer(c("n"), fac_levels, paste, sep = "_"))])
      n_eventsum <- sum(results_frame[,  as.vector(outer(c("n_event"), fac_levels, paste, sep = "_"))])
      if ((results_frame$n != nsum)|(results_frame$n_event != n_eventsum)){
        stop("Mismatch between numbers derived from different sources. Recheck model. May be that some rows are being excluded due to missing covariate data.")
      }
      
      
      # BIND INTO FINAL RESULTS FRAME AND DELETE
      results_tab <- rbind(results_tab, results_frame)
      
      rm(results_frame)
      
    }
  }
  
  assign(paste0(exposure, "_results_tab"), results_tab)
  write.csv(results_tab, paste0("outputs/", exposure, "_tab2yr.csv"))
  # print(results_tab)
  rm(results_tab, results_columns)
}
```



Modelling - 4 year
```{r}
for (exposure in exposures) {
  # SET UP RECORDING FRAME ======================================================
  results_columns <-   c("Exposure",
                         "Outcome",
                         "Model",
                         "Adjustment",
                         "n",
                         "n_event",
                         as.vector(outer(
                           # excessively complex code to get crossed column names
                           c(
                             "HR",
                             "Lower_CI",
                             "Upper_CI",
                             "floatedlnHR",
                             "floatedSE",
                             "floatedHR",
                             "floatedLower_CI",
                             "floatedUpper_CI",
                             "n",
                             "n_event", 
                             "mean_steps", 
                             "mean_acc",
                             "mean_cad30"
                           ),
                           unique(dat4yr[, exposure]),
                           paste,
                           sep = "_"
                         )))
  results_tab <-
    data.frame(matrix(ncol = length(results_columns) ,
                      nrow = 0))
  colnames(results_tab) <- results_columns
  
  # Do modelling: ===============================================================
  for (i in 1:length(outcomes)) {
    for (j in 1:length(adjustments)) {
      # Pull relevant values-----------------------------
      outcome <- outcomes[i]
      model_name <- names(adjustments)[j]
      covs <- adjustments[[j]]
      
      # Organise for model calculation-------------------
      cov_sum <- paste0(c(exposure, covs), collapse = "+")
      
      # Model--------------------------------------------
      form <-
        paste0("Surv(age_entry_days, age_exit_days, ",
               outcome,
               ") ~ ",
               cov_sum)
      model <-
        coxph(as.formula(form), dat4yr)
      print(cox.zph(model))
      
      # Extract values-----------------------------------
      results_frame <-
        data.frame(
          "Exposure" = exposure,
          "Outcome" = outcome,
          "Model" = model_name,
          "Adjustment" = paste0(covs, collapse = ","),
          "n" = model$n,
          "n_event" = model$nevent
        )
      
      # CYCLE THROUGH EXPOSURE LEVELS EXTRACTING INFO=============================================
      fac_levels <- levels(factor(dat4yr[, exposure]))
      for (cat in fac_levels) {
        
        # HRs=====================================================================================
        if (cat == fac_levels[1]) {
          HRs <- data.frame(matrix(c(1, 1, 1), nrow = 1))
        }
        else {
          HRs <-
            summary(model)$conf.int[paste0(exposure, cat), c("exp(coef)", "lower .95", "upper .95"), drop = FALSE]
          rownames(HRs) <- NULL
        }
        colnames(HRs) <-
          as.vector(outer(c("HR", "Lower_CI", "Upper_CI"), cat, paste, sep = "_"))
        
        # FLOATED HRs etc =======================================================================
        float_model <- float(model)# This is doing floating absolute risks using Epi package
        
        ## Values
        zval <- qnorm(0.975)
        se <- sqrt(float_model$var[cat])
        lnhr <- float_model$coef[cat]
        
        # Collate into frame
        HRs_floated <-
          data.frame(matrix(c(
            lnhr,
            se,
            exp(lnhr),
            exp(lnhr - se[cat] * zval),
            exp(lnhr + se[cat] * zval)
          ), nrow = 1))
        colnames(HRs_floated) <-
          as.vector(outer(
            c(
              "floatedlnHR",
              "floatedSE",
              "floatedHR",
              "floatedLower_CI",
              "floatedUpper_CI"
            ),
            cat,
            paste,
            sep = "_"
          ))
    
        # n and nevent ===================================================================================
        n_group <- nrow(dat4yr[dat4yr[, exposure] == cat, ])
        n_event_group <- nrow(dat4yr[(dat4yr[, exposure] == cat) & (dat4yr[, outcome]), ])
        n_by_group <- data.frame(matrix(c(n_group, n_event_group), nrow = 1))
        colnames(n_by_group) <-
          as.vector(outer(c("n", "n_event"), cat, paste, sep = "_"))
        
        # Mean exposure ==================================================================================
        col_me <- paste0("mean_steps_", cat)
        col_acc <- paste0("mean_acc_", cat)
        col_cad30 <- paste0("mean_cad30_", cat)
        mean_exposure <- data.frame(mean(dat4yr$med_steps[dat4yr[, exposure] == cat]), 
                                    mean(dat4yr$overall_activity[dat4yr[, exposure] == cat]),
                                    mean(dat4yr$CadencePeak30Adjusted.steps.min.[dat4yr[, exposure] == cat]))
        colnames(mean_exposure) <- c(col_me, col_acc,col_cad30)
        
        # Bind together in final data frame ==============================================================
        results_frame <-
          cbind(results_frame, HRs, HRs_floated, n_by_group, mean_exposure)
      }
      
      
      # CHECK THAT n and n events match by different routes
      nsum <- sum(results_frame[,  as.vector(outer(c("n"), fac_levels, paste, sep = "_"))])
      n_eventsum <- sum(results_frame[,  as.vector(outer(c("n_event"), fac_levels, paste, sep = "_"))])
      if ((results_frame$n != nsum)|(results_frame$n_event != n_eventsum)){
        stop("Mismatch between numbers derived from different sources. Recheck model. May be that some rows are being excluded due to missing covariate data.")
      }
      
      
      # BIND INTO FINAL RESULTS FRAME AND DELETE
      results_tab <- rbind(results_tab, results_frame)
      
      rm(results_frame)
      
    }
  }
  
  assign(paste0(exposure, "_results_tab"), results_tab)
  write.csv(results_tab, paste0("outputs/", exposure, "_tab4yr.csv"))
  # print(results_tab)
  rm(results_tab, results_columns)
}
```


## Estimated marginal means analyses

There is an issue with excessively large reference grid, when adjusting for all variables, meaning you tend to crash it with the full set. In which case, rather than an ad hoc covariate set, this code uses a minimally adjusted (i.e. just age and sex).

```{r}
source("useful_functions/load_disease_codelist.R")
disease_names <- paste0("prev_hes_", names(disease_codelist))
# emm_options(rg.limit = 2000000) # if include many covariates uncomment as needs bigger grid
rec_dat <- data.frame("Quality" = c(), "Mean" = c(), "LowerCI" = c(), "UpperCI" = c())
for (exposure in c("sr_overall_health", disease_names)){
  # Set up model
  form <- as.formula(paste0(c("med_steps ~ age_entry_days", "sex", exposure), collapse = "+"))
  model <- lm(form, data = dat)
  emmeans_model <- emmeans(model, specs = exposure)
  
  # Fill out data frame
  rec_dat_rows <- summary(emmeans_model)[,c(exposure, "emmean", "lower.CL", "upper.CL")]
  colnames(rec_dat_rows) <- c("Quality", "Mean","LowerCI", "UpperCI")
  
  # Do by binding - not ideal as slower than pre-assigning data frame but needed here for flexibility
  rec_dat <- rbind(rec_dat, rec_dat_rows)
  
  # Assign and tidy
  assign(paste0(exposure, "_model"), model)
  assign(paste0("emmeans_", exposure, "_model"), emmeans_model)
  rm(form, model, emmeans_model, rec_dat_rows)
}

rec_dat

```

```{r}
# Format rownames
rec_dat$Quality <- sub("prev_hes_", "", rec_dat$Quality)
rec_dat$Quality <- gsub("_", " ", rec_dat$Quality)
write.csv(rec_dat, "outputs/forest_plot_tab.csv")
```

## Additional analysis (Scott)

```{r}
cor.test(dat$overall_activity, dat$med_steps, method = "spearman", conf.level = 0.95) # Correlation between Steps and Acceleration

#Self reported health
health_model <- aov(data = dat, med_steps ~ sr_overall_health)
TukeyHSD(health_model)
plot(TukeyHSD(health_model))

#Self reported walking pace
dat_no_sr_usual_walking_pace <- dat[!(dat$sr_usual_walking_pace == "Missing"), ] # Note this should not be in people missing this data
pace_model <- aov(data = dat_no_sr_usual_walking_pace, mean_one_minute_cadence ~ sr_usual_walking_pace)
TukeyHSD(pace_model)
plot(TukeyHSD(pace_model))  
```

# Clear up some of the mess ahead of running future scripts

Not strictly necessary but hopefully avoids me accidentally relying on stuff in later scripts.

```{r}
rm(list = setdiff(ls(), lsf.str())) # this setdiff is listing everything then listing only functions. So it's saying remove everything that's not a function (see https://stackoverflow.com/questions/8305754/remove-all-variables-except-functions) 
```