---
title: "ForestPlots"
author: "Dylan McGagh"
date: "`r format(Sys.time(), '%a %b %d, %Y %X')`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    toc_depth: 5
    theme: lumen
---
```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This script runs models for 1 SD increase in exposure.

# Set up workspace

## Set-up useful functions and packages
```{r}
#Load packages
pkgs <- c("data.table", "rmarkdown", "plyr", "lubridate", "magrittr", "ggplot2", "data.table", "dplyr", "stringr", "splitstackshape", "scales", "haven", "Epi", "emmeans", "rtables", "table1",  "tableone", "boot", "knitr", "pspline", "survival", "xtable", "here", "forcats", "patchwork") # packages we need

pkgs_inst <- pkgs[!{pkgs %in% rownames(installed.packages())}] # check which are not present 
install.packages(pkgs_inst, repos = "https://www.stats.bris.ac.uk/R/") # install
lapply(pkgs, library, character.only = TRUE) #load all packages
```


## Load data
```{r, warning=FALSE, error=FALSE, message=FALSE, results = 'hide'}
## Outcome dfs
#updated file path 2023 10 03
steps_datLow <- fread(here("PRSbyGroup/steps_sdmodelresultsLow.csv"), data.table = FALSE)
steps_datInter <- fread(here("PRSbyGroup/steps_sdmodelresultsInter.csv"), data.table = FALSE)
steps_datHigh <- fread(here("PRSbyGroup/steps_sdmodelresultsHigh.csv"), data.table = FALSE)
```


# Forest plots
### Steps - prep dataframe
Combine results from sex specific models and main model
```{r, warning=FALSE, error=FALSE, message=FALSE, results = 'hide'}

# Drop rows where HR..95..CI. = "REF"
steps_datLowPlot <- steps_datLow %>%
  filter(!str_detect(`HR..95..CI.`, "REF"))

# Rename column
steps_datLowPlot <- steps_datLowPlot %>%
  rename(estimate = `HR..95..CI.`)

# Extract the lower and upper bounds of the estimates from the "estimate" column
steps_datLowPlot$lower <- as.numeric(gsub(".*\\((.*) -.*", "\\1", steps_datLowPlot$estimate))
steps_datLowPlot$upper <- as.numeric(gsub(".*- (.*)\\)", "\\1", steps_datLowPlot$estimate))
steps_datLowPlot$hr <- as.numeric(gsub(" \\(.*", "", steps_datLowPlot$estimate))

# Drop rows where HR..95..CI. = "REF"
steps_datInterPlot <- steps_datInter %>%
  filter(!str_detect(`HR..95..CI.`, "REF"))

# Rename column
steps_datInterPlot <- steps_datInterPlot %>%
  rename(estimate = `HR..95..CI.`)

# Extract the lower and upper bounds of the estimates from the "estimate" column
steps_datInterPlot$lower <-as.numeric(gsub(".*\\((.*) -.*", "\\1",steps_datInterPlot$estimate))
steps_datInterPlot$upper <-as.numeric(gsub(".*- (.*)\\)", "\\1",steps_datInterPlot$estimate))
steps_datInterPlot$hr <-as.numeric(gsub(" \\(.*", "",steps_datInterPlot$estimate))

# Drop rows where HR..95..CI. = "REF"
steps_datHighPlot <- steps_datHigh %>%
  filter(!str_detect(`HR..95..CI.`, "REF"))

# Rename column
steps_datHighPlot <- steps_datHighPlot %>%
  rename(estimate = `HR..95..CI.`)

# Extract the lower and upper bounds of the estimates from the "estimate" column
steps_datHighPlot$lower <-as.numeric(gsub(".*\\((.*) -.*", "\\1",steps_datHighPlot$estimate))
steps_datHighPlot$upper <-as.numeric(gsub(".*- (.*)\\)", "\\1",steps_datHighPlot$estimate))
steps_datHighPlot$hr <-as.numeric(gsub(" \\(.*", "",steps_datHighPlot$estimate))

```

Create new column in each subgroup specificing the PRS Risk Group 
```{r}
steps_datLowPlot$Group <- "PRS Low Risk Group"
steps_datInterPlot$Group <- "PRS Intermediate Risk Group"
steps_datHighPlot$Group <- "PRS High Risk Group"
```

For plotting - we want to display the 1sd results for each PRS subgroup on the one plot so need to create new dataframe with model outputs for each merged 

```{r}
datLow <- subset(steps_datLowPlot, Model == "Fully-adjusted")
datInter <- subset(steps_datInterPlot, Model == "Fully-adjusted")
datHigh <- subset(steps_datHighPlot, Model == "Fully-adjusted")

#Merge the 3 


# Assuming datLow, datInter, and datHigh are your data frames
datTotal <- bind_rows(datLow, datInter, datHigh)

custom_order <- c("PRS High Risk Group", "PRS Intermediate Risk Group", "PRS Low Risk Group")

# Use factor() to reorder "Group" based on the custom order
datTotal$Group <- factor(datTotal$Group, levels = custom_order)


```


#### Steps - fully adjusted by PRS risk group 

```{r}

# Create a named vector to define the custom order

# Use factor() to reorder "Group" based on the custom order within ggplot
plot <- ggplot(datTotal, aes(x = hr, y = factor(Group, levels = custom_order))) +
  geom_point(size = 3) +
  ggtitle("Association of Median Daily Step Count with Risk of\nIncident Rheumatoid Arthritis by PRS Risk Group")+
  geom_segment(aes(x = lower, xend = upper, y = factor(Group, levels = custom_order), yend = factor(Group, levels = custom_order)), size = 1) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  geom_text(aes(y = factor(Group, levels = custom_order), x = 0, 
                  label = paste0(format(round(hr, digits = 2), nsmall = 2), "
                                 (",format(round(lower, digits = 2), nsmall = 2), "-", 
                                 format(round(upper, digits = 2), nsmall = 2), ")")), size = 3, 
            hjust = 0) +
  labs(x = "HR per 1-SD increase in median daily step count (95% CI)") +
  ylab(NULL) +
  theme_bw() +
  scale_x_continuous(limits = c(0.2, 1.3),breaks = seq(0, 1.3, by = 0.2)) +
  theme(strip.background = element_blank(), 
        strip.placement = "outside", 
        axis.text.x = element_text(colour = "black"), 
        strip.text.y.left = element_text(angle = 90, hjust = 0.5, size = 8),
        plot.title = element_text(size = 12,hjust = 0.5))

plot

```

