GRAPPA Analysis

# First we need to install packages that aren't already present
```{r}
pkgs <- c("data.table", "ggplot2", "survival", "table1", "dplyr", "emmeans", "data.table","gridExtra","lubridate","table1") # packages we need
pkgs_inst <- pkgs[!{pkgs %in% rownames(installed.packages())}] # check which are not present 
install.packages(pkgs_inst, repos = "https://www.stats.bris.ac.uk/R/")

# Load packages
lapply(pkgs, library, character.only = TRUE)
# using lapply just allows us to load several packages on one line
```

```{bash}
cd ~/rap_wearables

dx download /users/mcgaghd/GP_final.csv -r 
```

We load the data we prepared earlier (noting again that the code is different depending on whether you're running it in the same session as the earlier scripts, or have come back to do this later): 
```{r}
dat <- fread("GP_final.csv", data.table = FALSE) # version if running in same session as 
```

We summarise to understand the data structure:
```{r}
# summary(dat) # uncomment to run, avoiding printing lots of information on internet
#We create factors and make sure they have sensible ordering of levels (e.g. large-ish reference category): 
dat$ethnicity <- factor(dat$ethnicity, levels = c("White", "Nonwhite"))
dat$tdi_cats <- factor(dat$tdi_cats, levels = c("Least Deprived", "2nd Quintile", "3rd Quintile", "4th Quintile", "Most Deprived"))
dat$smoking <- factor(dat$smoking, levels = c("Never", "Previous", "Current"))
dat$alcohol <- factor(dat$alcohol, levels = c("<3 times/week", "3+ times/week", "Never"))
#dat$cci_huang <- factor(dat$cci_huang, levels = c("None", "Mild to Moderate", "Severe"))
dat$overall_activity_quarters <- factor(dat$overall_activity_quarters, levels = c("Quarter 1", "Quarter 2", "Quarter 3", "Quarter 4"))
```

## Create a case control cohort 4:1 exact matching on age (by year), season wear and sex

```{r}
dat$age_10yrBand <-
  cut(
    dat$age_entry_years,
    breaks = c(40,50,60,70),
    right = FALSE
  )
```

Tidying variables for Scott steps analysis 
```{r} 
#Create a singular factor column which reads 0 - healthy controls i.e. non psoriatic disease, 1 psoriasis and 2 = PsA 

#First create non-psoriatic group 
dat$NonPsoriaticDisease <- !dat$Prevalent_PsO_lessPsA & !dat$total_prevalent_PsA 

#Now we want to create a new column called FactorAnalysis which the above 3 factors 
dat$AnalysisCategory <- ifelse(dat$NonPsoriaticDisease, 0,
                              ifelse(dat$Prevalent_PsO_lessPsA, 1,
                                     ifelse(dat$total_prevalent_PsA, 2, NA)))

#Set up a corresponding label to these categories 
labels2 <- c("NonPsoriaticDisease", "PsoriasisOnly", "PsA")

dat$AnalysisLabel <- factor(dat$AnalysisCategory, levels = 0:2, labels = labels2)
```

Generate a table 1 - baseline demographics by non-psoriatic, psoriasis and PsA 
Create column called groups which returns for each of the above 3
```{r}
dat <- dat %>%
  mutate(groups = case_when(
    NonPsoriaticDisease == TRUE ~ "NonPsoriatic",
    Prevalent_PsO_lessPsA == TRUE ~ "Psoriasis",
    total_prevalent_PsA == TRUE ~ "PsA",
    TRUE ~ NA_character_
  ))
```


We use the 'table1' package to 'cheat' and generate a nicely formatted table:
```{r}
# Add labels
label(dat$age_entry_years)   <- "Age at accelerometer wear"
label(dat$overall_activity_quarters) <- "Quarter of overall activity"
label(dat$sex) <- "Sex"
label(dat$ethnicity) <- "Ethnicity"
label(dat$tdi_quarters) <- "Quarter of Townsend Deprivation Index"
label(dat$age_education_revalued_raw) <- "Age completed full-time education"
label(dat$smoking) <- "Smoking status"
label(dat$alcohol) <- "Frequency of alcohol consumption"
label(dat$BMI) <- "BMI (continuous)"
label(dat$BMI_cats) <- "BMI (by categories)"
units(dat$age_entry_years) <- "years"
units(dat$overall_activity_quarters) <- "mg"

dat$groups <- factor(dat$groups, levels = c("NonPsoriatic", "Psoriasis", "PsA"))

# We'll customise how we render variables so rather than median (min, max) we present median (Q1, Q3)
#my_render_cont <- function(x){
 # with(
  #  stats.apply.rounding(stats.default(x)),
   # c(
    #  "",
     # `Mean (SD)` = sprintf("%s (%s)", MEAN, SD),
      #`Median [Q1, Q3]` = sprintf("%s [%s, %s]",
       #                             MEDIAN, Q1, Q3)
    #)
  #)
#}

# Make table
tab_desc <- table1::table1(~ age_entry_years + sex + ethnicity + tdi_quarters  + smoking + alcohol + BMI + BMI_cats  | groups, 
                           include_median = FALSE,
                           data = dat
                         #  ,render.cont = my_render_cont
)
tab_desc
write(tab_desc, "descriptive_table.html")
```

Report on comorbidities in HC, Psoriasis, and PsA 


Comorbidities table 
```{r}
label(dat$cci) <- "Charlson Comorbidity Index"
label(dat$major_coronary_disease_charlson) <- "Coronary artery disease"
label(dat$congestive_heart_failure_charlson_CH) <- "Congestive heart failure"
label(dat$diabetes_charlson_CH) <- "Diabetes"
label(dat$liver_disease_charlson_CH) <- "Chronic liver disease"
label(dat$cancer_charlson_CH) <- "Any cancer"
label(dat$depression) <- "Depression"
label(dat$cerebrovascular_disease_charlson_CH) <- "Cerebrovascular disease"
label(dat$copd_charlson_CH) <- "COPD"
label(dat$CCI_kim) <- "CCI by categories"
dat$CCI_kim <- factor(dat$CCI_kim, levels = c("None", "Mild to Moderate", "Severe"))


tab_descComorbid <- table1::table1(~ cci + major_coronary_disease_charlson_CH +cerebrovascular_disease_charlson_CH+ congestive_heart_failure_charlson_CH + diabetes_charlson_CH +liver_disease_charlson_CH+cancer_charlson_CH+copd_charlson_CH+depression  | groups, 
                           data = dat 
                          #  ,render.cont = my_render_cont
)
tab_descComorbid

tab_descSteps <- table1::table1(~ med_steps  | groups, 
                           data = dat 
                          #  ,render.cont = my_render_cont
)
tab_descSteps
 

write(tab_descComorbid, "descriptive_table2.html")
```


###Model for inclusion!! 
```{r}
model_steps1 <- lm(med_steps ~ AnalysisLabel + as.factor(sex)  + as.factor(age_gp) + as.factor(season_wear)+ as.factor(tdi_cats) + as.factor(alcohol) + as.factor(smoking),  data = dat)
summary(model_steps1)
  confint(model_steps1)
  
emmeans_model_steps1_np<- emmeans(model_steps1, specs = "AnalysisLabel")
emmeans_model_steps1 <- emmeans(model_steps1, specs = pairwise~ AnalysisLabel)
summary(emmeans_model_steps1_np)
summary(emmeans_model_steps1)
confint(emmeans_model_steps1)

```


```{r}
# Define the data frame for plotting
steps <- summary(emmeans_model_steps1_np, specs = "AnalysisLabel")[1:3, c(1, 2, 6, 5)]
names(steps) <- c("Quality", "Mean", "UpperCI", "LowerCI")
steps$Quality <- c("Non-Psoriatic", "Psoriasis only", "Psoriatic arthritis")
steps$Quality <- factor(steps$Quality, levels = c("Non-Psoriatic", "Psoriasis only", "Psoriatic arthritis"))

# Create the plot
steps_final <- ggplot() +
  geom_pointrange(data = steps, mapping = aes(color = factor(Quality), x = Quality, y = Mean, ymin = LowerCI, ymax = UpperCI),
                  size = 1, shape = 20) +  # Increase size
  theme_minimal(base_size = 14) +  # Increase base font size
  theme(panel.grid.minor = element_blank()) +
  theme(legend.position = "none") +
  scale_y_continuous(breaks = seq(0, 10100, by = 200), name = "Adjusted Mean Daily Step Count") +
  coord_cartesian(ylim = c(8000, 10100)) +
  xlab("") +
  geom_hline(yintercept = steps$Mean[1], linetype = "dashed", color = "black", size = 0.5) +  # Increase size
  theme(axis.title.x = element_text(vjust = -1, size = 14)) +  # Increase x-axis title font size
  theme(axis.text.y.left = element_text(size = 12)) +  # Increase y-axis text font size
  theme(axis.text.x.bottom = element_text(size = 12)) +  # Increase x-axis text font size
  theme(plot.caption = element_text(hjust = 0, size = 12))  # Increase caption font size

# Display the plot
print(steps_final)


# Close the TIFF graphics device
dev.off()
```

Final model

```{r}
model_cadencePartial<- lm(CadencePeak30Adjusted.steps.min. ~ AnalysisLabel + as.factor(age_gp)+as.factor(sex)+as.factor(season_wear)+ as.factor(tdi_cats) + as.factor(alcohol) + as.factor(smoking),  data = dat)
summary(model_cadencePartial)
  confint(model_cadencePartial)
  
emmeans_model_cadencePartial_np<- emmeans(model_cadencePartial, specs = "AnalysisLabel")
emmeans_model_cadencePartial<- emmeans(model_cadencePartial, specs = pairwise~ AnalysisLabel)
summary(emmeans_model_cadencePartial_np)
summary(emmeans_model_cadencePartial)
confint(emmeans_model_cadencePartial)
```

Final Cad model  
```{r}
cadencePartial <- summary(emmeans_model_cadencePartial_np, specs = "AnalysisLabel")[1:3,c(1,2,6,5)]
names(cadencePartial) <- c("Quality", "Mean", "UpperCI", "LowerCI")
cadencePartial$Quality <- c("Non-Psoriatic", "Psoriasis only", "Psoriatic arthritis")
cadencePartial$Quality <- factor(cadencePartial$Quality, levels = c("Non-Psoriatic", "Psoriasis only", "Psoriatic arthritis"))


cad_plotFinal <- ggplot() +
  geom_pointrange(data = cadencePartial, mapping = aes(color = factor(Quality), x = Quality, y = Mean, ymin = LowerCI, ymax = UpperCI),
                  size = 1, shape = 20) +  # Increase size
  theme_minimal(base_size = 14) +  # Increase base font size
  theme(panel.grid.minor = element_blank()) +
  theme(legend.position = "none") +
  scale_y_continuous(breaks = seq(0, 170, by = 2), name = "Adjusted Peak Cadence (Steps/in 30 mins)") +
  coord_cartesian(ylim = c(76, 96)) +
  xlab("") +
  geom_hline(yintercept = cadencePartial$Mean[1], linetype = "dashed", color = "black", size = 0.5) +  # Increase size
  theme(axis.title.x = element_text(vjust = -1, size = 14)) +  # Increase x-axis title font size
  theme(axis.text.y.left = element_text(size = 12)) +  # Increase y-axis text font size
  theme(axis.text.x.bottom = element_text(size = 12)) +  # Increase x-axis text font size
  theme(plot.caption = element_text(hjust = 0, size = 12)) +  # Increase caption font size
  theme(plot.caption = element_text(hjust = 0, size = 12))  # Increase caption font size

cad_plotFinal


``` 

```{r}
dat <- dat %>%
  mutate(HC = !total_prevalent_PsA & !Prevalent_PsO_lessPsA & !LongTermOpioid)


dat$PsALongOpioid <- with(dat, total_prevalent_PsA & LongTermOpioid)
dat$PsA_NoOpioid <- with(dat, total_prevalent_PsA & !LongTermOpioid)

dat$PsOLongOpioid  <- with(dat, Prevalent_PsO_lessPsA & LongTermOpioid)
dat$PsO_NoOpioid  <- with(dat, Prevalent_PsO_lessPsA & !LongTermOpioid)

dat <- dat %>% 
  mutate(PsoriaticLongOpioid = case_when(
    HC ~ 4,
    PsALongOpioid ~ 3,
    PsA_NoOpioid ~ 2,
    PsOLongOpioid ~ 1,
    PsO_NoOpioid ~ 0,
    TRUE ~ NA_real_  # default case if none of the above conditions are met
  ))
                                     
#Set up a corresponding label to these categories 
labels3 <- c("HC", "PsA + persistent opioid", "PsA + no opioid", "PsO + persistent opioid", "PsO + no opioid")

dat$PsoriaticLongOpioid <- factor(dat$PsoriaticLongOpioid, levels = 0:4, labels = labels3)

```

```{r}
model_steps1 <- lm(med_steps ~ PsoriaticLongOpioid + as.factor(sex)  + as.factor(age_gp) + as.factor(season_wear)+ as.factor(tdi_cats) + as.factor(alcohol) + as.factor(smoking),  data = dat)
summary(model_steps1)
  confint(model_steps1)
  
emmeans_model_steps1_np<- emmeans(model_steps1, specs = "PsoriaticLongOpioid")
emmeans_model_steps1 <- emmeans(model_steps1, specs = pairwise~ PsoriaticLongOpioid)
summary(emmeans_model_steps1_np)
summary(emmeans_model_steps1)
confint(emmeans_model_steps1)

```


```{r}
dat <- dat %>%
  mutate(HC2 = !total_prevalent_PsA & !Prevalent_PsO_lessPsA & !Systemic)


dat$PsAAcuteOpioid <- with(dat, total_prevalent_PsA & AnyOpioid)
dat$PsA_NoAcuteOpioid <- with(dat, total_prevalent_PsA & !AnyOpioid)

dat$PsOAcuteOpioid  <- with(dat, Prevalent_PsO_lessPsA & AnyOpioid)
dat$PsO_NoAcuteOpioid  <- with(dat, Prevalent_PsO_lessPsA & !AnyOpioid)

dat <- dat %>% 
  mutate(PsoriaticAcuteOpioid = case_when(
    HC2 ~ 4,
    PsAAcuteOpioid ~ 3,
    PsA_NoAcuteOpioid ~ 2,
    PsOAcuteOpioid ~ 1,
    PsO_NoAcuteOpioid ~ 0,
    TRUE ~ NA_real_  # default case if none of the above conditions are met
  ))
                                     
#Set up a corresponding label to these categories 
labels3 <- c("HC", "PsA + acute opioid", "PsA + no opioid", "PsO + acute opioid", "PsO + no opioid")

dat$PsoriaticAcuteOpioid <- factor(dat$PsoriaticAcuteOpioid, levels = 0:4, labels = labels3)

```

```{r}
model_steps1 <- lm(med_steps ~ PsoriaticAcuteOpioid + as.factor(sex)  + as.factor(age_gp) + as.factor(season_wear)+ as.factor(tdi_cats) + as.factor(alcohol) + as.factor(smoking),  data = dat)
summary(model_steps1)
  confint(model_steps1)
  
emmeans_model_steps1_np<- emmeans(model_steps1, specs = "PsoriaticAcuteOpioid")
emmeans_model_steps1 <- emmeans(model_steps1, specs = pairwise~ PsoriaticAcuteOpioid)
summary(emmeans_model_steps1_np)
summary(emmeans_model_steps1)
confint(emmeans_model_steps1)

```


datPsoriatic to look at mediating role of medications 
```{r}
datPsoriatic <- dat %>%
  filter(Prevalent_PsO_lessPsA | total_prevalent_PsA)
```

```{r}

datPsoriatic$PsAAcuteOpioid <- with(datPsoriatic, total_prevalent_PsA & LongTermOpioid)
datPsoriatic$PsA_NoAcuteOpioid <- with(datPsoriatic, total_prevalent_PsA & !LongTermOpioid)

datPsoriatic$PsOAcuteOpioid  <- with(datPsoriatic, Prevalent_PsO_lessPsA & LongTermOpioid)
datPsoriatic$PsO_NoAcuteOpioid  <- with(datPsoriatic, Prevalent_PsO_lessPsA & !LongTermOpioid)

datPsoriatic <- datPsoriatic %>% 
  mutate(PsoriaticAcuteOpioid = case_when(
    PsAAcuteOpioid ~ 3,
    PsA_NoAcuteOpioid ~ 2,
    PsOAcuteOpioid ~ 1,
    PsO_NoAcuteOpioid ~ 0,
    TRUE ~ NA_real_  # default case if none of the above conditions are met
  ))
                                     
#Set up a corresponding label to these categories 
labels3 <- c("PsA + acute opioid", "PsA + no opioid", "PsO + acute opioid", "PsO + no opioid")

datPsoriatic$PsoriaticAcuteOpioid <- factor(datPsoriatic$PsoriaticAcuteOpioid, levels = 0:3, labels = labels3)

```

```{r}
model_steps1 <- lm(med_steps ~ PsoriaticAcuteOpioid + as.factor(sex)  + as.factor(age_gp) + as.factor(season_wear)+ as.factor(tdi_cats) + as.factor(alcohol) + as.factor(smoking),  data = datPsoriatic)
summary(model_steps1)
  confint(model_steps1)
  
emmeans_model_steps1_np<- emmeans(model_steps1, specs = "PsoriaticAcuteOpioid")
emmeans_model_steps1 <- emmeans(model_steps1, specs = pairwise~ PsoriaticAcuteOpioid)
summary(emmeans_model_steps1_np)
summary(emmeans_model_steps1)
confint(emmeans_model_steps1)

```

```{r}

# Define the data frame for plotting
steps <- summary(emmeans_model_steps1_np, specs = "PsoriaticComorbid")[1:6, c(1, 2, 6, 5)]
names(steps) <- c("Quality", "Mean", "UpperCI", "LowerCI")
steps$Quality <- c("Non-\nPsoriatic","Non-Psoriatic+\ncomorbidity", "PsO", "PsO+\ncomorbidity", "PsA", "PsA+\ncomorbidity")
steps$Quality <- factor(steps$Quality, levels = c("Non-\nPsoriatic", "Non-Psoriatic+\ncomorbidity", "PsO", "PsO+\ncomorbidity", "PsA", "PsA+\ncomorbidity"))

# Create the plot
step_plotComorbid <- ggplot(steps, aes(x = Quality, y = Mean, ymin = LowerCI, ymax = UpperCI, color = factor(Quality))) +
  geom_pointrange(size = 1, shape = 20) +
  theme_minimal(base_size = 14) + 
  theme_light() +
  theme(panel.grid.minor = element_blank()) +
  theme(legend.position = "none") +
  scale_y_continuous(breaks = seq(0, 10000, by = 1000), name = "Adjusted Mean Daily Step Count") +
  coord_cartesian(ylim = c(4000, 10000)) +
  xlab("") +
  geom_hline(yintercept = steps$Mean[1], linetype = "dashed", color = "black", size = 0.5) +
  theme(axis.title.x = element_text(vjust = -1,size=14)) +
  theme(plot.caption = element_text(hjust = 0))+
  theme(axis.text.y.left = element_text(size = 12)) +  # Increase y-axis text font size
  theme(axis.text.x.bottom = element_text(size = 12)) +  # Increase x-axis text font size
  theme(plot.caption = element_text(hjust = 0, size = 12)) +  # Increase caption font size
  theme(plot.caption = element_text(hjust = 0, size = 12)) +
  theme(axis.title.y = element_text(size = 14))


# Display the plot
print(step_plotComorbid) 
```


