---
title: "Modeling_1sd"
author: "Dylan McGagh"
date: "`r format(Sys.time(), '%a %b %d, %Y %X')`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 5
    theme: lumen
---
```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction

This script runs models for 1 SD increase in exposure.

# Set up workspace
```{r}
dir.create("PRSbyGroup")
```

## Set-up useful functions and packages
```{r}
############################################
#Load packages
pkgs <- c("data.table", "rmarkdown", "plyr", "lubridate", "magrittr", "ggplot2", "data.table", "dplyr", "stringr", "splitstackshape", "scales", "haven", "Epi", "emmeans", "rtables", "table1",  "tableone", "boot", "knitr", "pspline", "survival", "xtable", "splines2", "htmltools", "here") # packages we need

pkgs_inst <- pkgs[!{pkgs %in% rownames(installed.packages())}] # check which are not present 
install.packages(pkgs_inst, repos = "https://www.stats.bris.ac.uk/R/") # install
lapply(pkgs, library, character.only = TRUE) #load all packages

## Set-up useful functions and packages
source("rounding_functions.R")
source("cut_by_quantile.R") 
```
## Load data

# Overall - SD models
Following [example code for cox model loops](https://github.com/shaankhurshid/ecg_ai/blob/main/cox_ipcw.R)

```{r}
dat <- fread("Data/prepped_steps.csv", data.table = FALSE) # this file location assumes running in same session as previous script
```

Sensitivity analysis group
```{r}
dat <- fread("Data/prepped_steps_RA2years.csv", data.table = FALSE)
```

## Additional data preprocessing

Done in this script not the previous to ensure reference values behave (easy to fix by relevelling otherwise).

```{r}
dat$step_quarters <- qtile_cut(dat$med_steps, probs = seq(0, 1, by = 0.25))
dat$acc_quarters <- qtile_cut(dat$overall_activity, probs = seq(0, 1, by = 0.25), dp_label = 1)
dat$tdi_quarters <- qtile_cut(dat$tdi_raw, probs = seq(0, 1, by = 0.25), dp_label = 1)
dat$age_gp_crude <- cut(dat$age_entry_years, seq(40, 80, by = 10), right = FALSE, labels = c("40-49", "50-59", "60-69", "70-79"))
dat$BMI_cats <-
  cut(dat$BMI,
      breaks = c(0, 25, 30, 10000),
      labels = c("<18.5-24.9", "25.0-29.9", "30.0+"),
      right = FALSE)
```

Get population level sd of steps and then subset data down to PRS risk groups 

```{r}

################Cox models
# Calc sd steps
sd(dat$med_steps) #3822.266
# Generate a standardization variable
dat$med_steps_std = (dat$med_steps - mean(dat$med_steps))/sd(dat$med_steps)
# Check distribution of variable
hist(dat$med_steps_std)

```

Subgroup analysis using Harry's method 

```{r}

library("multcomp", "foreign", "biostat3")
cox_BMI <- coxph((Surv(age_entry_days, age_exit_days, RA_incident)) ~ 
                           med_steps_std*BMI_cats+ strata(sex)+
                           as.factor(ethnicity) + as.factor(tdi_quarters) +as.factor(qualif)+
                           as.factor(season_wear) + as.factor(smoking) + as.factor(alcohol)+as.factor(processed_meat)+as.factor(fresh_fruit)+as.factor(oily_fish),
                         data=dat)


reduced_model <- coxph(Surv(age_entry_days, age_exit_days, RA_incident) ~ 
                       med_steps_std + BMI_cats + strata(sex) +
                       as.factor(ethnicity) + as.factor(tdi_quarters) +
                       as.factor(qualif) + as.factor(season_wear) + 
                       as.factor(smoking) + as.factor(alcohol) + 
                       as.factor(processed_meat) + as.factor(fresh_fruit) + 
                       as.factor(oily_fish), data = dat)

lrt <- anova(cox_BMI,reduced_model, test = "Chisq")

lrt
x_chi <- lrt$"Chisq"[13] 
p_value <- lrt$"Pr(>|Chi|)"[13] 
df <- lrt$"Df"[13]

summary(cox_BMI)


sub_BMI <- lincom(cox_BMI, c("med_steps_std",
                          "med_steps_std+med_steps_std:BMI_cats25.0-29.9",
                          "med_steps_std+med_steps_std:BMI_cats30.0+"),
       eform=TRUE)
res <- as.data.frame(sub_BMI)

sub_BMI_df <- bind_rows(res)

sub_BMI_df <- sub_BMI_df %>%
  mutate(model = c("BMI: <25", "BMI: 25 - 30", "BMI: >30")) %>%
  relocate(model, .before = 1)

sub_BMI_df <- sub_BMI_df[, c(1, 2, 3, 4, 6), drop = TRUE]

sub_BMI_df <- sub_BMI_df %>%
  rename(
    `Lower CI` = `2.5 %`,
    `Upper CI` = `97.5 %`
  )

new_row_names <- c("1", "2", "3")

# Assign the new row names to the dataframe
rownames(sub_BMI_df) <- new_row_names

sub_BMI_df$x_chi <- x_chi
sub_BMI_df$p_value <- p_value

```

```{r}
#Now get case counts and denominator for each 
counts <- dat %>%
  group_by(BMI_cats) %>%
  summarise(
    Total_Length = n(),  # Total number of observations
    RA_Incidents = sum(RA_incident, na.rm = TRUE)  # Total number of RA incidents, excluding NA values
  )

#Rename to match 
new_row_names <- c("BMI: <25", "BMI: 25 - 30", "BMI: >30")

# Assign the new row names to the dataframe
rownames(counts) <- new_row_names

counts <- counts %>%
  rename(model = BMI_cats)

counts <- counts %>%
  mutate(
    model = case_when(
      model == "<18.5-24.9" ~ "BMI: <25",
      model == "25.0-29.9" ~ "BMI: 25 - 30",
      model == "30.0+" ~ "BMI: >30")
  )

sub_BMI_df <- left_join(sub_BMI_df, counts, by = "model")

sub_BMI_df <- sub_BMI_df %>%
  mutate(
    across(2:5, ~ as.numeric(as.character(.)))
  )

sub_BMI_df <- sub_BMI_df %>%
  mutate(across(where(is.numeric), round, 3))
    
```

```{r}
cox_sex <- coxph((Surv(age_entry_days, age_exit_days, RA_incident)) ~ 
                   med_steps_std + med_steps_std:strata(sex)+as.factor(ethnicity) +
                   as.factor(tdi_quarters) +as.factor(qualif)+as.factor(season_wear) + 
                   as.factor(smoking)+as.factor(alcohol)+as.factor(processed_meat)+
                   as.factor(fresh_fruit)+as.factor(oily_fish),
                         data=dat)

reduced_model_sex <- coxph(Surv(age_entry_days, age_exit_days, RA_incident) ~ 
                       med_steps_std + strata(sex) +
                       as.factor(ethnicity) + as.factor(tdi_quarters) +
                       as.factor(qualif) + as.factor(season_wear) + 
                       as.factor(smoking) + as.factor(alcohol) + 
                       as.factor(processed_meat) + as.factor(fresh_fruit) + 
                       as.factor(oily_fish), data = dat)

lrt <- anova(cox_sex, reduced_model_sex, test = "Chisq")
lrt

#p-value on the interaction term between sex + med steps, if significnat  then suggests heterogeneity betweens males and females -> 0.118
summary(cox_sex)


subgroup_sex <- lincom(cox_sex, c("med_steps_std",
                                  "med_steps_std+med_steps_std:strata(sex)Male"),
                       eform=TRUE)
subgroup_sex


```

```{r}

cox_age <- coxph((Surv(age_entry_days, age_exit_days, RA_incident)) ~ 
                           med_steps_std + med_steps_std:age_gp_crude + strata(sex)+ as.factor(ethnicity) +
                   as.factor(tdi_quarters) +as.factor(qualif)+as.factor(season_wear) + 
                   as.factor(smoking)+as.factor(alcohol)+as.factor(processed_meat)+
                   as.factor(fresh_fruit)+as.factor(oily_fish),
                         data=dat)

cox_age_reduced <- coxph((Surv(age_entry_days, age_exit_days, RA_incident)) ~ 
                           med_steps_std + as.factor(age_gp_crude) + strata(sex)+  as.factor(ethnicity) +
                   as.factor(tdi_quarters) +as.factor(qualif)+as.factor(season_wear) + 
                   as.factor(smoking)+as.factor(alcohol)+as.factor(processed_meat)+
                   as.factor(fresh_fruit)+as.factor(oily_fish),
                         data=dat)

lrt <- anova(cox_age, cox_age_reduced, test = "Chisq")
lrt



#p-value on the interaction term between sex + med steps, if significnat  then suggests heterogeneity betweens males and females -> 0.118
summary(cox_age)


subgroup_age <- lincom(cox_age, c("med_steps_std",
                                  "med_steps_std+med_steps_std:age_gp_crude50-59",
                                  "med_steps_std+med_steps_std:age_gp_crude60-69",
                                  "med_steps_std+med_steps_std:age_gp_crude70-79"),
                                    eform=TRUE)
subgroup_age

```

```{r}

library("multcomp", "foreign", "biostat3")
cox_BMI_cad <- coxph((Surv(age_entry_days, age_exit_days, RA_incident)) ~ 
                           cad30_std*BMI_cats+ strata(sex)+
                           as.factor(ethnicity) + as.factor(tdi_quarters) +as.factor(qualif)+
                           as.factor(season_wear) + as.factor(smoking) + as.factor(alcohol)+as.factor(processed_meat)+as.factor(fresh_fruit)+as.factor(oily_fish),
                         data=dat)

summary(cox_BMI_cad)


subgroup_BMI_cad <- lincom(cox_BMI_cad, c("cad30_std",
                          "cad30_std+cad30_std:BMI_cats25.0-29.9",
                          "cad30_std+cad30_std:BMI_cats30.0+"),
       eform=TRUE)
subgroup_BMI_cad


```





Subset based on PRS risk 
```{r}
datLow <- dat %>%
  filter(PRS_RA_RiskGroup == "Low")
datInter <- dat %>%
  filter(PRS_RA_RiskGroup == "Intermediate")
datHigh <- dat %>%
  filter(PRS_RA_RiskGroup == "High")

```

Low risk group first 
```{r}
#########Steps - SD
### Define outcomes for loop

# Create an empty data frame to store the results
merged_steps_df <- data.frame()

# Loop over the variables

  ############################################# Model 1: Unadjusted
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_std,
    data = datLow
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and model
  res_dat <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Unadjusted model"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )

  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat)
  ############################################# Model 2: Fully-adjusted
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_std + strata(sex) + factor(ethnicity) + factor(tdi_quarters) + factor(season_wear)+ factor(qualif) + factor(smoking) + factor(alcohol) + factor(oily_fish)+factor(processed_meat)+factor(fresh_fruit),
    data = datLow
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and model
  res_dat2 <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Fully-adjusted"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )

  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat2)

 ############################################# Model 3: Fully adjusted + bmi
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_std + strata(sex) + factor(ethnicity) + factor(tdi_quarters) + factor(season_wear)+ factor(qualif) + factor(smoking) + factor(alcohol) + factor(oily_fish)+factor(processed_meat)+factor(fresh_fruit)+ factor(BMI_cats),
    data = datLow
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and mode
  res_dat3 <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Fully adjusted + bmi"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )
  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat3)
  merged_steps_dfLow <- merged_steps_df

  
# Print the merged data frame
# View(merged_steps_df)

#Save step sd results
write.csv(merged_steps_dfLow, file=here::here("PRSbyGroup", paste0("steps_sdmodelresultsLow.csv")), row.names=FALSE)
```


Intermediate risk group  
```{r}
#########Steps - SD
### Define outcomes for loop

# Create an empty data frame to store the results
merged_steps_df <- data.frame()

# Loop over the variables

  ############################################# Model 1: Unadjusted
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_std,
    data = datInter
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and model
  res_dat <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Unadjusted model"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )

  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat)
  ############################################# Model 2: Fully-adjusted
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_std + strata(sex) + factor(ethnicity) + factor(tdi_quarters) + factor(season_wear)+ factor(qualif) + factor(smoking) + factor(alcohol) + factor(oily_fish)+factor(processed_meat)+factor(fresh_fruit),
    data = datInter
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and model
  res_dat2 <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Fully-adjusted"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )

  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat2)

 ############################################# Model 3: Fully adjusted + bmi
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_std + strata(sex) + factor(ethnicity) + factor(tdi_quarters) + factor(season_wear)+ factor(qualif) + factor(smoking) + factor(alcohol) + factor(oily_fish)+factor(processed_meat)+factor(fresh_fruit)+ factor(BMI_cats),
    data = datInter
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and mode
  res_dat3 <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Fully adjusted + bmi"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )
  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat3)
  merged_steps_dfInter <- merged_steps_df

  
# Print the merged data frame
# View(merged_steps_df)

#Save step sd results
write.csv(merged_steps_dfInter, file=here::here("PRSbyGroup", paste0("steps_sdmodelresultsInter.csv")), row.names=FALSE)
```

High risk group  
```{r}
#########Steps - SD
### Define outcomes for loop

# Create an empty data frame to store the results
merged_steps_df <- data.frame()

# Loop over the variables

  ############################################# Model 1: Unadjusted
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_std,
    data = datHigh
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and model
  res_dat <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Unadjusted model"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )

  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat)
  ############################################# Model 2: Fully-adjusted
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_std + strata(sex) + factor(ethnicity) + factor(tdi_quarters) + factor(season_wear)+ factor(qualif) + factor(smoking) + factor(alcohol) + factor(oily_fish)+factor(processed_meat)+factor(fresh_fruit),
    data = datHigh
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and model
  res_dat2 <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Fully-adjusted"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )

  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat2)

 ############################################# Model 3: Fully adjusted + bmi
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_std + strata(sex) + factor(ethnicity) + factor(tdi_quarters) + factor(season_wear)+ factor(qualif) + factor(smoking) + factor(alcohol) + factor(oily_fish)+factor(processed_meat)+factor(fresh_fruit)+ factor(BMI_cats),
    data = datHigh
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and mode
  res_dat3 <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Fully adjusted + bmi"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )
  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat3)
  merged_steps_dfHigh <- merged_steps_df  

  
# Print the merged data frame
# View(merged_steps_df)

#Save step sd results
write.csv(merged_steps_dfHigh, file=here::here("PRSbyGroup", paste0("steps_sdmodelresultsHigh.csv")), row.names=FALSE)
```


Sensitivity analysis of those 2 years post 

Done in this script not the previous to ensure reference values behave (easy to fix by relevelling otherwise).

```{r}
dat2yr$step_fifths <- qtile_cut(dat2yr$med_steps, probs = seq(0, 1, by = 0.2))
dat2yr$acc_fifths <- qtile_cut(dat2yr$overall_activity, probs = seq(0, 1, by = 0.2), dp_label = 1)
dat2yr$tdi_quarters <- qtile_cut(dat2yr$tdi_raw, probs = seq(0, 1, by = 0.25), dp_label = 1)
dat2yr$age_gp_crude <- cut(dat2yr$age_entry_years, seq(40, 80, by = 10), right = FALSE, labels = c("40-49", "50-59", "60-69", "70-79"))
dat2yr$BMI_cats <-
  cut(dat2yr$BMI,
      breaks = c(0, 25, 30, 10000),
      labels = c("<18.5-24.9", "25.0-29.9", "30.0+"),
      right = FALSE)
```

Get population level sd of steps and then subset data down to PRS risk groups 

```{r}

################Cox models
# Calc sd steps
sd(dat2yr$med_steps) #3815.4
# Generate a standardization variable
dat2yr$med_steps_standardization = (dat2yr$med_steps - mean(dat2yr$med_steps))/sd(dat2yr$med_steps)
# Check distribution of variable
hist(dat2yr$med_steps_standardization)

```


Subset based on PRS risk 
```{r}
datLow <- dat2yr %>%
  filter(PRS_RA_RiskGroup == "Low")
datInter <- dat2yr %>%
  filter(PRS_RA_RiskGroup == "Intermediate")
datHigh <- dat2yr %>%
  filter(PRS_RA_RiskGroup == "High")

```

Low risk group first 
```{r}
#########Steps - SD
### Define outcomes for loop

# Create an empty data frame to store the results
merged_steps_df <- data.frame()

# Loop over the variables

  ############################################# Model 1: Unadjusted
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_standardization,
    data = datLow
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and model
  res_dat <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Unadjusted model"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )

  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat)
  ############################################# Model 2: Fully-adjusted
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_standardization + factor(sex) + factor(ethnicity) + factor(tdi_quarters) + factor(season_wear)+ factor(qualif) + factor(smoking) + factor(alcohol),
    data = datLow
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and model
  res_dat2 <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Fully-adjusted"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )

  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat2)

 ############################################# Model 3: Fully adjusted + bmi
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_standardization + factor(sex) + factor(ethnicity) + factor(tdi_quarters) + factor(season_wear)+ factor(qualif) + factor(smoking) + factor(alcohol) + factor(BMI_cats),
    data = datLow
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and mode
  res_dat3 <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Fully adjusted + bmi"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )
  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat3)
  merged_steps_dfLow <- merged_steps_df

  
# Print the merged data frame
# View(merged_steps_df)

#Save step sd results
write.csv(merged_steps_dfLow, file=here::here("PRSbyGroup", paste0("steps_sdmodelresultsLow2yr.csv")), row.names=FALSE)
```


Intermediate risk group  
```{r}
#########Steps - SD
### Define outcomes for loop

# Create an empty data frame to store the results
merged_steps_df <- data.frame()

# Loop over the variables

  ############################################# Model 1: Unadjusted
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_standardization,
    data = datInter
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and model
  res_dat <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Unadjusted model"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )

  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat)
  ############################################# Model 2: Fully-adjusted
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_standardization + factor(sex) + factor(ethnicity) + factor(tdi_quarters) + factor(season_wear)+ factor(qualif) + factor(smoking) + factor(alcohol),
    data = datInter
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and model
  res_dat2 <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Fully-adjusted"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )

  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat2)

 ############################################# Model 3: Fully adjusted + bmi
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_standardization + factor(sex) + factor(ethnicity) + factor(tdi_quarters) + factor(season_wear)+ factor(qualif) + factor(smoking) + factor(alcohol) + factor(BMI_cats),
    data = datInter
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and mode
  res_dat3 <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Fully adjusted + bmi"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )
  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat3)
  merged_steps_dfInter <- merged_steps_df

  
# Print the merged data frame
# View(merged_steps_df)

#Save step sd results
write.csv(merged_steps_dfInter, file=here::here("PRSbyGroup", paste0("steps_sdmodelresultsInter2yr.csv")), row.names=FALSE)
```

High risk group  
```{r}
#########Steps - SD
### Define outcomes for loop

# Create an empty data frame to store the results
merged_steps_df <- data.frame()

# Loop over the variables

  ############################################# Model 1: Unadjusted
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_standardization,
    data = datHigh
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and model
  res_dat <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Unadjusted model"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )

  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat)
  ############################################# Model 2: Fully-adjusted
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_standardization + factor(sex) + factor(ethnicity) + factor(tdi_quarters) + factor(season_wear)+ factor(qualif) + factor(smoking) + factor(alcohol),
    data = datHigh
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and model
  res_dat2 <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Fully-adjusted"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )

  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat2)

 ############################################# Model 3: Fully adjusted + bmi
  # Cox regression model ==================
  model <- coxph(
    Surv(age_entry_days, age_exit_days, RA_incident) ~ med_steps_standardization + factor(sex) + factor(ethnicity) + factor(tdi_quarters) + factor(season_wear)+ factor(qualif) + factor(smoking) + factor(alcohol) + factor(BMI_cats),
    data = datHigh
  )
  res <- summary(model)
  
  # Extraction
  aic <- AIC(model)
  rr <- res$coefficients[1:1, 1]
  uci <- exp(rr + qnorm(0.975) * res$coefficients[1:1, 3])
  lci <- exp(rr - qnorm(0.975) * res$coefficients[1:1, 3])
  p <- res$coefficients[1:1, 5]

  # Create a data frame for the current variable and mode
  res_dat3 <- data.frame(
    Outcome= c("RA_incident"),
    Model = c("Fully adjusted + bmi"),
    `HR (95% CI)` = c("REF", paste0(round(exp(rr), 2), " (", round(lci, 2), " - ", round(uci, 2), ")")),
    `P trend` = c(round(p, 4)),
    AIC = c(round(aic, 2))
  )
  # Append the current data frame to the merged data frame
  merged_steps_df <- rbind(merged_steps_df, res_dat3)
  merged_steps_dfHigh <- merged_steps_df  

  
# Print the merged data frame
# View(merged_steps_df)

#Save step sd results
write.csv(merged_steps_dfHigh, file=here::here("PRSbyGroup", paste0("steps_sdmodelresultsHigh2yr.csv")), row.names=FALSE)
```



