---
title: "3_prep_data_for_analysis"
format: html
editor: visual
---

## Continue environment setup

## Load packages

Load packages for use in subsequent scripts:

```{r}
library(data.table)
library(plyr)
library(dplyr)
library(magrittr)
library(lubridate)
library(readr)
library(survival)
library(Epi)
library(emmeans)
library(ggplot2)
##library(ckbplotr)
```

## Set-up useful functions

```{r}
source("rounding_functions.R")
source("cut_by_quantile.R")
source("table_lookup_whole.R")
```

## Create necessary folders

```{r}
dir.create("data")
dir.create("outputs")
```

```{bash}
cd ~/rap_wearables

dx download /users/mcgaghd/Data/participant_data.csv -r
dx download /users/mcgaghd/Data/hes_data.csv -r
dx download /users/mcgaghd/Data/death_data.csv -r
dx download /users/mcgaghd/Data/death_cause_data.csv -r


```

We load data:
```{r}
# Note that in order to be able to load the data from these locations, you will need to start a new Jupyter Lab session relative to the previous notebook.
# Alternatively you could replace these locations with appropriate file locations from the current session's storage, or you could try remounting (see: 
# community.dnanexus.com/s/question/0D5t000003vAWYxCAO/it-seems-that-the-recently-dx-uploaded-files-does-not-show-up-on-mntproject-until-i-restart-the-whole-jupyter-lab-vm)
dat <- fread("participant_data.csv", data.table = FALSE) # fread is a function from the data.table package for fast reading of large data
dat_hes <- fread("hes_data.csv", data.table = FALSE)
dat_death <- fread("death_data.csv", data.table = FALSE)
dat_death_cause <- fread("death_cause_data.csv", data.table = FALSE)
```



## Basic R formatting

```{r}
cols_dat <- c("eid", "sex", "year_birth", "month_birth", "ethnicity_raw",
              "ukb_assess_cent",  "date_baseline", "date_inst_1", "date_inst_2", "date_lost_followup",
              "tdi_raw", "age_education_raw", "qualif_raw", "alcohol_raw", "smoking_raw", "BMI_raw",
              "self_report_cvd_baseline", "self_report_cvd_inst_1", "self_report_cvd_inst_2", 
              "date_end_accel","conditions_at_baseline_2006","medications_at_baseline_2006", "IPAQ_group", "summed_MET_minutes", "Overall health rating | Instance 0", "self_report_disability",
              colnames(dat)[grepl("acceleration", colnames(dat))])
cols_dat_hes <- c("eid","dnx_hesin_id", "dnx_hesin_diag_id", 
                  "dateepiimp",  "ins_index", "arr_index", "level",
                  "diag_icd9", "diag_icd9_nb", "diag_icd10", "diag_icd10_nb")

dat <- dat[, cols_dat]
dat_hes <- dat_hes[, cols_dat_hes]
```


We inspect the data structure to check all columns are the types we expect:

```{r}
for (data in list(dat, dat_hes, dat_death, dat_death_cause)){
    str(data, vec.len = 0) # vec.len = 0 avoids accidentally printing data
}
``` 

Mostly this looks sensible, but there are some things to address. For example, why is "age_education_raw" formatted as character? It turns out it's because there are some special values for [age completed full time education](https://biobank.ndph.ox.ac.uk/showcase/field.cgi?id=845), which need to be removed before it can be coerced to numeric. Let's reformat it appropriately.

```{r}
dat$age_education_revalued_raw <- plyr::revalue(dat$age_education_raw, 
                                                c("Do not know" =  NA, 
                                                  "Prefer not to answer" = NA, 
                                                  "Never went to school" = 0))
dat$age_education_numeric_raw <- as.numeric(dat$age_education_revalued_raw)
```

We also do some simple formatting of date columns:
```{r}
# Tabular participant data
dat$date_lost_followup <- as.Date(dat$date_lost_followup, format = "%Y-%m-%d")
for (suffix in c("baseline", "inst_1", "inst_2")){
 dat[, paste0("date_", suffix)] <- as.Date(dat[, paste0("date_", suffix)], 
                                           format = "%Y-%m-%d") 
}

# Hospital data
dat_hes$date_hes <- as.Date(dat_hes$dateepiimp, format = "%Y-%m-%d")


# Death data
dat_death$date_death <-
  as.Date(dat_death$date_of_death, format = "%Y-%m-%d")
# A very small number of participants have duplicate records in death data (e.g. perhaps from a second death certificate after post-mortem)
# In this dataset we keep just one record per participant: they should have the same date, and we will use the death_cause dataset for any 
# other records related to death. It also only affects a very small number of participants.
dat_death <-
  dat_death[dat_death$ins_index == 0, ]
```

We'll do more involved processing later on, but this just ensures we have a sensibly coded dataset to work with.

## Find the first occurrence in hospital record data


```{r}
# 1 Self reported
dat$self_reported_PsO <- grepl("psoriasis", dat$conditions_at_baseline_2006)
dat$self_reported_PsA <- grepl("psoriatic arthropathy", dat$conditions_at_baseline_2006)
```



```{r}
# The lists of ICD codes we will consider
RA_icd10_codes <- "^M05.*$|^M06.*$|^M08.0$|^M08.2$|^M08.3$|^M08.4$"
RA_icd9_codes <- "^714.*$"

# 1 Self reported
dat$self_reported_RA <- grepl("rheumatoid arthritis", dat$conditions_at_baseline_2006)


# Then we look-up those with rheumatoid arthritis
dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, RA_icd10_codes, RA_icd9_codes,
                     "prevalent", "hes_prevalent_RA"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)


##create new total RA 
dat$total_prevalent_RA <- (dat$self_reported_RA | dat$hes_prevalent_RA)
```

```{r}
AS_icd10_codes <- "^M45$|^M46.9$|M469|M468|^M481"
AS_icd9_codes <- "^720$"

# 1 Self reported
dat$self_reported_AS <- grepl("ankylosing spondylitis", dat$conditions_at_baseline_2006)


# Then we look-up those with Ankylosing Spondylitis 
dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, AS_icd10_codes, AS_icd9_codes,
                     "prevalent", "hes_prevalent_AS"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

##create new total Ank Spond arthritis 
dat$total_prevalent_AS <- (dat$self_reported_AS | dat$hes_prevalent_AS)

duplicate_cols <- duplicated(names(dat))
dat <- dat[, !duplicate_cols]
```


Get self-report and HES prevalent data 
```{r}
PsA_icd10_codes <- "^M07[^4|^5|^6].*$|^L405$|^M090.*$"
PsA_icd9_codes <- "^6960$"
PsO_icd10_codes <- "^L40[^5].*$" 
PsO_icd9_codes <- "^6961$"


# 2 Death record
dat <- merge(
  dat,
  death_disease_lookup(dat_death_cause, dat, paste(PsA_icd10_codes),
                       "Death_PsA"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, PsA_icd10_codes, PsA_icd9_codes,
                     "prevalent", "hes_prevalent_PsA"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

# Then we look-up those with psoriasis 
dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, PsO_icd10_codes, PsO_icd9_codes,
                     "prevalent", "hes_prevalent_PsO"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

##create new total psoriasis 
dat$total_prevalent_PsO <- (dat$self_reported_PsO | dat$hes_prevalent_PsO) 

dat$total_prevalent_PsA <- (dat$self_reported_PsA | dat$hes_prevalent_PsA)



##Create new variable which is psoriasis ONLY
dat$Prevalent_PsO_lessPsA <- dat$total_prevalent_PsO & !dat$total_prevalent_PsA

```

Incident RA
```{r}
dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, RA_icd10_codes, RA_icd9_codes,
                     "incident", "hes_incident_RA"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)


dat$RA_incident <- dat$hes_incident_RA & (!dat$total_prevalent_RA)


dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, RA_icd10_codes, RA_icd9_codes,
                     "date", "RA_incident_date"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

dat$RA_incident_date[!dat$RA_incident] <- NA
dat$RA_incident_date <- as.Date(dat$RA_incident_date, format = "%Y-%m-%d")
```


Incident AS 
```{r}
dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, AS_icd10_codes, AS_icd9_codes,
                     "incident", "hes_incident_AS"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)


dat$AS_incident <- dat$hes_incident_AS & (!dat$total_prevalent_AS)


dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, AS_icd10_codes, AS_icd9_codes,
                     "date", "AS_incident_date"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

dat$AS_incident_date[!dat$AS_incident] <- NA
dat$AS_incident_date <- as.Date(dat$AS_incident_date, format = "%Y-%m-%d")
```

HES incident cases of PsA
```{r}
dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, PsA_icd10_codes, PsA_icd9_codes,
                     "incident", "hes_incident_PsA"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)


dat$PsA_incident <- dat$hes_incident_PsA & !(dat$total_prevalent_PsA)

 
dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, PsA_icd10_codes, PsA_icd9_codes,
                     "date", "PsA_incident_date"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

dat$PsA_incident_date[!dat$PsA_incident] <- NA
dat$PsA_incident_date <- as.Date(dat$PsA_incident_date, format = "%Y-%m-%d")
```

Probable PsA group - ICD-10/9 code 

Probable Incident PsA 
Sero-negative (Rheumatoid) arthritis and axial spondyloarthritis
```{r}

ProbPsA_icd10_codes <- "^M06[^3].*$|^M07[^0|^1|^2|^3].*$|^M130.*$|^M45.*$|^M46[^2|^3|^4|^5].*$|^M479.*$"
ProbPsA_icd9_codes <- "^7133$|^7149$|^7165$|^720$|^721$|^71499$"


dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, ProbPsA_icd10_codes, ProbPsA_icd9_codes,
                     "incident", "hes_incident_ProbPsA"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

dat$total_prevalent_IA <- (dat$total_prevalent_PsA | dat$total_prevalent_RA |
                             dat$total_prevalent_AS)

dat$ProbPsA_incident <- ((dat$hes_incident_ProbPsA & dat$Prevalent_PsO_lessPsA)) & (!dat$total_prevalent_IA) 
  


dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, ProbPsA_icd10_codes, ProbPsA_icd9_codes,
                     "date", "ProbPsA_incident_date"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)

dat$ProbPsA_incident_date[!dat$ProbPsA_incident] <- NA
dat$ProbPsA_incident_date <- as.Date(dat$ProbPsA_incident_date, format = "%Y-%m-%d")
```

Total PsA
```{r}
dat$PsATotal_incident <- dat$PsA_incident | dat$ProbPsA_incident

# Creating the PsATotal_incident_date column
# Initialize the column with NA dates first
dat$PsATotal_incident_date <- as.Date(NA)

# Loop through each individual
for (i in 1:nrow(dat)) {
  # Check if the individual has an incident date in either or both columns
  if (!is.na(dat$PsA_incident_date[i]) & !is.na(dat$ProbPsA_incident_date[i])) {
    # If the individual has dates in both columns, choose the earliest
    dat$PsATotal_incident_date[i] <- min(dat$PsA_incident_date[i], dat$ProbPsA_incident_date[i])
  } else if (!is.na(dat$PsA_incident_date[i])) {
    # If the individual has a date only in PsA_incident_date, use that
    dat$PsATotal_incident_date[i] <- dat$PsA_incident_date[i]
  } else if (!is.na(dat$ProbPsA_incident_date[i])) {
    # If the individual has a date only in ProbPsA_incident_date, use that
    dat$PsATotal_incident_date[i] <- dat$ProbPsA_incident_date[i]
  }
  # If there's no date in either, the date will remain NA
}
```

Seropositive RA - if ever receives a code for incident seropositive RA then remove from total_PsA 

```{r}
SeropositiveRA_icd10_codes <- "^M05.*$"
SeropositiveRA_icd9_codes <- "^E971$"

dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, SeropositiveRA_icd10_codes,SeropositiveRA_icd9_codes,
                     "incident", "hes_incident_seropositiveRA"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)


dat$SeropositiveRA_incident <- dat$hes_incident_seropositiveRA & (!dat$total_prevalent_IA) 
```

Exclusions 

Exclude prevalent osteoarthritis 
Need a few sub groupings. We will generate a total OA which is all hospital coded OA diagnoses and self-report 

```{r}
# The lists of ICD codes we will consider
OA_icd10_codes <- "^M15.*$|^M16.*$|^M17.*$|^M18.*$|^M19.*$"
OA_icd9_codes <- "^715.*$"

# 1 Self reported
dat$self_reported_OA <- grepl("osteoarthritis", dat$conditions_at_baseline_2006)


# Then we look-up those with OA
dat <- merge(
  dat,
  hes_disease_lookup(dat_hes, dat, OA_icd10_codes, OA_icd9_codes,
                     "prevalent", "hes_prevalent_OA"),
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)


##create new total RA 
dat$total_prevalent_OA <- dat$self_reported_OA | dat$hes_prevalent_OA
```

## Variables

### Age

Age at accelerometer wear:

```{r}
# Add date of birth
dat$approx_dob <-
  as.Date(paste(dat$year_birth, dat$month_birth, "15", sep = "-"),
          "%Y-%B-%d") # UK Biobank doesn't contain day of birth as it would be unnecessary identifying information, so we roughly impute it as the 15th of the birth month.

# Add age at entry in days
dat$age_entry_days <-
  difftime(dat$date_baseline,
           dat$approx_dob,
           units = "days")

# Convert to age at entry in years
dat$age_entry_years <- as.double(dat$age_entry_days)/365.25
```

### Sex

Male, female

Should need no additional prep

### Age

40-44 $$note this is really 43-44$$; 45-49; 75-79

```{r}
# Add age groups
dat$age_gp <-
  cut(
    dat$age_entry_years,
    breaks = c(40,45, 50,55, 60,65, 70,75, 80),
    right = FALSE
  )
```

### Ethnicity

White, non-white

```{r}
# Ethnicity
dat$ethnicity <-
  plyr::revalue(
    dat$ethnicity_raw,
    c(
      "British" = "White",
      "Any other white background" = "White",
      "Irish" = "White",
      "White and Asian" = "Nonwhite",
      "Caribbean" = "Nonwhite",
      "Chinese"   = "Nonwhite",
      "Pakistani"  = "Nonwhite",
      "White and Black African" = "Nonwhite",
      "Other ethnic group"  = "Nonwhite",
      "Any other mixed background" = "Nonwhite",
      "African"    = "Nonwhite",
      "White and Black Caribbean" = "Nonwhite",
      "Prefer not to answer" = NA,
      "Indian"  = "Nonwhite",
      "White" = "White",
      "Do not know" = NA,
      "Any other Black background" = "Nonwhite",
      "Any other Asian background"  = "Nonwhite",
      "Bangladeshi"  = "Nonwhite",
      "Mixed"  = "Nonwhite",
      "Asian or Asian British"  = "Nonwhite",
      "Black or Black British"  = "Nonwhite"
    )
  )
```

### Education

School leaver, further education, higher education

```{r}
dat$qualif <- NA
dat$qualif[grepl("degree", dat$qualif_raw)] <-
  "Higher education"
dat$qualif[is.na(dat$qualif) & grepl("A level|NVQ|professional", dat$qualif_raw)] <- "Further education"
dat$qualif[is.na(dat$qualif) & grepl("GCSEs|CSEs|None", dat$qualif_raw)] <- "School leaver"
```


##Medication indicator -> systemic treatment in PsO

Medications
```{r}
dat_baselineMeds <- fread("Data/medication_wacc_data.csv", data.table = FALSE)
```

Extract medications of interest
```{r}
dat_baselineMeds <- dat_baselineMeds %>%
  select(eid,medications_at_baseline_2006)
```

Methotrexate
```{r}
dat_baselineMeds <- dat_baselineMeds %>%
        mutate(methotrexate = grepl("methotrexate|metoject", medications_at_baseline_2006, 
                                ignore.case =TRUE))
```


Ciclosporin
```{r}
dat_baselineMeds <- dat_baselineMeds %>%
        mutate(ciclosporin = grepl("Ciclosporin|Cyclosporin|neoral",
                                   medications_at_baseline_2006, 
                                ignore.case =TRUE))
```

Acitretin
```{r}
dat_baselineMeds <- dat_baselineMeds %>%
        mutate(acitretin = grepl("acitretin|Neotigason",
                                   medications_at_baseline_2006, 
                                ignore.case =TRUE))
```


bDMARD  
```{r}
dat_baselineMeds <- dat_baselineMeds %>%
        mutate(bDMARD = grepl("humira|adalimumab|Etanercept|efalizumab|
        Infliximab|golimumab|Ustekinumab",
                                     medications_at_baseline_2006, 
                                ignore.case =TRUE))
```

Join this up with dat
```{r}
dat <- dat %>%
  full_join(dat_baselineMeds, by = "eid")
```

Link in GP medications 

```{bash}
dx download /users/mcgaghd/PsAfinal/medications/GP -r
```

link in methotrexate data 
Link dat with medication data 

#Subset based on eid and date
```{r}
dat_accel <- dat %>%
  select(eid, date_end_accel)
```

Read in medication data 
#"Data/GP_scripts/methotrexate.csv" usually use this 
```{r}
methotrexate <- fread("GP/methotrexate.csv", data.table = FALSE) # fread is a function
bDMARD <- fread("GP/bDMARD.csv", data.table = FALSE) # fread is a function
ciclosporin <- fread("GP/ciclosporin.csv", data.table = FALSE) # fread is a function
acitretin <- fread("GP/acitretin.csv", data.table = FALSE) # fread is a function
```

Join
```{r}
joined_data <- methotrexate %>%
  left_join(dat_accel, by = "eid")
```

```{r}
filtered_data <- joined_data %>%
  filter(issue_date <= date_end_accel | is.na(date_end_accel))
```

```{r}
filtered_data$GP_MTX <- TRUE
```

Only keep unique EID
```{r}
filtered_data <- distinct(filtered_data, eid, .keep_all = TRUE)
```

```{r}
filtered_data <- select(filtered_data, eid, GP_MTX)
```

Join up with dat
```{r}
dat <- dat %>%
  left_join(filtered_data, by = "eid")
```

Replace NA with FALSE
```{r}
dat <- dat %>%
  mutate(GP_MTX = ifelse(is.na(GP_MTX), FALSE, GP_MTX))
```

Ciclosporin join up 
Join
```{r}
joined_data <- ciclosporin %>%
  left_join(dat_accel, by = "eid")
```

```{r}
filtered_data <- joined_data %>%
  filter(issue_date <= date_end_accel | is.na(date_end_accel))
```

```{r}
filtered_data$ciclosporin_GP <- TRUE
```

Only keep unique EID
```{r}
filtered_data <- distinct(filtered_data, eid, .keep_all = TRUE)
```

```{r}
filtered_data <- select(filtered_data, eid, ciclosporin_GP)
```

Join up with dat
```{r}
dat <- dat %>%
  left_join(filtered_data, by = "eid")
```

Replace NA with FALSE
```{r}
dat <- dat %>%
  mutate(ciclosporin_GP = ifelse(is.na(ciclosporin_GP), FALSE, ciclosporin_GP))
```




Acitretin
Join
```{r}
joined_data <- acitretin %>%
  left_join(dat_accel, by = "eid")
```

```{r}
filtered_data <- joined_data %>%
  filter(issue_date <= date_end_accel | is.na(date_end_accel))
```

```{r}
filtered_data$acitretin_GP <- TRUE
```

Only keep unique EID
```{r}
filtered_data <- distinct(filtered_data, eid, .keep_all = TRUE)
```

```{r}
filtered_data <- select(filtered_data, eid, acitretin_GP)
```

Join up with dat
```{r}
dat <- dat %>%
  left_join(filtered_data, by = "eid")
```

Replace NA with FALSE
```{r}
dat <- dat %>%
  mutate(acitretin_GP = ifelse(is.na(acitretin_GP), FALSE, acitretin_GP))
```

###bDMARD_GP
Join
```{r}
joined_data <- bDMARD %>%
  left_join(dat_accel, by = "eid")
```

```{r}
filtered_data <- joined_data %>%
  filter(issue_date <= date_end_accel | is.na(date_end_accel))
```

```{r}
filtered_data$bDMARD_GP <- TRUE
```

Only keep unique EID
```{r}
filtered_data <- distinct(filtered_data, eid, .keep_all = TRUE)
```

```{r}
filtered_data <- select(filtered_data, eid, bDMARD_GP)
```

Join up with dat
```{r}
dat <- dat %>%
  left_join(filtered_data, by = "eid")
```

Replace NA with FALSE
```{r}
dat <- dat %>%
  mutate(bDMARD_GP = ifelse(is.na(bDMARD_GP), FALSE, bDMARD_GP))
```

###Total indicators 
```{r}
dat$MTX_any <- (dat$MTXmonitoring_gp | dat$GP_MTX |
                             dat$methotrexate)

dat$bDMARD_any <- (dat$bDMARD_GP | dat$bDMARD)

dat$ciclosporin_any <- (dat$ciclosporin_GP | dat$ciclosporin)

dat$acitretin_any <- (dat$acitretin_GP | dat$acitretin)

dat$SystemicTreatment <- (dat$MTX_any | dat$bDMARD_any | dat$ciclosporin_any |
                            dat$acitretin_any)
```


Rheumatoid factor tidying - taken from McQueenie and Mair PMID: 33234629
```{r}
dat <- dat %>%
  mutate(RF_binary = case_when(
    RheumFactor_status == "Not reportable at assay (too high)" ~ "Positive",
    RheumFactor_status == "Not reportable at assay (too low)" ~ "Negative",
    RheumFactor_baseline > 20 ~ "Positive",
    RheumFactor_baseline >= 10 & RheumFactor_baseline <= 20 ~ "Negative",
    TRUE ~ NA_character_  # Set to NA for other cases
  ))

```

Also reformat to get a continuous variable for RF 
```{r}
dat <- dat %>%
  mutate(RF_cont = case_when(
    RheumFactor_status == "Not reportable at assay (too low)" ~ 0,
    RheumFactor_status == "Not reportable at assay (too high)" ~ 120,
    TRUE ~ RheumFactor_baseline))
#6084 NAs
```


### BMI

```{r}
# BMI
dat$BMI <- dat$BMI_raw

dat$BMI_cats <-
  cut(dat$BMI,
      breaks = c(0, 25, 30, 10000),
      labels = c("<25", "25.0-29.9", "30.0+"),
      right = FALSE)
```

### Smoking status

Never, Former, Current

```{r}
# Smoking
dat$smoking <-
  plyr::revalue(dat$smoking_raw, replace = c("Prefer not to answer" = NA))
```

### Alcohol consumption

```{r}
# Alcohol
dat$alcohol <-
  plyr::revalue(
    dat$alcohol_raw,
    replace = c(
      "Prefer not to answer" = NA,
      "Three or four times a week" = "3+ times/week",
      "Special occasions only" = "<3 times/week",
      "One to three times a month" = "<3 times/week",
      "Daily or almost daily" = "3+ times/week",
      "Once or twice a week" = "<3 times/week"
    )
  )
```

### TDI

By quarter in population, based on UK census

```{r}
dat$tdi_cats <- cut(
    dat$tdi_raw,
    breaks = c(-6.3627, -2.4167, -0.4373, 1.7863, 4.7426, 13.5881),
    labels = c("Least Deprived", "2nd Quintile", "3rd Quintile", "4th Quintile", "Most Deprived"),
    right = FALSE
  )
```


### Fresh fruit

```{r}
dat$fresh_fruit_numeric <-
  plyr::revalue(
    dat$fresh_fruit_intake,
    replace = c(
      "Less than one" = "0.5",
      "Do not know" = NA,
      "Prefer not to answer" = NA
    )
  )
dat$fresh_fruit <-
  cut(as.double(dat$fresh_fruit_numeric),
      c(0, 1.999, 2.999, 3.999, 100000),
      right = FALSE)
```

#Salad Raw Veg
```{r}
dat$salad_raw_vegetable_numeric <-
  plyr::revalue(
    dat$salad_raw_vegetable_intake,
    replace = c(
      "Less than one" = "0.5",
      "Do not know" = NA,
      "Prefer not to answer" = NA
    )
  )
dat$salad_raw_veg <-
  cut(as.double(dat$salad_raw_vegetable_numeric),
      c(0, 1.999, 2.999, 3.999, 100000),
      right = FALSE)
```

##Cooked veg
```{r}
dat$cooked_vegetable_numeric <-
  plyr::revalue(
    dat$cooked_vegetable_intake,
    replace = c(
      "Less than one" = "0.5",
      "Do not know" = NA,
      "Prefer not to answer" = NA
    )
  )
dat$cooked_veg <-
  cut(as.double(dat$cooked_vegetable_numeric),
      c(0, 1.999, 2.999, 3.999, 100000),
      right = FALSE)
```

###FruitAndVeg

```{r}
#convert to numeric
dat <- dat %>%
  mutate(across(c(salad_raw_vegetable_numeric, fresh_fruit_numeric, cooked_vegetable_numeric),
                ~as.numeric(.), .names = "numeric_{.col}")) %>%
  mutate(fruit_veg_numeric = rowSums(select(., numeric_salad_raw_vegetable_numeric, numeric_fresh_fruit_numeric, numeric_cooked_vegetable_numeric), na.rm = TRUE))

# Replace fruit_veg_numeric with NA where all three columns are NA
dat <- dat %>%
  mutate(fruit_veg_numeric = ifelse(rowSums(is.na(select(., numeric_salad_raw_vegetable_numeric, numeric_fresh_fruit_numeric, numeric_cooked_vegetable_numeric))) == 3, NA_real_, fruit_veg_numeric))

```

##Binned fruit_veg intake 
```{r}
dat$fruit_veg_total <-
  cut(as.double(dat$fruit_veg_numeric),
      c(0, 3.999, 6.999, 10.999, 100000),
      right = FALSE)
```

### Processed meat

```{r}
dat$processed_meat <-
  plyr::revalue(
    dat$processed_meat_intake,
    replace = c(
      "Do not know" = NA,
      "Prefer not to answer" = NA,
      "Less than once a week" = "Less than twice a week",
      "Once a week" = "Less than twice a week",
      "5-6 times a week" = "5 or more times a week",
      "Once or more daily" = "5 or more times a week"
    )
  )
```

```{r}
dat$processed_meat <-
  plyr::revalue(
    dat$processed_meat_intake,
    replace = c(
      "Do not know" = NA,
      "Prefer not to answer" = NA,
      "Never" = "0",
      "Less than once a week" = "0.5",
      "Once a week" = "1",
      "2-4 times a week" = "3",
      "5-6 times a week" = "5.5",
      "Once or more daily" = "7"
    )
  )

dat$processed_meat <- ifelse(dat$processed_meat == "", NA, dat$processed_meat)

```

##Beef intake 
```{r}
dat$beef_meat <-
  plyr::revalue(
    dat$beef_intake,
    replace = c(
      "Do not know" = NA,
      "Prefer not to answer" = NA,
      "Never" = "0",
      "Less than once a week" = "0.5",
      "Once a week" = "1",
      "2-4 times a week" = "3",
      "5-6 times a week" = "5.5",
      "Once or more daily" = "7"
    )
  )

dat$beef_meat <- ifelse(dat$beef_meat == "", NA, dat$beef_meat)

```

##MuttonLamb 
```{r}
dat$lamb_mutton_meat <-
  plyr::revalue(
    dat$lamb_mutton_intake,
    replace = c(
      "Do not know" = NA,
      "Prefer not to answer" = NA,
      "Never" = "0",
      "Less than once a week" = "0.5",
      "Once a week" = "1",
      "2-4 times a week" = "3",
      "5-6 times a week" = "5.5",
      "Once or more daily" = "7"
    )
  )

dat$lamb_mutton_meat <- ifelse(dat$lamb_mutton_meat == "", NA, dat$lamb_mutton_meat)

```

##Pork 
```{r}
dat$pork_meat <-
  plyr::revalue(
    dat$pork_intake,
    replace = c(
      "Do not know" = NA,
      "Prefer not to answer" = NA,
      "Never" = "0",
      "Less than once a week" = "0.5",
      "Once a week" = "1",
      "2-4 times a week" = "3",
      "5-6 times a week" = "5.5",
      "Once or more daily" = "7"
    )
  )

dat$pork_meat <- ifelse(dat$pork_meat == "", NA, dat$pork_meat)

```

##Sum red and processed meat 
```{r}
#convert to numeric
dat <- dat %>%
  mutate(across(c(pork_meat, lamb_mutton_meat, beef_meat, processed_meat),
                ~as.numeric(.), .names = "numeric_{.col}")) %>%
  mutate(red_processed_numeric = rowSums(select(., numeric_pork_meat, numeric_lamb_mutton_meat, numeric_beef_meat, numeric_processed_meat), na.rm = TRUE))

# Replace fruit_veg_numeric with NA where all three columns are NA
dat <- dat %>%
  mutate(red_processed_numeric = ifelse(rowSums(is.na(select(., numeric_pork_meat, numeric_lamb_mutton_meat, numeric_beef_meat, numeric_processed_meat))) == 4, NA_real_, red_processed_numeric))

```

##Condition that if participant answered less than once a week for all meats then fall into lowest category in binning 
```{r}
dat <- dat %>%
  mutate(red_processed_numeric = ifelse(numeric_processed_meat == 0.5 & numeric_beef_meat == 0.5 & numeric_lamb_mutton_meat == 0.5 & numeric_pork_meat == 0.5, 
                                        1.5, 
                                        rowSums(select(., numeric_processed_meat, numeric_beef_meat, numeric_lamb_mutton_meat, numeric_pork_meat), na.rm = TRUE)))
```

##Binned red_processed_meat
```{r}
dat$red_processed_total <-
  cut(as.double(dat$red_processed_numeric),
      c(0, 1.999, 3.999, 5.999, 100000),
      right = FALSE)
```



### Oily fish

```{r}
dat$oily_fish <- plyr::revalue(
  dat$oily_fish_intake,
  replace = c(
    "Do not know" = NA,
    "Prefer not to answer" = NA,
    "Less than once a week" = "Less than twice a week",
    "Once a week" = "Less than twice a week",
    "5-6 times a week" = "5 or more times a week",
    "Once or more daily" = "5 or more times a week"
  )
) 
```

### Wear season

Spring, Summer, Autumn, Winter

```{r}
dat$month_wear <- month(dat$date_end_accel)
dat$season_wear <- plyr::mapvalues(dat$month_wear,
                                   c(12, 1:11),
                                   c(
                                     rep("Winter", 3),
                                     rep("Spring", 3),
                                     rep("Summer", 3),
                                     rep("Autumn", 3)
                                   ))
table(dat$month_wear, dat$season_wear) # showing Dec-Feb assigned to winter (based on end time of accelerometer wear) and so on
```
## Add outcome

Set up censoring dates:

```{r}
ind_wales <-
  dat$ukb_assess_cent %in% c("Cardiff", "Wrexham", "Swansea")
ind_scotland <- 
  dat$ukb_assess_cent %in% c("Edinburgh", "Glasgow")

dat$date_cens <- "2022-10-31"
dat$date_cens[ind_scotland] <- "2022-08-31"
dat$date_cens[ind_wales] <- "2022-05-31"
dat$date_cens <- as.Date(dat$date_cens)
```

\[Note: if there is a new data release you can update these. But beware to:

-   update the outcomes in our project - even if there's been a new release, they won't update in our project unless someone triggers it.
-   rerun all dataset generation code, including the lower level script
-   sense check the results: do they end in the expected month?
-   check the censoring dates by region are correctly entered\]

Participants with a recorded loss-to-follow-up date should be censored at loss-to-follow-up:

```{r}
# People who were lost to follow-up are censored at earliest of loss-to-follow-up and overall censoring
dat$date_cens <- pmin(dat$date_cens, dat$date_lost_followup, na.rm = TRUE)
```

```{r}
dat <- merge(
  dat,
  dat_death,
  by = "eid",
  all.x = TRUE,
  suffixes = c("", "dup")
)
```

Participants who died should be censored at death, provided this occurred before the end of records:

```{r}
# People who died are followed up to earliest of date of death and overall censoring
dat$date_fu <- dat$date_cens
dat$date_fu[!is.na(dat$date_death)] <- 
  pmin(dat$date_cens[!is.na(dat$date_death)], dat$date_death[!is.na(dat$date_death)])
```

```{r}
# Incident cases are followed up to date of diagnosis

dat$date_fu[dat$PsATotal_incident] <- dat$PsATotal_incident_date[dat$PsATotal_incident]
```

We calculate follow up time (i.e. total time on study):
```{r}
dat$fu_time <-as.double(difftime(dat$date_fu, dat$date_baseline, units = "days"))
```

We want to analyse the data using age as the timescale, so we add a variable for age at exit in days:
```{r}
dat$age_exit_days <- as.double(dat$age_entry_days + dat$fu_time)
dat$age_exit_days2 <-  as.double(difftime(dat$date_fu, dat$approx_dob, units = "days")) # calculation in an alternative way just so we can implement a logic check

# Logic check 
if (!isTRUE(all.equal(dat$age_exit_days, dat$age_exit_days2))){
    stop("Different methods of calculating age at exit give different answers")
}
```
 

## Writing out the data for reuse

As previously, we need to write out the data so we can reuse it, and upload it to the RAP system.

```{r}
write.csv(dat, "prepped_steps_before_exclude.csv", row.names = FALSE)
```


## Now create dataset which excludes those with poor wear time etc 


We will record how many participants are excluded at each of the steps, for a flow diagram:
## Only run step below if new working environment
```{r}
dat <- fread("prepped_steps_before_exclude.csv", data.table = FALSE)
```

```{r}
tab_exc <- data.frame("Exclusion" = "Starting cohort", "Number_excluded" = NA, "Number_remaining" = nrow(dat))
```



Missing data in adjustment variables:
```{r}
for (
  cov in c(
    "age_entry_years",
    "sex",
    "BMI",
   # "ethnicity", (excluding as <2% non-white ethnicity group with psoriasis)
    "tdi_raw",
    "qualif",
    "smoking",
    "alcohol", 
   "summed_MET_minutes"
    #"red_processed_total",
   # "fruit_veg_total"
  )
){
  nb <- nrow(dat)
  print(cov)
  missing_cov <- is.na(dat[, cov])|(as.character(dat[, cov]) == "") |(as.character(dat[, cov]) == "Missing") # for safety coerce to character for second check as can return NA on some classes e.g. Date
  dat <- dat[!missing_cov,]
  tab_exc <- rbind(
    tab_exc,
    data.frame(
      "Exclusion" = paste0("Missing ", cov),
      "Number_excluded" = nb - nrow(dat),
      "Number_remaining" = nrow(dat)
    )
  )
}
```

Remove anyone with prevalent IA 
```{r}
nb <- nrow(dat)
dat <- dat[dat$total_prevalent_IA == FALSE, ] 

tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Prevalent IA at time of accelerometer wear",
      "Number_excluded" = nb - nrow(dat),
      "Number_remaining" = nrow(dat)
    )
  )
```



We will also exclude those with PsA diagnosis up to 2 years and 4 years after accelerometer wear for a sensitivity analysis

```{r}
nb <- nrow(dat)

TwoYear_cases <- (dat$PsATotal_incident & dat$fu_time <= 730)

dat_TwoYear_PsA <- dat[!(TwoYear_cases), ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Sens: Prevalent PsA disease up to 2 years after accelerometer wear",
      "Number_excluded" = nb - nrow(dat_TwoYear_PsA),
      "Number_remaining" = nrow(dat_TwoYear_PsA)
    )
  )
```

Four years
```{r}
nb <- nrow(dat)

FourYear_cases <- (dat$PsA_incident & dat$fu_time <= 1461)

dat_FourYear_PsA <- dat[!(FourYear_cases), ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Sens: Prevalent PsA disease up to 4 years after accelerometer wear",
      "Number_excluded" = nb - nrow(dat_FourYear_PsA),
      "Number_remaining" = nrow(dat_FourYear_PsA)
    )
  )
```



## Write out
```{r}
write.csv(dat, "prepped_steps_PsAfinal.csv")
```


```{r}
write.csv(dat_TwoYear_PsA, "prepped_steps_PsA2years.csv")
```

```{r}
print(tab_exc)
```

```{r}
write.csv(tab_exc, "tab_exc.csv")
```

In terminal, run the following code: `dx upload prepped_steps_PsAfinal.csv --dest users/mcgaghd/PsAfinal/prepped_steps_PsAfinal.csv`

`dx upload prepped_steps_RA2years.csv --dest users/mcgaghd/prepped_steps_RA2years.csv``dx upload prepped_steps_RA2years.csv --dest users/mcgaghd/prepped_steps_RA2years.csv`

## Clear up some of the mess ahead of running future scripts

Not strictly necessary but hopefully avoids accidentally relying on leftover data in later scripts.

```{r}
rm(list = setdiff(ls(), lsf.str())) # this setdiff is listing everything then listing only functions. So it's saying remove everything that's not a function (see https://stackoverflow.com/questions/8305754/remove-all-variables-except-functions) 
```



```{r}
dat$MVPA <- dat$MVPA_overall_average*1440
dat$LIPA <- dat$light_overall_average*1440
dat$SB <- dat$sedentary_overall_average*1440
dat$sleep <- dat$sleep_overall_average*1440
```

```{r}
datPsO <- dat %>%
  filter(total_prevalent_PsO == TRUE)
```

```{r}
TwoYear_cases <- (datPsO$PsATotal_incident & datPsO$fu_time <= 730)

datPso_2year <- datPsO[!(TwoYear_cases), ]
```

```{r}
dat$age_gp_crude <- cut(dat$age_entry_years, seq(40, 80, by = 10), right = FALSE, labels = c("40-49", "50-59", "60-69", "70-79"))
dat$age_2cat <- cut(dat$age_entry_years, seq(40, 80, by = 20), right = FALSE, labels = c("40-59", "60-79"))
dat$BMI_cats <-
  cut(dat$BMI,
      breaks = c(0, 25, 30, 10000),
      labels = c("<24.9", "25.0-29.9", "30.0+"),
      right = FALSE)
dat$tdi_quarters <- qtile_cut(dat$tdi_raw, probs = seq(0, 1, by = 0.25), dp_label = 1)

dat$MET_thirds <- qtile_cut(dat$summed_MET_minutes, probs = seq(0, 1, by = 1/3), dp_label = 1)
```

```{r}
dat <- dat %>%
  filter(total_prevalent_PsO == TRUE)
```

```{r}
dat_2year <- dat_TwoYear_PsA %>%
  filter(total_prevalent_PsO == TRUE)
```

```{r}
dat_4year <- dat_FourYear_PsA %>%
  filter(total_prevalent_PsO == TRUE)
```

```{r}
dat_4year$age_gp_crude <- cut(dat_4year$age_entry_years, seq(40, 80, by = 10), right = FALSE, labels = c("40-49", "50-59", "60-69", "70-79"))
dat_4year$BMI_cats <-
  cut(dat_4year$BMI,
      breaks = c(0, 25, 30, 10000),
      labels = c("<24.9", "25.0-29.9", "30.0+"),
      right = FALSE)
dat_4year$tdi_quarters <- qtile_cut(dat_4year$tdi_raw, probs = seq(0, 1, by = 0.25), dp_label = 1)

dat_4year$MET_thirds <- qtile_cut(dat_4year$summed_MET_minutes, probs = seq(0, 1, by = 1/3), dp_label = 1)
```


```{r}
final <- coxph((Surv(age_entry_days, age_exit_days, PsATotal_incident)) ~ 
                           (IPAQ_group) +as.factor(qualif) + as.factor(smoking) +as.factor(tdi_quarters) +as.factor(alcohol)+as.factor(BMI_cats),
                          data=dat)

summary(final)

```



```{r}
final <- coxph((Surv(age_entry_days, age_exit_days, PsATotal_incident)) ~ 
                           (IPAQ_group) +as.factor(qualif) + as.factor(smoking) +as.factor(tdi_quarters) +as.factor(alcohol)+as.factor(BMI_cats),
                          data=dat_4year)

summary(final)

```

```{r}
final <- coxph((Surv(age_entry_days, age_exit_days, PsATotal_incident)) ~ 
                           (IPAQ_group) +as.factor(qualif) + as.factor(smoking) +as.factor(tdi_quarters) +as.factor(alcohol)+as.factor(BMI_cats),
                          data=dat)

summary(final)

```



```{r}
cox_BMI_cad <- coxph((Surv(age_entry_days, age_exit_days, PsATotal_incident)) ~ 
                           (med_steps_std) +as.factor(qualif) + as.factor(smoking) +as.factor(alcohol)+as.factor(BMI_cats),
                          data=datPsO)

summary(cox_BMI_cad)

```

```{r}
dat$step_thirds <- qtile_cut(dat$med_steps, probs = seq(0, 1, by = 1/3))
dat$age_gp_crude <- cut(dat$age_entry_years, seq(40, 80, by = 10), right = FALSE, labels = c("40-49", "50-59", "60-69", "70-79"))
dat$BMI_cats <-
  cut(dat$BMI,
      breaks = c(0, 25, 30, 10000),
      labels = c("<24.9", "25.0-29.9", "30.0+"),
      right = FALSE)
```

```{r}
sd(dat$med_steps) #3822.266
# Generate a standardization variable
dat$med_steps_std = (dat$med_steps - mean(dat$med_steps))/sd(dat$med_steps)
# Check distribution of variable
hist(dat$med_steps_std)
```

```{r}
datPsO <- dat[dat$total_prevalent_PsO == TRUE, ]
```

```{r}
cox_BMI_cad <- coxph((Surv(age_entry_days, age_exit_days, PsATotal_incident)) ~ 
                           med_steps_std+
                           as.factor(smoking),
                          data=datPsO)

summary(cox_BMI_cad)

```
