---
title: "PsA_self_report_PA_analysis"
output: html_document
---


## Load data
```{r}
library(data.table)
library(plyr)
library(dplyr)
library(magrittr) 
library(lubridate)
library(readr) 
library(survival)
library(Epi)
#library(emmeans)
library(ggplot2)
#library(survminer)
#library(ckbplotr)
```

## Set-up useful functions

```{r}
source("rounding_functions.R")
source("cut_by_quantile.R")
source("cox_functions.R")
```

```{r}
dat <- fread("PsAfinal/self_report_PsA_final.csv", data.table = FALSE)


dat$V1 <- NULL
```


```{r}
dat$age_gp_crude <- cut(dat$age_entry_years, seq(40, 80, by = 10), right = FALSE, labels = c("40-49", "50-59", "60-69", "70-79"))

dat$BMI_cats <-
  cut(dat$BMI,
      breaks = c(0, 25, 30, 10000),
      labels = c("<24.9", "25.0-29.9", "30.0+"),
      right = FALSE)
dat$tdi_quarters <- qtile_cut(dat$tdi_raw, probs = seq(0, 1, by = 0.25), dp_label = 1)

dat$MET_tertiles <- qtile_cut(dat$summed_MET_minutes, probs = seq(0, 1, by = 1/3), dp_label = 1)

dat$MET_quartiles <- qtile_cut(dat$summed_MET_minutes, probs = seq(0, 1, by = 1/4), dp_label = 1)

dat$MET_quintiles <- qtile_cut(dat$summed_MET_minutes, probs = seq(0, 1, by = 1/5), dp_label = 1)


dat <- dat %>%
  mutate(MET_quartiles = factor(MET_quartiles,
                                levels = c("<756.0", "756.0-1,755.9", "1,756.0-3,491.9", "3,492.0+")))

```

2 and 4 year sensitivity analysis 
```{r}
dat2year_cases <- (dat$ProbDefPsA_incident & dat$fu_time_2 <= 760)
dat4year_cases <- (dat$ProbDefPsA_incident & dat$fu_time_2 <= 1461)
dat6year_cases <- (dat$ProbDefPsA_incident & dat$fu_time_2 <= 2192)

dat2year <- dat[!(dat2year_cases), ]
dat4year <- dat[!(dat4year_cases), ]
dat6year <- dat[!(dat6year_cases), ]
```


Get median follow up time and age of the cohort at accelerometer wear. 
```{r}
# Calculate the median of age_entry_years

dat$age_entry_years_numeric <- as.numeric(as.character(dat$age_entry_years))

median_age <- median(dat$age_entry_years_numeric, na.rm = TRUE)

# Calculate the interquartile range of age_entry_years
iqr_age <- IQR(dat$age_entry_years_numeric, na.rm = TRUE)

# Display the results
cat("Median of age_entry_years:", median_age, "\n")
cat("Interquartile Range (IQR) of age_entry_years:", iqr_age, "\n")

median_age <- 58.109
iqr_age <- 13.136

# Calculate the first and third quartiles
Q1 <- median_age - (iqr_age / 2)
Q3 <- median_age + (iqr_age / 2)

# Format the output string
output <- sprintf("%.1f (IQR: %.1f–%.1f) years", median_age, Q1, Q3)

# Print the formatted output
print(output)
```


```{r}
# Convert fu_time from days to years
dat$fu_time_years <- dat$fu_time_2 / 365.25

# Calculate the mean of fu_time in years
mean_fu_time_years <- mean(dat$fu_time_years, na.rm = TRUE)

# Calculate the standard deviation of fu_time in years
sd_fu_time_years <- sd(dat$fu_time_years, na.rm = TRUE)

# Format and print the results
output <- sprintf("Mean (SD) years: %.2f (%.2f)", mean_fu_time_years, sd_fu_time_years)
print(output)

```

Full analysis
```{r}
exposures <- c("MET_tertiles", "MET_quartiles", "MET_quintiles")
outcomes <- c("ProbDefPsA_incident") 

multivar_covs <-
c(
"sex",
"ethnicity",
strata("tdi_quarters"), 
"qualif",
"smoking",
"alcohol"
)

adjustments <-
list(
"model1" = NULL,
"model2" = c(multivar_covs),
"model3" = c(multivar_covs, "BMI_cats"),
"model4" = c(multivar_covs, "SystemicTreatment")
)

```

```{r}
for (exposure in exposures) {
  # Set up recording frame
  results_columns <- c("Exposure", "Outcome", "Model", "Adjustment", "n", "n_event",
                       as.vector(outer(c("HR", "Lower_CI", "Upper_CI", "floatedlnHR", "floatedSE", "floatedHR", 
                                         "floatedLower_CI", "floatedUpper_CI", "n", "n_event", "mean_MET"), 
                                        unique(dat[, exposure]), paste, sep = "_")))
  results_tab <- data.frame(matrix(ncol = length(results_columns), nrow = 0))
  colnames(results_tab) <- results_columns
  
  # Do modeling
  for (outcome in outcomes) {
    for (model_name in names(adjustments)) {
      covs <- adjustments[[model_name]]
      cov_sum <- paste0(c(exposure, covs), collapse = "+")
      
      # Model
      form <- paste0("Surv(age_entry_days, age_exit_days_2, ", outcome, ") ~ ", cov_sum)
      model <- coxph(as.formula(form), data = dat)
      print(cox.zph(model))
      
      # Extract values
      results_frame <- data.frame("Exposure" = exposure, "Outcome" = outcome, "Model" = model_name,
                                  "Adjustment" = paste0(covs, collapse = ","), "n" = model$n, "n_event" = model$nevent)
      
      # Cycle through exposure levels extracting info
      fac_levels <- levels(factor(dat[, exposure]))
      for (cat in fac_levels) {
        # HRs
        if (cat == fac_levels[1]) {
          HRs <- data.frame(matrix(c(1, 1, 1), nrow = 1))
        } else {
          HRs <- summary(model)$conf.int[paste0(exposure, cat), c("exp(coef)", "lower .95", "upper .95"), drop = FALSE]
          rownames(HRs) <- NULL
        }
        colnames(HRs) <- as.vector(outer(c("HR", "Lower_CI", "Upper_CI"), cat, paste, sep = "_"))
        
        # Floated HRs etc.
        float_model <- float(model)  # This is doing floating absolute risks using Epi package
        zval <- qnorm(0.975)
        se <- sqrt(float_model$var[cat])
        lnhr <- float_model$coef[cat]
        HRs_floated <- data.frame(matrix(c(lnhr, se, exp(lnhr), exp(lnhr - se * zval), exp(lnhr + se * zval)), nrow = 1))
        colnames(HRs_floated) <- as.vector(outer(c("floatedlnHR", "floatedSE", "floatedHR", "floatedLower_CI", "floatedUpper_CI"), cat, paste, sep = "_"))
        
        # n and n_event
        n_group <- nrow(dat[dat[, exposure] == cat, ])
        n_event_group <- nrow(dat[(dat[, exposure] == cat) & (dat[, outcome]), ])
        n_by_group <- data.frame(matrix(c(n_group, n_event_group), nrow = 1))
        colnames(n_by_group) <- as.vector(outer(c("n", "n_event"), cat, paste, sep = "_"))
        
        # Mean exposure
        col_me <- paste0("mean_MET_", cat)
        mean_exposure <- data.frame(mean(dat$summed_MET_minutes[dat[, exposure] == cat]))
        colnames(mean_exposure) <- col_me
        
        # Bind together in final data frame
        results_frame <- cbind(results_frame, HRs, HRs_floated, n_by_group, mean_exposure)
      }
      
      # Check that n and n_event match by different routes
      nsum <- sum(results_frame[, as.vector(outer(c("n"), fac_levels, paste, sep = "_"))])
      n_eventsum <- sum(results_frame[, as.vector(outer(c("n_event"), fac_levels, paste, sep = "_"))])
      if ((results_frame$n != nsum) | (results_frame$n_event != n_eventsum)) {
        stop("Mismatch between numbers derived from different sources. Recheck model. May be that some rows are being excluded due to missing covariate data.")
      }
      
      # Bind into final results frame and delete
      results_tab <- rbind(results_tab, results_frame)
      rm(results_frame)
    }
  }
  
  assign(paste0(exposure, "_results_tab"), results_tab)
  write.csv(results_tab, paste0("outputs/", exposure, "_tab.csv"))
  rm(results_tab, results_columns)
}


```

2 year sensitivity analysis
```{r}
for (exposure in exposures) {
  # Set up recording frame
  results_columns <- c("Exposure", "Outcome", "Model", "Adjustment", "n", "n_event",
                       as.vector(outer(c("HR", "Lower_CI", "Upper_CI", "floatedlnHR", "floatedSE", "floatedHR", 
                                         "floatedLower_CI", "floatedUpper_CI", "n", "n_event", "mean_MET"), 
                                        unique(dat2year[, exposure]), paste, sep = "_")))
  results_tab <- data.frame(matrix(ncol = length(results_columns), nrow = 0))
  colnames(results_tab) <- results_columns
  
  # Do modeling
  for (outcome in outcomes) {
    for (model_name in names(adjustments)) {
      covs <- adjustments[[model_name]]
      cov_sum <- paste0(c(exposure, covs), collapse = "+")
      
      # Model
      form <- paste0("Surv(age_entry_days, age_exit_days_2, ", outcome, ") ~ ", cov_sum)
      model <- coxph(as.formula(form), data = dat2year)
      print(cox.zph(model))
      
      # Extract values
      results_frame <- data.frame("Exposure" = exposure, "Outcome" = outcome, "Model" = model_name,
                                  "Adjustment" = paste0(covs, collapse = ","), "n" = model$n, "n_event" = model$nevent)
      
      # Cycle through exposure levels extracting info
      fac_levels <- levels(factor(dat2year[, exposure]))
      for (cat in fac_levels) {
        # HRs
        if (cat == fac_levels[1]) {
          HRs <- data.frame(matrix(c(1, 1, 1), nrow = 1))
        } else {
          HRs <- summary(model)$conf.int[paste0(exposure, cat), c("exp(coef)", "lower .95", "upper .95"), drop = FALSE]
          rownames(HRs) <- NULL
        }
        colnames(HRs) <- as.vector(outer(c("HR", "Lower_CI", "Upper_CI"), cat, paste, sep = "_"))
        
        # Floated HRs etc.
        float_model <- float(model)  # This is doing floating absolute risks using Epi package
        zval <- qnorm(0.975)
        se <- sqrt(float_model$var[cat])
        lnhr <- float_model$coef[cat]
        HRs_floated <- data.frame(matrix(c(lnhr, se, exp(lnhr), exp(lnhr - se * zval), exp(lnhr + se * zval)), nrow = 1))
        colnames(HRs_floated) <- as.vector(outer(c("floatedlnHR", "floatedSE", "floatedHR", "floatedLower_CI", "floatedUpper_CI"), cat, paste, sep = "_"))
        
        # n and n_event
        n_group <- nrow(dat2year[dat2year[, exposure] == cat, ])
        n_event_group <- nrow(dat2year[(dat2year[, exposure] == cat) & (dat2year[, outcome]), ])
        n_by_group <- data.frame(matrix(c(n_group, n_event_group), nrow = 1))
        colnames(n_by_group) <- as.vector(outer(c("n", "n_event"), cat, paste, sep = "_"))
        
        # Mean exposure
        col_me <- paste0("mean_MET_", cat)
        mean_exposure <- data.frame(mean(dat2year$summed_MET_minutes[dat2year[, exposure] == cat]))
        colnames(mean_exposure) <- col_me
        
        # Bind together in final data frame
        results_frame <- cbind(results_frame, HRs, HRs_floated, n_by_group, mean_exposure)
      }
      
      # Check that n and n_event match by different routes
      nsum <- sum(results_frame[, as.vector(outer(c("n"), fac_levels, paste, sep = "_"))])
      n_eventsum <- sum(results_frame[, as.vector(outer(c("n_event"), fac_levels, paste, sep = "_"))])
      if ((results_frame$n != nsum) | (results_frame$n_event != n_eventsum)) {
        stop("Mismatch between numbers derived from different sources. Recheck model. May be that some rows are being excluded due to missing covariate data.")
      }
      
      # Bind into final results frame and delete
      results_tab <- rbind(results_tab, results_frame)
      rm(results_frame)
    }
  }
  
  assign(paste0(exposure, "_results_tab_2yr"), results_tab)
  write.csv(results_tab, paste0("outputs/", exposure, "_tab2yr.csv"))
  rm(results_tab, results_columns)
}


```

4 year sensitivity analysis
```{r}
for (exposure in exposures) {
  # Set up recording frame
  results_columns <- c("Exposure", "Outcome", "Model", "Adjustment", "n", "n_event",
                       as.vector(outer(c("HR", "Lower_CI", "Upper_CI", "floatedlnHR", "floatedSE", "floatedHR", 
                                         "floatedLower_CI", "floatedUpper_CI", "n", "n_event", "mean_MET"), 
                                        unique(dat4year[, exposure]), paste, sep = "_")))
  results_tab <- data.frame(matrix(ncol = length(results_columns), nrow = 0))
  colnames(results_tab) <- results_columns
  
  # Do modeling
  for (outcome in outcomes) {
    for (model_name in names(adjustments)) {
      covs <- adjustments[[model_name]]
      cov_sum <- paste0(c(exposure, covs), collapse = "+")
      
      # Model
      form <- paste0("Surv(age_entry_days, age_exit_days_2, ", outcome, ") ~ ", cov_sum)
      model <- coxph(as.formula(form), data = dat4year)
      print(cox.zph(model))
      
      # Extract values
      results_frame <- data.frame("Exposure" = exposure, "Outcome" = outcome, "Model" = model_name,
                                  "Adjustment" = paste0(covs, collapse = ","), "n" = model$n, "n_event" = model$nevent)
      
      # Cycle through exposure levels extracting info
      fac_levels <- levels(factor(dat4year[, exposure]))
      for (cat in fac_levels) {
        # HRs
        if (cat == fac_levels[1]) {
          HRs <- data.frame(matrix(c(1, 1, 1), nrow = 1))
        } else {
          HRs <- summary(model)$conf.int[paste0(exposure, cat), c("exp(coef)", "lower .95", "upper .95"), drop = FALSE]
          rownames(HRs) <- NULL
        }
        colnames(HRs) <- as.vector(outer(c("HR", "Lower_CI", "Upper_CI"), cat, paste, sep = "_"))
        
        # Floated HRs etc.
        float_model <- float(model)  # This is doing floating absolute risks using Epi package
        zval <- qnorm(0.975)
        se <- sqrt(float_model$var[cat])
        lnhr <- float_model$coef[cat]
        HRs_floated <- data.frame(matrix(c(lnhr, se, exp(lnhr), exp(lnhr - se * zval), exp(lnhr + se * zval)), nrow = 1))
        colnames(HRs_floated) <- as.vector(outer(c("floatedlnHR", "floatedSE", "floatedHR", "floatedLower_CI", "floatedUpper_CI"), cat, paste, sep = "_"))
        
        # n and n_event
        n_group <- nrow(dat4year[dat4year[, exposure] == cat, ])
        n_event_group <- nrow(dat4year[(dat4year[, exposure] == cat) & (dat4year[, outcome]), ])
        n_by_group <- data.frame(matrix(c(n_group, n_event_group), nrow = 1))
        colnames(n_by_group) <- as.vector(outer(c("n", "n_event"), cat, paste, sep = "_"))
        
        # Mean exposure
        col_me <- paste0("mean_MET_", cat)
        mean_exposure <- data.frame(mean(dat4year$summed_MET_minutes[dat4year[, exposure] == cat]))
        colnames(mean_exposure) <- col_me
        
        # Bind together in final data frame
        results_frame <- cbind(results_frame, HRs, HRs_floated, n_by_group, mean_exposure)
      }
      
      # Check that n and n_event match by different routes
      nsum <- sum(results_frame[, as.vector(outer(c("n"), fac_levels, paste, sep = "_"))])
      n_eventsum <- sum(results_frame[, as.vector(outer(c("n_event"), fac_levels, paste, sep = "_"))])
      if ((results_frame$n != nsum) | (results_frame$n_event != n_eventsum)) {
        stop("Mismatch between numbers derived from different sources. Recheck model. May be that some rows are being excluded due to missing covariate data.")
      }
      
      # Bind into final results frame and delete
      results_tab <- rbind(results_tab, results_frame)
      rm(results_frame)
    }
  }
  
  assign(paste0(exposure, "_results_tab4yr"), results_tab)
  write.csv(results_tab, paste0("outputs/", exposure, "_tab_4yr.csv"))
  rm(results_tab, results_columns)
}


```

6 year sensitivity analysis
```{r}
for (exposure in exposures) {
  # Set up recording frame
  results_columns <- c("Exposure", "Outcome", "Model", "Adjustment", "n", "n_event",
                       as.vector(outer(c("HR", "Lower_CI", "Upper_CI", "floatedlnHR", "floatedSE", "floatedHR", 
                                         "floatedLower_CI", "floatedUpper_CI", "n", "n_event", "mean_MET"), 
                                        unique(dat6year[, exposure]), paste, sep = "_")))
  results_tab <- data.frame(matrix(ncol = length(results_columns), nrow = 0))
  colnames(results_tab) <- results_columns
  
  # Do modeling
  for (outcome in outcomes) {
    for (model_name in names(adjustments)) {
      covs <- adjustments[[model_name]]
      cov_sum <- paste0(c(exposure, covs), collapse = "+")
      
      # Model
      form <- paste0("Surv(age_entry_days, age_exit_days_2, ", outcome, ") ~ ", cov_sum)
      model <- coxph(as.formula(form), data = dat6year)
      print(cox.zph(model))
      
      # Extract values
      results_frame <- data.frame("Exposure" = exposure, "Outcome" = outcome, "Model" = model_name,
                                  "Adjustment" = paste0(covs, collapse = ","), "n" = model$n, "n_event" = model$nevent)
      
      # Cycle through exposure levels extracting info
      fac_levels <- levels(factor(dat6year[, exposure]))
      for (cat in fac_levels) {
        # HRs
        if (cat == fac_levels[1]) {
          HRs <- data.frame(matrix(c(1, 1, 1), nrow = 1))
        } else {
          HRs <- summary(model)$conf.int[paste0(exposure, cat), c("exp(coef)", "lower .95", "upper .95"), drop = FALSE]
          rownames(HRs) <- NULL
        }
        colnames(HRs) <- as.vector(outer(c("HR", "Lower_CI", "Upper_CI"), cat, paste, sep = "_"))
        
        # Floated HRs etc.
        float_model <- float(model)  # This is doing floating absolute risks using Epi package
        zval <- qnorm(0.975)
        se <- sqrt(float_model$var[cat])
        lnhr <- float_model$coef[cat]
        HRs_floated <- data.frame(matrix(c(lnhr, se, exp(lnhr), exp(lnhr - se * zval), exp(lnhr + se * zval)), nrow = 1))
        colnames(HRs_floated) <- as.vector(outer(c("floatedlnHR", "floatedSE", "floatedHR", "floatedLower_CI", "floatedUpper_CI"), cat, paste, sep = "_"))
        
        # n and n_event
        n_group <- nrow(dat6year[dat6year[, exposure] == cat, ])
        n_event_group <- nrow(dat6year[(dat6year[, exposure] == cat) & (dat6year[, outcome]), ])
        n_by_group <- data.frame(matrix(c(n_group, n_event_group), nrow = 1))
        colnames(n_by_group) <- as.vector(outer(c("n", "n_event"), cat, paste, sep = "_"))
        
        # Mean exposure
        col_me <- paste0("mean_MET_", cat)
        mean_exposure <- data.frame(mean(dat6year$summed_MET_minutes[dat6year[, exposure] == cat]))
        colnames(mean_exposure) <- col_me
        
        # Bind together in final data frame
        results_frame <- cbind(results_frame, HRs, HRs_floated, n_by_group, mean_exposure)
      }
      
      # Check that n and n_event match by different routes
      nsum <- sum(results_frame[, as.vector(outer(c("n"), fac_levels, paste, sep = "_"))])
      n_eventsum <- sum(results_frame[, as.vector(outer(c("n_event"), fac_levels, paste, sep = "_"))])
      if ((results_frame$n != nsum) | (results_frame$n_event != n_eventsum)) {
        stop("Mismatch between numbers derived from different sources. Recheck model. May be that some rows are being excluded due to missing covariate data.")
      }
      
      # Bind into final results frame and delete
      results_tab <- rbind(results_tab, results_frame)
      rm(results_frame)
    }
  }
  
  assign(paste0(exposure, "_results_tab6yr"), results_tab)
  write.csv(results_tab, paste0("outputs/", exposure, "_tab_6yr.csv"))
  rm(results_tab, results_columns)
}


```

Other definition HES probable only age_exit_days_3, ProbPsA_incident
```{r}
exposures <- c("MET_tertiles", "MET_quartiles", "MET_quintiles")
outcomes <- c("ProbPsA_incident") 

multivar_covs <-
c(
"sex",
"ethnicity",
strata("tdi_quarters"), 
"qualif",
"smoking",
"alcohol"
)

adjustments <-
list(
"model1" = NULL,
"model2" = c(multivar_covs),
"model3" = c(multivar_covs, "BMI_cats"),
"model4" = c(multivar_covs, "SystemicTreatment")
)

```

```{r}
for (exposure in exposures) {
  # Set up recording frame
  results_columns <- c("Exposure", "Outcome", "Model", "Adjustment", "n", "n_event",
                       as.vector(outer(c("HR", "Lower_CI", "Upper_CI", "floatedlnHR", "floatedSE", "floatedHR", 
                                         "floatedLower_CI", "floatedUpper_CI", "n", "n_event", "mean_MET"), 
                                        unique(dat[, exposure]), paste, sep = "_")))
  results_tab <- data.frame(matrix(ncol = length(results_columns), nrow = 0))
  colnames(results_tab) <- results_columns
  
  # Do modeling
  for (outcome in outcomes) {
    for (model_name in names(adjustments)) {
      covs <- adjustments[[model_name]]
      cov_sum <- paste0(c(exposure, covs), collapse = "+")
      
      # Model
      form <- paste0("Surv(age_entry_days, age_exit_days_3, ", outcome, ") ~ ", cov_sum)
      model <- coxph(as.formula(form), data = dat)
      print(cox.zph(model))
      
      # Extract values
      results_frame <- data.frame("Exposure" = exposure, "Outcome" = outcome, "Model" = model_name,
                                  "Adjustment" = paste0(covs, collapse = ","), "n" = model$n, "n_event" = model$nevent)
      
      # Cycle through exposure levels extracting info
      fac_levels <- levels(factor(dat[, exposure]))
      for (cat in fac_levels) {
        # HRs
        if (cat == fac_levels[1]) {
          HRs <- data.frame(matrix(c(1, 1, 1), nrow = 1))
        } else {
          HRs <- summary(model)$conf.int[paste0(exposure, cat), c("exp(coef)", "lower .95", "upper .95"), drop = FALSE]
          rownames(HRs) <- NULL
        }
        colnames(HRs) <- as.vector(outer(c("HR", "Lower_CI", "Upper_CI"), cat, paste, sep = "_"))
        
        # Floated HRs etc.
        float_model <- float(model)  # This is doing floating absolute risks using Epi package
        zval <- qnorm(0.975)
        se <- sqrt(float_model$var[cat])
        lnhr <- float_model$coef[cat]
        HRs_floated <- data.frame(matrix(c(lnhr, se, exp(lnhr), exp(lnhr - se * zval), exp(lnhr + se * zval)), nrow = 1))
        colnames(HRs_floated) <- as.vector(outer(c("floatedlnHR", "floatedSE", "floatedHR", "floatedLower_CI", "floatedUpper_CI"), cat, paste, sep = "_"))
        
        # n and n_event
        n_group <- nrow(dat[dat[, exposure] == cat, ])
        n_event_group <- nrow(dat[(dat[, exposure] == cat) & (dat[, outcome]), ])
        n_by_group <- data.frame(matrix(c(n_group, n_event_group), nrow = 1))
        colnames(n_by_group) <- as.vector(outer(c("n", "n_event"), cat, paste, sep = "_"))
        
        # Mean exposure
        col_me <- paste0("mean_MET_", cat)
        mean_exposure <- data.frame(mean(dat$summed_MET_minutes[dat[, exposure] == cat]))
        colnames(mean_exposure) <- col_me
        
        # Bind together in final data frame
        results_frame <- cbind(results_frame, HRs, HRs_floated, n_by_group, mean_exposure)
      }
      
      # Check that n and n_event match by different routes
      nsum <- sum(results_frame[, as.vector(outer(c("n"), fac_levels, paste, sep = "_"))])
      n_eventsum <- sum(results_frame[, as.vector(outer(c("n_event"), fac_levels, paste, sep = "_"))])
      if ((results_frame$n != nsum) | (results_frame$n_event != n_eventsum)) {
        stop("Mismatch between numbers derived from different sources. Recheck model. May be that some rows are being excluded due to missing covariate data.")
      }
      
      # Bind into final results frame and delete
      results_tab <- rbind(results_tab, results_frame)
      rm(results_frame)
    }
  }
  
  assign(paste0(exposure, "_results_tab"), results_tab)
  write.csv(results_tab, paste0("outputs/", exposure, "_tab.csv"))
  rm(results_tab, results_columns)
}


```

Other definition HES Definite only -> age_exit_days_1, PsA_incident_HES
```{r}
exposures <- c("MET_tertiles", "MET_quartiles", "MET_quintiles")
outcomes <- c("PsA_incident_HES") 

multivar_covs <-
c(
"sex",
"ethnicity",
strata("tdi_quarters"), 
"qualif",
"smoking",
"alcohol"
)

adjustments <-
list(
"model1" = NULL,
"model2" = c(multivar_covs),
"model3" = c(multivar_covs, "BMI_cats"),
"model4" = c(multivar_covs, "SystemicTreatment")
)

```

```{r}
for (exposure in exposures) {
  # Set up recording frame
  results_columns <- c("Exposure", "Outcome", "Model", "Adjustment", "n", "n_event",
                       as.vector(outer(c("HR", "Lower_CI", "Upper_CI", "floatedlnHR", "floatedSE", "floatedHR", 
                                         "floatedLower_CI", "floatedUpper_CI", "n", "n_event", "mean_MET"), 
                                        unique(dat[, exposure]), paste, sep = "_")))
  results_tab <- data.frame(matrix(ncol = length(results_columns), nrow = 0))
  colnames(results_tab) <- results_columns
  
  # Do modeling
  for (outcome in outcomes) {
    for (model_name in names(adjustments)) {
      covs <- adjustments[[model_name]]
      cov_sum <- paste0(c(exposure, covs), collapse = "+")
      
      # Model
      form <- paste0("Surv(age_entry_days, age_exit_days_1, ", outcome, ") ~ ", cov_sum)
      model <- coxph(as.formula(form), data = dat)
      print(cox.zph(model))
      
      # Extract values
      results_frame <- data.frame("Exposure" = exposure, "Outcome" = outcome, "Model" = model_name,
                                  "Adjustment" = paste0(covs, collapse = ","), "n" = model$n, "n_event" = model$nevent)
      
      # Cycle through exposure levels extracting info
      fac_levels <- levels(factor(dat[, exposure]))
      for (cat in fac_levels) {
        # HRs
        if (cat == fac_levels[1]) {
          HRs <- data.frame(matrix(c(1, 1, 1), nrow = 1))
        } else {
          HRs <- summary(model)$conf.int[paste0(exposure, cat), c("exp(coef)", "lower .95", "upper .95"), drop = FALSE]
          rownames(HRs) <- NULL
        }
        colnames(HRs) <- as.vector(outer(c("HR", "Lower_CI", "Upper_CI"), cat, paste, sep = "_"))
        
        # Floated HRs etc.
        float_model <- float(model)  # This is doing floating absolute risks using Epi package
        zval <- qnorm(0.975)
        se <- sqrt(float_model$var[cat])
        lnhr <- float_model$coef[cat]
        HRs_floated <- data.frame(matrix(c(lnhr, se, exp(lnhr), exp(lnhr - se * zval), exp(lnhr + se * zval)), nrow = 1))
        colnames(HRs_floated) <- as.vector(outer(c("floatedlnHR", "floatedSE", "floatedHR", "floatedLower_CI", "floatedUpper_CI"), cat, paste, sep = "_"))
        
        # n and n_event
        n_group <- nrow(dat[dat[, exposure] == cat, ])
        n_event_group <- nrow(dat[(dat[, exposure] == cat) & (dat[, outcome]), ])
        n_by_group <- data.frame(matrix(c(n_group, n_event_group), nrow = 1))
        colnames(n_by_group) <- as.vector(outer(c("n", "n_event"), cat, paste, sep = "_"))
        
        # Mean exposure
        col_me <- paste0("mean_MET_", cat)
        mean_exposure <- data.frame(mean(dat$summed_MET_minutes[dat[, exposure] == cat]))
        colnames(mean_exposure) <- col_me
        
        # Bind together in final data frame
        results_frame <- cbind(results_frame, HRs, HRs_floated, n_by_group, mean_exposure)
      }
      
      # Check that n and n_event match by different routes
      nsum <- sum(results_frame[, as.vector(outer(c("n"), fac_levels, paste, sep = "_"))])
      n_eventsum <- sum(results_frame[, as.vector(outer(c("n_event"), fac_levels, paste, sep = "_"))])
      if ((results_frame$n != nsum) | (results_frame$n_event != n_eventsum)) {
        stop("Mismatch between numbers derived from different sources. Recheck model. May be that some rows are being excluded due to missing covariate data.")
      }
      
      # Bind into final results frame and delete
      results_tab <- rbind(results_tab, results_frame)
      rm(results_frame)
    }
  }
  
  assign(paste0(exposure, "_results_tab"), results_tab)
  write.csv(results_tab, paste0("outputs/", exposure, "_tab.csv"))
  rm(results_tab, results_columns)
}


```

2 year
```{r}
for (exposure in exposures) {
  # Set up recording frame
  results_columns <- c("Exposure", "Outcome", "Model", "Adjustment", "n", "n_event",
                       as.vector(outer(c("HR", "Lower_CI", "Upper_CI", "floatedlnHR", "floatedSE", "floatedHR", 
                                         "floatedLower_CI", "floatedUpper_CI", "n", "n_event", "mean_MET"), 
                                        unique(dat2year[, exposure]), paste, sep = "_")))
  results_tab <- data.frame(matrix(ncol = length(results_columns), nrow = 0))
  colnames(results_tab) <- results_columns
  
  # Do modeling
  for (outcome in outcomes) {
    for (model_name in names(adjustments)) {
      covs <- adjustments[[model_name]]
      cov_sum <- paste0(c(exposure, covs), collapse = "+")
      
      # Model
      form <- paste0("Surv(age_entry_days, age_exit_days_1, ", outcome, ") ~ ", cov_sum)
      model <- coxph(as.formula(form), data = dat2year)
      print(cox.zph(model))
      
      # Extract values
      results_frame <- data.frame("Exposure" = exposure, "Outcome" = outcome, "Model" = model_name,
                                  "Adjustment" = paste0(covs, collapse = ","), "n" = model$n, "n_event" = model$nevent)
      
      # Cycle through exposure levels extracting info
      fac_levels <- levels(factor(dat2year[, exposure]))
      for (cat in fac_levels) {
        # HRs
        if (cat == fac_levels[1]) {
          HRs <- data.frame(matrix(c(1, 1, 1), nrow = 1))
        } else {
          HRs <- summary(model)$conf.int[paste0(exposure, cat), c("exp(coef)", "lower .95", "upper .95"), drop = FALSE]
          rownames(HRs) <- NULL
        }
        colnames(HRs) <- as.vector(outer(c("HR", "Lower_CI", "Upper_CI"), cat, paste, sep = "_"))
        
        # Floated HRs etc.
        float_model <- float(model)  # This is doing floating absolute risks using Epi package
        zval <- qnorm(0.975)
        se <- sqrt(float_model$var[cat])
        lnhr <- float_model$coef[cat]
        HRs_floated <- data.frame(matrix(c(lnhr, se, exp(lnhr), exp(lnhr - se * zval), exp(lnhr + se * zval)), nrow = 1))
        colnames(HRs_floated) <- as.vector(outer(c("floatedlnHR", "floatedSE", "floatedHR", "floatedLower_CI", "floatedUpper_CI"), cat, paste, sep = "_"))
        
        # n and n_event
        n_group <- nrow(dat2year[dat2year[, exposure] == cat, ])
        n_event_group <- nrow(dat2year[(dat2year[, exposure] == cat) & (dat2year[, outcome]), ])
        n_by_group <- data.frame(matrix(c(n_group, n_event_group), nrow = 1))
        colnames(n_by_group) <- as.vector(outer(c("n", "n_event"), cat, paste, sep = "_"))
        
        # Mean exposure
        col_me <- paste0("mean_MET_", cat)
        mean_exposure <- data.frame(mean(dat2year$summed_MET_minutes[dat2year[, exposure] == cat]))
        colnames(mean_exposure) <- col_me
        
        # Bind together in final data frame
        results_frame <- cbind(results_frame, HRs, HRs_floated, n_by_group, mean_exposure)
      }
      
      # Check that n and n_event match by different routes
      nsum <- sum(results_frame[, as.vector(outer(c("n"), fac_levels, paste, sep = "_"))])
      n_eventsum <- sum(results_frame[, as.vector(outer(c("n_event"), fac_levels, paste, sep = "_"))])
      if ((results_frame$n != nsum) | (results_frame$n_event != n_eventsum)) {
        stop("Mismatch between numbers derived from different sources. Recheck model. May be that some rows are being excluded due to missing covariate data.")
      }
      
      # Bind into final results frame and delete
      results_tab <- rbind(results_tab, results_frame)
      rm(results_frame)
    }
  }
  
  assign(paste0(exposure, "_results_tab"), results_tab)
  write.csv(results_tab, paste0("outputs/", exposure, "_tab_DEF_2yr.csv"))
  rm(results_tab, results_columns)
}


```



4 year DEFINITE 
```{r}
for (exposure in exposures) {
  # Set up recording frame
  results_columns <- c("Exposure", "Outcome", "Model", "Adjustment", "n", "n_event",
                       as.vector(outer(c("HR", "Lower_CI", "Upper_CI", "floatedlnHR", "floatedSE", "floatedHR", 
                                         "floatedLower_CI", "floatedUpper_CI", "n", "n_event", "mean_MET"), 
                                        unique(dat4year[, exposure]), paste, sep = "_")))
  results_tab <- data.frame(matrix(ncol = length(results_columns), nrow = 0))
  colnames(results_tab) <- results_columns
  
  # Do modeling
  for (outcome in outcomes) {
    for (model_name in names(adjustments)) {
      covs <- adjustments[[model_name]]
      cov_sum <- paste0(c(exposure, covs), collapse = "+")
      
      # Model
      form <- paste0("Surv(age_entry_days, age_exit_days_1, ", outcome, ") ~ ", cov_sum)
      model <- coxph(as.formula(form), data = dat4year)
      print(cox.zph(model))
      
      # Extract values
      results_frame <- data.frame("Exposure" = exposure, "Outcome" = outcome, "Model" = model_name,
                                  "Adjustment" = paste0(covs, collapse = ","), "n" = model$n, "n_event" = model$nevent)
      
      # Cycle through exposure levels extracting info
      fac_levels <- levels(factor(dat4year[, exposure]))
      for (cat in fac_levels) {
        # HRs
        if (cat == fac_levels[1]) {
          HRs <- data.frame(matrix(c(1, 1, 1), nrow = 1))
        } else {
          HRs <- summary(model)$conf.int[paste0(exposure, cat), c("exp(coef)", "lower .95", "upper .95"), drop = FALSE]
          rownames(HRs) <- NULL
        }
        colnames(HRs) <- as.vector(outer(c("HR", "Lower_CI", "Upper_CI"), cat, paste, sep = "_"))
        
        # Floated HRs etc.
        float_model <- float(model)  # This is doing floating absolute risks using Epi package
        zval <- qnorm(0.975)
        se <- sqrt(float_model$var[cat])
        lnhr <- float_model$coef[cat]
        HRs_floated <- data.frame(matrix(c(lnhr, se, exp(lnhr), exp(lnhr - se * zval), exp(lnhr + se * zval)), nrow = 1))
        colnames(HRs_floated) <- as.vector(outer(c("floatedlnHR", "floatedSE", "floatedHR", "floatedLower_CI", "floatedUpper_CI"), cat, paste, sep = "_"))
        
        # n and n_event
        n_group <- nrow(dat4year[dat4year[, exposure] == cat, ])
        n_event_group <- nrow(dat4year[(dat4year[, exposure] == cat) & (dat4year[, outcome]), ])
        n_by_group <- data.frame(matrix(c(n_group, n_event_group), nrow = 1))
        colnames(n_by_group) <- as.vector(outer(c("n", "n_event"), cat, paste, sep = "_"))
        
        # Mean exposure
        col_me <- paste0("mean_MET_", cat)
        mean_exposure <- data.frame(mean(dat4year$summed_MET_minutes[dat4year[, exposure] == cat]))
        colnames(mean_exposure) <- col_me
        
        # Bind together in final data frame
        results_frame <- cbind(results_frame, HRs, HRs_floated, n_by_group, mean_exposure)
      }
      
      # Check that n and n_event match by different routes
      nsum <- sum(results_frame[, as.vector(outer(c("n"), fac_levels, paste, sep = "_"))])
      n_eventsum <- sum(results_frame[, as.vector(outer(c("n_event"), fac_levels, paste, sep = "_"))])
      if ((results_frame$n != nsum) | (results_frame$n_event != n_eventsum)) {
        stop("Mismatch between numbers derived from different sources. Recheck model. May be that some rows are being excluded due to missing covariate data.")
      }
      
      # Bind into final results frame and delete
      results_tab <- rbind(results_tab, results_frame)
      rm(results_frame)
    }
  }
  
  assign(paste0(exposure, "_results_tab"), results_tab)
  write.csv(results_tab, paste0("outputs/", exposure, "_tab_DEF_4yr.csv"))
  rm(results_tab, results_columns)
}


```


4 year DEFINITE 
```{r}
for (exposure in exposures) {
  # Set up recording frame
  results_columns <- c("Exposure", "Outcome", "Model", "Adjustment", "n", "n_event",
                       as.vector(outer(c("HR", "Lower_CI", "Upper_CI", "floatedlnHR", "floatedSE", "floatedHR", 
                                         "floatedLower_CI", "floatedUpper_CI", "n", "n_event", "mean_MET"), 
                                        unique(dat6year[, exposure]), paste, sep = "_")))
  results_tab <- data.frame(matrix(ncol = length(results_columns), nrow = 0))
  colnames(results_tab) <- results_columns
  
  # Do modeling
  for (outcome in outcomes) {
    for (model_name in names(adjustments)) {
      covs <- adjustments[[model_name]]
      cov_sum <- paste0(c(exposure, covs), collapse = "+")
      
      # Model
      form <- paste0("Surv(age_entry_days, age_exit_days_1, ", outcome, ") ~ ", cov_sum)
      model <- coxph(as.formula(form), data = dat6year)
      print(cox.zph(model))
      
      # Extract values
      results_frame <- data.frame("Exposure" = exposure, "Outcome" = outcome, "Model" = model_name,
                                  "Adjustment" = paste0(covs, collapse = ","), "n" = model$n, "n_event" = model$nevent)
      
      # Cycle through exposure levels extracting info
      fac_levels <- levels(factor(dat6year[, exposure]))
      for (cat in fac_levels) {
        # HRs
        if (cat == fac_levels[1]) {
          HRs <- data.frame(matrix(c(1, 1, 1), nrow = 1))
        } else {
          HRs <- summary(model)$conf.int[paste0(exposure, cat), c("exp(coef)", "lower .95", "upper .95"), drop = FALSE]
          rownames(HRs) <- NULL
        }
        colnames(HRs) <- as.vector(outer(c("HR", "Lower_CI", "Upper_CI"), cat, paste, sep = "_"))
        
        # Floated HRs etc.
        float_model <- float(model)  # This is doing floating absolute risks using Epi package
        zval <- qnorm(0.975)
        se <- sqrt(float_model$var[cat])
        lnhr <- float_model$coef[cat]
        HRs_floated <- data.frame(matrix(c(lnhr, se, exp(lnhr), exp(lnhr - se * zval), exp(lnhr + se * zval)), nrow = 1))
        colnames(HRs_floated) <- as.vector(outer(c("floatedlnHR", "floatedSE", "floatedHR", "floatedLower_CI", "floatedUpper_CI"), cat, paste, sep = "_"))
        
        # n and n_event
        n_group <- nrow(dat6year[dat6year[, exposure] == cat, ])
        n_event_group <- nrow(dat6year[(dat6year[, exposure] == cat) & (dat6year[, outcome]), ])
        n_by_group <- data.frame(matrix(c(n_group, n_event_group), nrow = 1))
        colnames(n_by_group) <- as.vector(outer(c("n", "n_event"), cat, paste, sep = "_"))
        
        # Mean exposure
        col_me <- paste0("mean_MET_", cat)
        mean_exposure <- data.frame(mean(dat6year$summed_MET_minutes[dat6year[, exposure] == cat]))
        colnames(mean_exposure) <- col_me
        
        # Bind together in final data frame
        results_frame <- cbind(results_frame, HRs, HRs_floated, n_by_group, mean_exposure)
      }
      
      # Check that n and n_event match by different routes
      nsum <- sum(results_frame[, as.vector(outer(c("n"), fac_levels, paste, sep = "_"))])
      n_eventsum <- sum(results_frame[, as.vector(outer(c("n_event"), fac_levels, paste, sep = "_"))])
      if ((results_frame$n != nsum) | (results_frame$n_event != n_eventsum)) {
        stop("Mismatch between numbers derived from different sources. Recheck model. May be that some rows are being excluded due to missing covariate data.")
      }
      
      # Bind into final results frame and delete
      results_tab <- rbind(results_tab, results_frame)
      rm(results_frame)
    }
  }
  
  assign(paste0(exposure, "_results_tab"), results_tab)
  write.csv(results_tab, paste0("outputs/", exposure, "_tab_DEF_6yr.csv"))
  rm(results_tab, results_columns)
}


```

```{r}
dir.create("outputs")
```
