---
title: "Compositional analysis"
author: "Dylan McGagh"
date: "`r format(Sys.time(), '%a %b %d, %Y %X')`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 5
    theme: lumen
---
```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This script runs the compositional analysis based on Rosemary's scirpt for Jemma's project that can be found [here](https://github.com/OxWearables/AF_RW/blob/main/scripts_pipeline/compositional_script.R).

# Set up workspace
```{r}
#Load standard packages
pkgs <-
  c(
    "data.table",
    "rmarkdown",
    "plyr",
    "lubridate",
    "magrittr",
    "ggplot2",
    "data.table",
    "dplyr",
    "stringr",
    "splitstackshape",
    "scales",
    "haven",
    "Epi",
    #"rtables",
    "table1",
   # "tableone",
    "boot",
    "knitr",
    "pspline",
    "survival",
    "xtable",
    "here",
    "ggtern",
    "devtools"
  ) # packages we need
#Note - At install, load epicoda using Rosemary's instructions from GitHub - https://github.com/OxWearables/epicoda

#install
pkgs_inst <-
  pkgs[!{
    pkgs %in% rownames(installed.packages())
  }] # check which are not present
install.packages(pkgs_inst, repos = "https://www.stats.bris.ac.uk/R/") # install

lapply(pkgs, library, character.only = TRUE) #load all packages

#renv::install(pkgs)


## Source helper functions---------------------------------------------------
source(here("compare_plot_linear.R"))
source("rounding_functions.R")
source("cut_by_quantile.R")
source("cox_functions.R")
source(here("arrange_plots_matrix.R"))
source(here("med_and_iqr.R"))
source(here("forest_plot_examples_with_evals.R"))
source(here("compare_plot2.R"))
source(here("compare_plot3.R"))
source(here("plot_transfers_ism.R"))

```

Walmsley plots 
```{r}
library(devtools)
library(epicoda)
```

## Load data
```{r}
#############read in data
#updated file path 2023 10 16
dat <- fread("PsAfinal/prepped_steps_PsAfinalAccelerometer.csv", data.table = FALSE)



dat$V1 <- NULL

#dat <- dat %>% filter(!is.na(ethnicity))

```

Restrict the accelerometer analysis to white ethnicity only 
```{r}
dat <- dat %>% filter(ethnicity == "White")
```


```{r}
dir.create("outputs")
```

```{r}
dat$tdi_quarters <- qtile_cut(dat$tdi_raw, probs = seq(0, 1, by = 0.25), dp_label = 1)
dat$BMI_cats <-
  cut(dat$BMI,
      breaks = c(0, 25, 30, 10000),
      labels = c("<24.9", "25.0-29.9", "30.0+"),
      right = FALSE)

dat$sleep <- NULL
dat$SB <- NULL
dat$LIPA <- NULL
dat$MVPA <- NULL

dat$MVPA <- dat$MVPA_overall_average
dat$sleep <- dat$sleep_overall_average
dat$SB <- dat$sedentary_overall_average
dat$LIPA <- dat$light_overall_average 

#1 participant with NA for MVPA
dat <- dat[!is.na(dat$MVPA), ]
```

```{r}
dat2year_cases <- (dat$ProbDefPsA_incident & dat$fu_time_2 <= 760)
dat4year_cases <- (dat$ProbDefPsA_incident & dat$fu_time_2 <= 1461)


dat <- dat[!(dat2year_cases), ]
#dat <- dat[!(dat4year_cases), ]

```

# Prep dataset
```{r}
# Set up for modelling=====================================
### Alaina updated covariates to match other scripts
covariates <-
  c(
"sex",
"tdi_quarters",
"qualif",
"smoking",
"alcohol"
  )

# AHS add code for time points
dat$agein <- dat$age_entry_days
dat$ageout <- dat$age_exit_days_2

# Define the names of the comp variables in the main group
comp_labels <- c("sleep", "SB", "LIPA", "MVPA")
# Check sum to 1
dat$sum_sleep_SB_LIPA_MVPA <- rowSums(dat[, c("sleep", "SB", "LIPA", "MVPA")])
table(dat$sum_sleep_SB_LIPA_MVPA)

### for  models that combine activity behavior in a single category
# Combine exposure categories and make new labels
dat$LIPA_MVPA <- (dat$LIPA + dat$MVPA)
# Define the names of the compositional variables
comp_labels_three <- c("sleep", "SB", "LIPA_MVPA")

# Define the names of the compositional variables for the cutpoint based analysis
comp_labels_cutpoint <- c("SB_cutpoint", "sleep", "LIPA_cutpoint","MVPA_cutpoint")
# Check sum to 1
#dat$sum_SB_cutpoint_sleep_LIPA_cutpoint_MVPA_cutpoint <- rowSums(dat[, c("SB_cutpoint", "sleep", "LIPA_cutpoint", "MVPA_cutpoint")])
#table(dat$sum_SB_cutpoint_sleep_LIPA_cutpoint_MVPA_cutpoint)

# Define the names of the compositional variables for the cutpoint based analysis with VPA seprate
comp_labels_cutpoint_fivecat <- c("SB_cutpoint", "sleep", "LIPA_cutpoint","MPA_cutpoint","VPA_cutpoint")
```

# Modeling - all
## Matrix - main 
Model compositional matrix for risk of PsA
```{r}
# Run model ==============================================
######### Incident PsA  
main_model <- comp_model(
  type = "cox",
  covariates = covariates,
  outcome = Surv(
    time = dat$agein,
    time2 = dat$ageout,
    event = dat$ProbDefPsA_incident),
  data = dat,
  comp_labels = comp_labels,
  rounded_zeroes = TRUE,
  det_limit = min(dat$MVPA[dat$MVPA >0]))


# predictions (note that this takes a while to run)
tab_coefs(scale_type = "exp",
          level = 0.95,
          type = "cox",
  covariates = covariates,
  outcome = Surv(
    time = dat$agein,
    time2 = dat$ageout,
    event = dat$ProbDefPsA_incident),
  data = dat,
  comp_labels = comp_labels,
  rounded_zeroes = TRUE,
  det_limit = min(dat$MVPA[dat$MVPA >0]))



# Check proportional hazards==============================
# Note From Rosemary's paper - Participants with missing data in any adjustment variable were excluded. The proportional hazards assumption was tested component-wise and globally using the Grambsch-Therneau test with the Kaplan-Meier transformation, and there was no evidence (at the 5% level) that it was violated in the main analysis. Plots of the Schoenfeld residuals were also examined.
cox.zph(main_model) # Grambsch-Therneau test
plot(cox.zph(main_model)) # Plots of scaled Schoenfeld residuals.

# Get information for plot ==============================
summary(main_model)
# This is the value of all the movement behaviours at the compositional mean behaviours
# i.e. the point marked with the vertical line on the graph.
get_cm_from_model(main_model, comp_labels, transf_labels(comp_labels, "ilr"))$cm*24
get_cm_from_model(main_model, comp_labels, transf_labels(comp_labels, "ilr"))$cm*24*60

# Main matrix plot ============================================
# This makes and writes to file the plot
svg(
  file=here::here("outputs", paste0("comp_figure.svg")), width = 15, height = 10
)

# Save ggplot object as SVG in the "/outputs/" folder - revisit when ready to print figures
output_folder <- here::here("outputs")
svg_file <- file.path(output_folder, (paste0("comp_figure.svg")))
plot <- plot_maxi_matrix_transfers(
  comp_model = main_model,
  comp_labels = comp_labels,
  yllimit = 0.8,
  yulimit = 1.3,
  plot_log = TRUE,
  granularity = 4000,
  point_specification = ggplot2::geom_point(size = 0.5),
  theme = theme_classic() + theme(axis.line = element_blank())
)
ggsave(filename = svg_file, plot = plot, device = "svg", width = 15, height = 10)

#Save as a jpg for poster if needed
output_folder <- here::here("outputs")
svg_file <- file.path(output_folder, paste0("comp_figure_tall.svg"))
ggsave(filename = svg_file, plot = plot, device = "svg", width = 8, height = 8)
```

## Matrix - Sensitivity analyses
Before and after adjustment for reverse causation (remove first 2 years of FU)
Before and after adjustment for BMI

```{r}
############ Build models for comparison
#main_model is defined above
## Minimal adjustment----------------------------
### Run model
minimal_model <-
  comp_model(
    type = "cox",
    outcome = Surv(
    time = dat$agein,
    time2 = dat$ageout,
    event = dat$ProbDefPsA_incident),
    data = dat,
    comp_labels = comp_labels,
    rounded_zeroes = TRUE,
    det_limit = min(dat$MVPA[dat$MVPA >0])
  )

## Sensitivity analyses for reverse causation - 2 year----------------------------
### Prep dataset
all_only_fu <- dat
all_only_fu$age_entry_days <- all_only_fu$age_entry_days + (365.25 * 2)
all_only_fu <-  all_only_fu[all_only_fu$age_entry_days < all_only_fu$age_exit_days_2, ]

# Recode relevant time points
all_only_fu$agein <- all_only_fu$age_entry_days
all_only_fu$ageout <- all_only_fu$age_exit_days_2

### Run model
only_fu_model <-
  comp_model(
    type = "cox",
    covariates = c(covariates),
    outcome = Surv(
    time = all_only_fu$agein,
    time2 = all_only_fu$ageout,
    event = all_only_fu$ProbDefPsA_incident),
    data = all_only_fu,
    comp_labels = comp_labels,
    rounded_zeroes = TRUE,
    det_limit = min(all_only_fu$MVPA[all_only_fu$MVPA >0])
  )

## Sensitivity analyses for reverse causation -4 year----------------------------
### Prep dataset
all_only_fu_4yr <- dat
all_only_fu_4yr$age_entry_days <- all_only_fu_4yr$age_entry_days + (365.25 * 4)
all_only_fu_4yr <-  all_only_fu_4yr[all_only_fu_4yr$age_entry_days < all_only_fu_4yr$age_exit_days_2, ]

# Recode relevant time points
all_only_fu_4yr$agein <- all_only_fu_4yr$age_entry_days
all_only_fu_4yr$ageout <- all_only_fu_4yr$age_exit_days_2

### Run model
only_fu_4yr_model <-
  comp_model(
    type = "cox",
    covariates = c(covariates),
    outcome = Surv(
    time = all_only_fu_4yr$agein,
    time2 = all_only_fu_4yr$ageout,
    event = all_only_fu_4yr$ProbDefPsA_incident),
    data = all_only_fu_4yr,
    comp_labels = comp_labels,
    rounded_zeroes = TRUE,
    det_limit = min(all_only_fu_4yr$MVPA[all_only_fu_4yr$MVPA >0])
  )

## Sensitivity analyses for BMI adjustment----------------------------
### Run model
add_adj_bmi_model <- comp_model(
  type = "cox",
  covariates = c(covariates, "BMI_cats") ,
  outcome = Surv(
    time = dat$agein,
    time2 = dat$ageout,
    event = dat$ProbDefPsA_incident),
  data = dat,
  comp_labels = comp_labels,
  rounded_zeroes = TRUE,
  det_limit = min(dat$MVPA[dat$MVPA >0]))

# Plotting ----------------------------------------------------------------
## Matrix plots: model comparison-----------------------------------------
tl <-
  transf_labels(comp_labels,
                transformation_type = "ilr")

cm <-
  get_cm_from_model(model = main_model,
                    comp_labels = comp_labels,
                    transf_labels = tl)$cm
### Set up a data frame which will hold key numbers----------------------
numbers_df <- data.frame(matrix(nrow = 0, ncol = 8))
colnames(numbers_df) <-
  c(
    "Model",
    "n",
    "n_event",
    "CM: Sleep",
    "CM: SB",
    "CM: LIPA",
    "CM: MVPA",
    "CM: MVPA (min/day)"
  )

### Iterate over model pairs generating the relevant plot and adding detail to numbers_df -------------------
model_pair_list <- list(
  list("main", "minimal"),
  list("main", "only_fu"),
  list("main", "only_fu_4yr"),
  list("main", "add_adj_bmi")
)

for (pair in model_pair_list) {
  # Get models
  model1 <- get(paste0(pair[[1]], "_model"))
  model2 <- get(paste0(pair[[2]], "_model"))
  mlist <- list(model1, model2)

  # Write main components of df
  for (i in c(1, 2)) {
    model <- mlist[[i]]
    name <- pair[[i]]

    cm_hrs <-
      get_cm_from_model(model, comp_labels = comp_labels, transf_labels = tl)$cm *
      24
    cm_mins <-
      get_cm_from_model(model, comp_labels = comp_labels, transf_labels = tl)$cm *
      24 * 60
    numbers_df <-
      rbind(
        numbers_df,
        data.frame(
          "Model" = name,
          "n" = model$n,
          "n_event" = model$nevent,
          "CM: Sleep" = cm_hrs$sleep,
          "CM: SB" = cm_hrs$SB,
          "CM: LIPA" = cm_hrs$LIPA,
          "CM: MVPA" = cm_hrs$MVPA,
          "CM: MVPA (min/day)" = cm_mins$MVPA
        )
      )
  }

  # Main matrix plot ============================================
# This makes and writes to file the plot
svg(
  file=here::here("outputs", paste0("comp_figure_", pair[[1]], "_" , pair[[2]], ".svg")), width = 22, height = 12
)

# Save ggplot object as SVG in the "/outputs/" folder - revisit when ready to print figures
output_folder <- here::here("outputs")
svg_file <- file.path(output_folder, (paste0("comp_figure_", pair[[1]], "_" , pair[[2]], ".svg")))
plot <- compare_all_transfers_side_by_side(
    comp_model = model1,
    comp_model2 = model2,
    comp_labels = comp_labels,
    yllimit = 0.8,
    yulimit = 1.3,
    plot_log =  TRUE,
    units = "hr/day",
    granularity = 1000,
    point_specification = ggplot2::geom_point(size = 2)
  )
ggsave(filename = svg_file, plot = plot, device = "svg", width = 22, height = 12)
}

### Write numbers_df --------------------------------------------------
 write.csv(numbers_df, file=here::here("outputs", paste0("details_for_plots.csv")), row.names=FALSE)
```


Main
```{r}
## Specific predictions Moving 1:proportional Followed code outlined in [Rosemary's compositional paper repo](https://github.com/OxWearables/manuscript_ml_behaviours_cvd_2021/blob/main/epiAnalysis/scripts/analysis.R#L44)
##  Forest plot to show 1 hour change in behavior  ---------------------------------------------------------------
# Changes in activity by one extra hour
# Forest plot
cm_df <- as.data.frame(comp_mean(
  data = dat,
  comp_labels = comp_labels, 
  det_limit = min(dat$MVPA[dat$MVPA >0]),
    units = "hr/day"
))
new_comp1 <- change_composition(
  composition = cm_df,
  main_part = "MVPA",
  main_change = +1.0,
  comp_labels = comp_labels)
new_comp2 <- change_composition(
  composition = cm_df,
  main_part = "sleep",
  main_change = +1.0,
  comp_labels = comp_labels)
new_comp3 <- change_composition(
  composition = cm_df,
  main_part = "SB",
  main_change = +1.0,
  comp_labels = comp_labels)
new_comp4 <- change_composition(
  composition = cm_df,
  main_part = "LIPA",
  main_change = +1.0,
  comp_labels = comp_labels)
new_comp1_halfhour <- change_composition(
  composition = cm_df,
  main_part = "MVPA",
  main_change = +0.25,
  comp_labels = comp_labels)
######################## Plot 1 hour change 
list_for_plot <- list ("1 hour/day more sleep" = new_comp2, 
                       "1 hour/day more SB" = new_comp3, 
                       "1 hour/day more LIPA" = new_comp4,
                       "1 hour/day more MVPA" = new_comp1
                      )
#### Main plot (run all at once to get to print)
forestplot_main <- forest_plot_comp(
  composition_list = list_for_plot,
  model = main_model,
  comp_labels = comp_labels,
  pred_name = "Hazard Ratio (95% CI)",
  text_settings = forestplot::fpTxtGp(
    label = grid::gpar(
      fontfamily = "sans",
      cex = 0.90,
      fontface = 2
    ),
    xlab = grid::gpar(
      fontfamily = "sans",
      cex = 0.90,
      fontface = 2
    ),
    ticks = grid::gpar(cex = 0.80, fontface = 2)
  ),
  boxsize = 0.2,
  x_label = "Estimated Hazard Ratio"
)
forestplot_main
#manually save forest plot
######################## Plot 1 hour change for all but MVPA
list_for_plot_halfhour <- list ("1 hour/day more sleep" = new_comp2, 
                       "1 hour/day more SB" = new_comp3, 
                       "1 hour/day more LIPA" = new_comp4,
                       "15 mins/day more MVPA" = new_comp1_halfhour
                      )
#### Main plot
forestplot_main_halfhour <-forest_plot_comp(composition_list=list_for_plot_halfhour,
                 model = main_model, 
                 comp_labels = comp_labels,
                 pred_name = "Hazard Ratio (95% CI)",
  text_settings = forestplot::fpTxtGp(
    label = grid::gpar(
      fontfamily = "sans",
      cex = 0.90,
      fontface = 2
    ),
    xlab = grid::gpar(
      fontfamily = "sans",
      cex = 0.90,
      fontface = 2
    ),
    ticks = grid::gpar(cex = 0.75, fontface = 2)
  ),
  boxsize = 0.2,
  x_label = "Estimated Hazard Ratio")

svg("forestplot_main_halfhour.svg", width = 10, height = 5)
print(forestplot_main_halfhour)
dev.off()
```



## Specific predictions - Grouping MVPA and LIPA together for single activity category
```{r}
## Main model ----------------------------
main_model_threecat <- comp_model(
  type = "cox",
  covariates = covariates,
  outcome = Surv(
    time = dat$agein,
    time2 = dat$ageout,
    event = dat$ProbDefPsA_incident),
  data = dat,
  comp_labels = comp_labels_three,
  rounded_zeroes = TRUE,
  det_limit = min(dat$LIPA_MVPA[dat$LIPA_MVPA >0]))


## Sensitivity analyses with minimal adjustment----------------------------
### Run model
minimal_model_threecat <-
  comp_model(
    type = "cox",
    outcome = Surv(
    time = dat$agein,
    time2 = dat$ageout,
    event = dat$ProbDefPsA_incident),
    data = dat,
    comp_labels = comp_labels_three,
    rounded_zeroes = TRUE,
    det_limit = min(dat$LIPA_MVPA[dat$LIPA_MVPA >0])
  )

## Sensitivity analyses for BMI adjustment----------------------------
### Run model
add_adj_bmi_model_threecat <- comp_model(
  type = "cox",
  covariates = c(covariates, "BMI_cats") ,
  outcome = Surv(
    time = dat$agein,
    time2 = dat$ageout,
    event = dat$ProbDefPsA_incident),
  data = dat,
  comp_labels = comp_labels_three,
  rounded_zeroes = TRUE,
  det_limit = min(dat$LIPA_MVPA[dat$LIPA_MVPA >0]))

```

## Specific_predictions 1:1 change
```{r}
## Specific_predictions ---------------------------------------------------------------
tl <-
  transf_labels(comp_labels,
                transformation_type = "ilr")
cm <-
  get_cm_from_model(model = main_model,
                    comp_labels = comp_labels,
                    transf_labels = tl)$cm
new <-
  rbind(
    cm,
    change_composition(
      cm,
      main_part = "SB",
      at_expense_of = "LIPA",
      main_change = 1 / 24,
      comp_labels = comp_labels
    ),
    change_composition(
      cm,
      main_part = "LIPA",
      at_expense_of = "SB",
      main_change = 1 / 24,
      comp_labels = comp_labels
    ),
    change_composition(
      cm,
      main_part = "MVPA",
      at_expense_of = "LIPA",
      main_change = 1 / 24,
      comp_labels = comp_labels
    ),
    change_composition(
      cm,
      main_part = "MVPA",
      at_expense_of = "SB",
      main_change = 1 / 24,
      comp_labels = comp_labels
    ),
      change_composition(
      cm,
      main_part = "MVPA",
      at_expense_of = "sleep",
      main_change = 1 / 24,
      comp_labels = comp_labels
    ),
    change_composition(
      cm,
      main_part = "LIPA",
      at_expense_of = "MVPA",
      main_change = 0.25 / 24,
      comp_labels = comp_labels
    ),
    change_composition(
      cm,
      main_part = "MVPA",
      at_expense_of = "LIPA",
      main_change = 0.25 / 24,
      comp_labels = comp_labels
    ), 
    change_composition(
      cm,
      main_part = "SB",
      at_expense_of = "MVPA",
      main_change = 0.25 / 24,
      comp_labels = comp_labels
    ),
    change_composition(
      cm,
      main_part = "MVPA",
      at_expense_of = "SB",
      main_change = 0.25 / 24,
      comp_labels = comp_labels
    ),
    change_composition(
      cm,
      main_part = "sleep",
      at_expense_of = "MVPA",
      main_change = 0.25 / 24,
      comp_labels = comp_labels
    ),
    change_composition(
      cm,
      main_part = "MVPA",
      at_expense_of = "sleep",
      main_change = 0.25 / 24,
      comp_labels = comp_labels
    )
  )
preds <-  predict_fit_and_ci(
  model =  main_model,
  new_data = new,
  comp_labels = comp_labels,
  units = "hr/day"
)
preds <- preds[, c("fit", "lower_CI", "upper_CI")]
rownames(preds) <-
  c(
    "Compositional mean",
    "1 hr/day extra SB from LIPA",
    "1 hr/day extra LIPA from SB",
    "1 hr/day extra MVPA from LIPA",
    "1 hr/day extra MVPA from SB",
    "1 hr/day extra MVPA from sleep",
    "15 min/day extra LIPA from MVPA",
    "15 min/day extra MVPA from LIPA",
    "15 min/day extra SB from MVPA",
    "15 min/day extra MVPA from SB",
    "15 min/day extra sleep from MVPA",
    "15 min/day extra MVPA from sleep"
  )
p <- format(round(preds, digits = 2), nsmall = 2)

# make row labels as separate variable column
p$changes_in_composition <- rownames(p)
# Reorder columns to make 'changes_in_composition' the first column
p <- p[, c("changes_in_composition", names(p)[-ncol(p)])]
p$lower_CI <- as.numeric(p$lower_CI)
p$upper_CI <- as.numeric(p$upper_CI)
p$fit <- as.numeric(p$fit)
#combine estimate and confidence intervals, keeping 2 digits after decimal including 0s
p$hr_cis_multivaradj <- sprintf("%.2f (%.2f-%.2f)", p$fit, p$lower_CI, p$upper_CI)


########################
#### Main
preds <-  predict_fit_and_ci(
  model =  main_model,
  new_data = new,
  comp_labels = comp_labels,
  units = "hr/day"
)
preds <- preds[, c("fit", "lower_CI", "upper_CI")]
rownames(preds) <-
  c(
    "Compositional mean",
    "1 hr/day extra SB from LIPA",
    "1 hr/day extra LIPA from SB",
    "1 hr/day extra MVPA from LIPA",
    "1 hr/day extra MVPA from SB",
    "1 hr/day extra MVPA from sleep",
    "15 min/day extra LIPA from MVPA",
    "15 min/day extra MVPA from LIPA",
    "15 min/day extra SB from MVPA",
    "15 min/day extra MVPA from SB",
    "15 min/day extra sleep from MVPA",
    "15 min/day extra MVPA from sleep"
  )
p <- format(round(preds, digits = 2), nsmall = 2)

# make row labels as separate variable column
p$changes_in_composition <- rownames(p)
# Reorder columns to make 'changes_in_composition' the first column
p <- p[, c("changes_in_composition", names(p)[-ncol(p)])]
p$lower_CI <- as.numeric(p$lower_CI)
p$upper_CI <- as.numeric(p$upper_CI)
p$fit <- as.numeric(p$fit)
#combine estimate and confidence intervals, keeping 2 digits after decimal including 0s
p$hr_cis_multivaradj <- sprintf("%.2f (%.2f-%.2f)", p$fit, p$lower_CI, p$upper_CI)
# subset variables
p <- p %>% dplyr::select(changes_in_composition, hr_cis_multivaradj) 

########################
#### Minimal
preds_minimal <-  predict_fit_and_ci(
  model =  minimal_model,
  new_data = new,
  comp_labels = comp_labels,
  units = "hr/day"
)
preds_minimal <- preds_minimal[, c("fit", "lower_CI", "upper_CI")]
rownames(preds_minimal) <-
  c(
    "Compositional mean",
    "1 hr/day extra SB from LIPA",
    "1 hr/day extra LIPA from SB",
    "1 hr/day extra MVPA from LIPA",
    "1 hr/day extra MVPA from SB",
    "1 hr/day extra MVPA from sleep",
    "15 min/day extra LIPA from MVPA",
    "15 min/day extra MVPA from LIPA",
    "15 min/day extra SB from MVPA",
    "15 min/day extra MVPA from SB",
    "15 min/day extra sleep from MVPA",
    "15 min/day extra MVPA from sleep"
  )
p_minimal <- format(round(preds_minimal, digits = 2), nsmall = 2)

# make row labels as separate variable column
p_minimal$changes_in_composition <- rownames(p_minimal)
# Reorder columns to make 'changes_in_composition' the first column
p_minimal <- p_minimal[, c("changes_in_composition", names(p_minimal)[-ncol(p_minimal)])]
p_minimal$lower_CI <- as.numeric(p_minimal$lower_CI)
p_minimal$upper_CI <- as.numeric(p_minimal$upper_CI)
p_minimal$fit <- as.numeric(p_minimal$fit)
#combine estimate and confidence intervals, keeping 2 digits after decimal including 0s
p_minimal$hr_cis_minimal <- sprintf("%.2f (%.2f-%.2f)", p_minimal$fit, p_minimal$lower_CI, p_minimal$upper_CI)
# subset variables
p_minimal <- p_minimal %>% dplyr::select(changes_in_composition, hr_cis_minimal) 

########################
#### main + bmi
preds_multi_bmi <-  predict_fit_and_ci(
  model =  add_adj_bmi_model,
  new_data = new,
  comp_labels = comp_labels,
  units = "hr/day"
)
preds_multi_bmi <- preds_multi_bmi[, c("fit", "lower_CI", "upper_CI")]
rownames(preds_multi_bmi) <-
  c(
    "Compositional mean",
    "1 hr/day extra SB from LIPA",
    "1 hr/day extra LIPA from SB",
    "1 hr/day extra MVPA from LIPA",
    "1 hr/day extra MVPA from SB",
    "1 hr/day extra MVPA from sleep",
    "15 min/day extra LIPA from MVPA",
    "15 min/day extra MVPA from LIPA",
    "15 min/day extra SB from MVPA",
    "15 min/day extra MVPA from SB",
    "15 min/day extra sleep from MVPA",
    "15 min/day extra MVPA from sleep"
  )
p_multi_bmi <- format(round(preds_multi_bmi, digits = 2), nsmall = 2)

# make row labels as separate variable column
p_multi_bmi$changes_in_composition <- rownames(p_multi_bmi)
# Reorder columns to make 'changes_in_composition' the first column
p_multi_bmi <- p_multi_bmi[, c("changes_in_composition", names(p_multi_bmi)[-ncol(p_multi_bmi)])]
p_multi_bmi$lower_CI <- as.numeric(p_multi_bmi$lower_CI)
p_multi_bmi$upper_CI <- as.numeric(p_multi_bmi$upper_CI)
p_multi_bmi$fit <- as.numeric(p_multi_bmi$fit)
#combine estimate and confidence intervals, keeping 2 digits after decimal including 0s
p_multi_bmi$hr_cis_multi_bmi <- sprintf("%.2f (%.2f-%.2f)", p_multi_bmi$fit, p_multi_bmi$lower_CI, p_multi_bmi$upper_CI)
# subset variables
p_multi_bmi <- p_multi_bmi %>% dplyr::select(changes_in_composition, hr_cis_multi_bmi)  

# Write dataset as csv
# Assuming 'p', 'p_minimal', and 'p_multi' are your datasets 
combined_predictions <- p_minimal %>%
  left_join(p, by = "changes_in_composition") %>%
  left_join(p_multi_bmi, by = "changes_in_composition")


# Get information for plot ==============================
# This is the value of all the movement behaviours at the compositional mean behaviours
# i.e. the point marked with the vertical line on the graph.
get_cm_from_model(main_model_threecat, comp_labels_three, transf_labels(comp_labels_three, "ilr"))$cm*24
get_cm_from_model(main_model_threecat, comp_labels_three, transf_labels(comp_labels_three, "ilr"))$cm*24*60

# Main matrix plot ============================================
# This makes and writes to file the plot
svg(
  file=here::here("outputs", paste0("comp_figure_three.svg")), width = 15, height = 10
)

# Save ggplot object as SVG in the "/outputs/" folder - revisit when ready to print figures
output_folder <- here::here("outputs")
svg_file <- file.path(output_folder, (paste0("comp_figure_three.svg")))
plot <- plot_transfers(from_part = "SB",
               to_part = "LIPA_MVPA",
               model = main_model_threecat ,
               comp_labels = comp_labels_three,
               y_label = "Model-estimated difference in outcome",
               units = "hr/day",
                point_specification = ggplot2::geom_point(size = 0.5),
  theme = theme_classic() + theme(axis.line = element_blank()))

ggsave(filename = svg_file, plot = plot, device = "svg", width = 15, height = 10)
plot

## plot to get other comp mean values
plot_transfers(from_part = "SB",
               to_part = "sleep",
               model = main_model_threecat ,
               comp_labels = comp_labels_three,
               y_label = "Model-estimated difference in outcome",
               units = "hr/day",
                point_specification = ggplot2::geom_point(size = 0.5),
  theme = theme_classic() + theme(axis.line = element_blank()))

plot_transfers(from_part = "sleep",
               to_part = "LIPA_MVPA",
               model = main_model_threecat ,
               comp_labels = comp_labels_three,
               y_label = "Model-estimated difference in outcome",
               units = "hr/day",
                point_specification = ggplot2::geom_point(size = 0.5),
  theme = theme_classic() + theme(axis.line = element_blank()))


# Get information for plot ==============================
# This is the value of all the movement behaviours at the compositional mean behaviours
# i.e. the point marked with the vertical line on the graph.
get_cm_from_model(comp_labels, transf_labels(comp_labels, "ilr"))$cm*24
get_cm_from_model(comp_labels, transf_labels(comp_labels, "ilr"))$cm*24*60

```
