---
title: "9_modeling_compositional"
author: "Alaina Shreves"
date: "`r format(Sys.time(), '%a %b %d, %Y %X')`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 5
    theme: lumen
---
```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This script runs the compositional analysis based on Rosemary's scirpt for Jemma's project that can be found [here](https://github.com/OxWearables/AF_RW/blob/main/scripts_pipeline/compositional_script.R).

# Set up workspace

```{r}
#Load standard packages
pkgs <-
  c(
    "data.table",
    "rmarkdown",
    "plyr",
    "lubridate",
    "magrittr",
    "ggplot2",
    "data.table",
    "dplyr",
    "stringr",
    "splitstackshape",
    "scales",
    "haven",
    "Epi",
    #"rtables",
    "table1",
   # "tableone",
    "boot",
    "knitr",
    "pspline",
    "survival",
    "xtable",
    "here",
    "ggtern",
    "devtools"
  ) # packages we need
#Note - At install, load epicoda using Rosemary's instructions from GitHub - https://github.com/OxWearables/epicoda

#install
pkgs_inst <-
  pkgs[!{
    pkgs %in% rownames(installed.packages())
  }] # check which are not present
install.packages(pkgs_inst, repos = "https://www.stats.bris.ac.uk/R/") # install

lapply(pkgs, library, character.only = TRUE) #load all packages

#renv::install(pkgs)


## Source helper functions---------------------------------------------------
source(here("compare_plot_linear.R"))
source("rounding_functions.R")
source("cut_by_quantile.R")
source("cox_functions.R")
source(here("arrange_plots_matrix.R"))
source(here("med_and_iqr.R"))
source(here("forest_plot_examples_with_evals.R"))
source(here("compare_plot2.R"))
source(here("compare_plot3.R"))
source(here("plot_transfers_ism.R"))

```

```{r}
library(devtools)
devtools::install_github("activityMonitoring/epicoda",  build_opts = c("--no-resave-data"), build_vignettes = FALSE, build_manual = FALSE)

library(epicoda)
```
## Load data
```{r}
#############read in data
#updated file path 2023 10 16
dat <- fread(here("PsAfinal/prepped_steps_PsAfinal.csv"), data.table = FALSE)

dat$V1 <- NULL

dat <- dat %>%
  filter(total_prevalent_PsO == TRUE)
```

```{r}
dir.create("outputs")
```

```{r}
dat$tdi_quarters <- qtile_cut(dat$tdi_raw, probs = seq(0, 1, by = 0.25), dp_label = 1)
dat$BMI_cats <-
  cut(dat$BMI,
      breaks = c(0, 25, 30, 10000),
      labels = c("<24.9", "25.0-29.9", "30.0+"),
      right = FALSE)

dat$sleep <- NULL
dat$SB <- NULL
dat$LIPA <- NULL
dat$MVPA <- NULL

dat$MVPA <- dat$MVPA_overall_average
dat$sleep <- dat$sleep_overall_average
dat$SB <- dat$sedentary_overall_average
dat$LIPA <- dat$light_overall_average 

#1 participant with NA for MVPA
dat <- dat[!is.na(dat$MVPA), ]
```

# Prep dataset
```{r}
# Set up for modelling=====================================
### Alaina updated covariates to match other scripts
covariates <-
  c(
"sex",
"tdi_quarters",
"qualif",
"smoking",
"alcohol"
  )

# AHS add code for time points
dat$agein <- dat$age_entry_days
dat$ageout <- dat$age_exit_days

# Define the names of the comp variables in the main group
comp_labels <- c("sleep", "SB", "LIPA", "MVPA")
# Check sum to 1
dat$sum_sleep_SB_LIPA_MVPA <- rowSums(dat[, c("sleep", "SB", "LIPA", "MVPA")])
table(dat$sum_sleep_SB_LIPA_MVPA)

### for  models that combine activity behavior in a single category
# Combine exposure categories and make new labels
dat$LIPA_MVPA <- (dat$LIPA + dat$MVPA)
# Define the names of the compositional variables
comp_labels_three <- c("sleep", "SB", "LIPA_MVPA")

# Define the names of the compositional variables for the cutpoint based analysis
comp_labels_cutpoint <- c("SB_cutpoint", "sleep", "LIPA_cutpoint","MVPA_cutpoint")
# Check sum to 1
#dat$sum_SB_cutpoint_sleep_LIPA_cutpoint_MVPA_cutpoint <- rowSums(dat[, c("SB_cutpoint", "sleep", "LIPA_cutpoint", "MVPA_cutpoint")])
#table(dat$sum_SB_cutpoint_sleep_LIPA_cutpoint_MVPA_cutpoint)

# Define the names of the compositional variables for the cutpoint based analysis with VPA seprate
comp_labels_cutpoint_fivecat <- c("SB_cutpoint", "sleep", "LIPA_cutpoint","MPA_cutpoint","VPA_cutpoint")
```

# Modeling - all
## Matrix - main 
Model compositional matrix for risk of Incident PA-related cancer
```{r}
# Run model ==============================================
######### Incident PA-related cancers 
main_model <- comp_model(
  type = "cox",
  covariates = covariates,
  outcome = Surv(
    time = dat$agein,
    time2 = dat$ageout,
    event = dat$PsATotal_incident),
  data = dat,
  comp_labels = comp_labels,
  rounded_zeroes = TRUE,
  det_limit = min(dat$MVPA[dat$MVPA >0]))


# predictions (note that this takes a while to run)
tab_coefs(scale_type = "exp",
          level = 0.95,
          type = "cox",
  covariates = covariates,
  outcome = Surv(
    time = dat$agein,
    time2 = dat$ageout,
    event = dat$PsATotal_incident),
  data = dat,
  comp_labels = comp_labels,
  rounded_zeroes = TRUE,
  det_limit = min(dat$MVPA[dat$MVPA >0]))



# Check proportional hazards==============================
# Note From Rosemary's paper - Participants with missing data in any adjustment variable were excluded. The proportional hazards assumption was tested component-wise and globally using the Grambsch-Therneau test with the Kaplan-Meier transformation, and there was no evidence (at the 5% level) that it was violated in the main analysis. Plots of the Schoenfeld residuals were also examined.
cox.zph(main_model) # Grambsch-Therneau test
plot(cox.zph(main_model)) # Plots of scaled Schoenfeld residuals.

# Get information for plot ==============================
summary(main_model)
# This is the value of all the movement behaviours at the compositional mean behaviours
# i.e. the point marked with the vertical line on the graph.
get_cm_from_model(main_model, comp_labels, transf_labels(comp_labels, "ilr"))$cm*24
get_cm_from_model(main_model, comp_labels, transf_labels(comp_labels, "ilr"))$cm*24*60

# Main matrix plot ============================================
# This makes and writes to file the plot
svg(
  file=here::here("outputs", paste0("comp_figure.svg")), width = 15, height = 10
)

# Save ggplot object as SVG in the "/outputs/" folder - revisit when ready to print figures
output_folder <- here::here("outputs")
svg_file <- file.path(output_folder, (paste0("comp_figure.svg")))
plot <- plot_maxi_matrix_transfers(
  comp_model = main_model,
  comp_labels = comp_labels,
  yllimit = 0.8,
  yulimit = 1.3,
  plot_log = TRUE,
  granularity = 4000,
  point_specification = ggplot2::geom_point(size = 0.5),
  theme = theme_classic() + theme(axis.line = element_blank())
)
ggsave(filename = svg_file, plot = plot, device = "svg", width = 15, height = 10)

# Save as a jpg for poster if needed
# output_folder <- here::here("outputs")
# svg_file <- file.path(output_folder, paste0("comp_figure_tall.svg"))
# ggsave(filename = svg_file, plot = plot, device = "svg", width = 8, height = 8)
```
# Specific predictions
## Specific predictions Moving 1:proportional
Followed code outlined in [Rosemary's compositional paper repo](https://github.com/OxWearables/manuscript_ml_behaviours_cvd_2021/blob/main/epiAnalysis/scripts/analysis.R#L44)

Note- since the mean of MVPA is less than 1 hour, have to make the unit of change for MVPA less than 1 hour. ///
I used Rosemary's suggestion of 15 min/day change in MVPA and kept other behaviors at 1 hour/day change. 
```{r}
## Specific predictions Moving 1:proportional Followed code outlined in [Rosemary's compositional paper repo](https://github.com/OxWearables/manuscript_ml_behaviours_cvd_2021/blob/main/epiAnalysis/scripts/analysis.R#L44)

##  Forest plot to show 1 hour change in behavior  ---------------------------------------------------------------
# Changes in activity by one extra hour
# Forest plot
cm_df <- as.data.frame(comp_mean(
  data = dat,
  comp_labels = comp_labels, 
  det_limit = min(dat$MVPA[dat$MVPA >0]),
    units = "hr/day"
))

new_comp1 <- change_composition(
  composition = cm_df,
  main_part = "MVPA",
  main_change = +1.0,
  comp_labels = comp_labels)
new_comp2 <- change_composition(
  composition = cm_df,
  main_part = "sleep",
  main_change = +1.0,
  comp_labels = comp_labels)
new_comp3 <- change_composition(
  composition = cm_df,
  main_part = "SB",
  main_change = +1.0,
  comp_labels = comp_labels)
new_comp4 <- change_composition(
  composition = cm_df,
  main_part = "LIPA",
  main_change = +1.0,
  comp_labels = comp_labels)

new_comp1_halfhour <- change_composition(
  composition = cm_df,
  main_part = "MVPA",
  main_change = +0.5,
  comp_labels = comp_labels)

######################## Plot 1 hour change 
list_for_plot <- list ("1 hour/day more sleep" = new_comp2, 
                       "1 hour/day more SB" = new_comp3, 
                       "1 hour/day more LIPA" = new_comp4,
                       "1 hour/day more MVPA" = new_comp1
                      )

#### Main plot (run all at once to get to print)
forestplot_main <- forest_plot_comp(
  composition_list = list_for_plot,
  model = main_model,
  comp_labels = comp_labels,
  pred_name = "Hazard Ratio (95% CI)",
  text_settings = forestplot::fpTxtGp(
    label = grid::gpar(
      fontfamily = "sans",
      cex = 0.90,
      fontface = 2
    ),
    xlab = grid::gpar(
      fontfamily = "sans",
      cex = 0.90,
      fontface = 2
    ),
    ticks = grid::gpar(cex = 0.80, fontface = 2)
  ),
  boxsize = 0.2,
  x_label = "Estimated Hazard Ratio"
)
forestplot_main
#manually save forest plot

######################## Plot 1 hour change for all but MVPA
list_for_plot_halfhour <- list ("1 hour/day more sleep" = new_comp2, 
                       "1 hour/day more SB" = new_comp3, 
                       "1 hour/day more LIPA" = new_comp4,
                       "0.5 hour/day more MVPA" = new_comp1_halfhour
                      )
#### Main plot
forestplot_main_halfhour <-forest_plot_comp(composition_list=list_for_plot_halfhour,
                 model = main_model, 
                 comp_labels = comp_labels,
                 pred_name = "Hazard Ratio (95% CI)",
  text_settings = forestplot::fpTxtGp(
    label = grid::gpar(
      fontfamily = "sans",
      cex = 0.90,
      fontface = 2
    ),
    xlab = grid::gpar(
      fontfamily = "sans",
      cex = 0.90,
      fontface = 2
    ),
    ticks = grid::gpar(cex = 0.75, fontface = 2)
  ),
  boxsize = 0.2,
  x_label = "Estimated Hazard Ratio")
forestplot_main_halfhour
```

## Specific predictions - Grouping MVPA and LIPA together for single activity category
```{r}
## Main model ----------------------------
main_model_threecat <- comp_model(
  type = "cox",
  covariates = covariates,
  outcome = Surv(
    time = dat$agein,
    time2 = dat$ageout,
    event = dat$PsATotal_incident),
  data = dat,
  comp_labels = comp_labels_three,
  rounded_zeroes = TRUE,
  det_limit = min(dat$LIPA_MVPA[dat$LIPA_MVPA >0]))


## Sensitivity analyses with minimal adjustment----------------------------
### Run model
minimal_model_threecat <-
  comp_model(
    type = "cox",
    outcome = Surv(
    time = dat$agein,
    time2 = dat$ageout,
    event = dat$PsATotal_incident),
    data = dat,
    comp_labels = comp_labels_three,
    rounded_zeroes = TRUE,
    det_limit = min(dat$LIPA_MVPA[dat$LIPA_MVPA >0])
  )

## Sensitivity analyses for BMI adjustment----------------------------
### Run model
add_adj_bmi_model_threecat <- comp_model(
  type = "cox",
  covariates = c(covariates, "BMI_cats") ,
  outcome = Surv(
    time = dat$agein,
    time2 = dat$ageout,
    event = dat$PsATotal_incident),
  data = dat,
  comp_labels = comp_labels_three,
  rounded_zeroes = TRUE,
  det_limit = min(dat$LIPA_MVPA[dat$LIPA_MVPA >0]))

```
