---
title: "RiskAnalysis"
format: html
editor: visual
---

## Continue environment setup

## Load packages

Load packages for use in subsequent scripts:

```{r}
library(data.table)
library(plyr)
library(dplyr)
library(magrittr)
library(lubridate)
library(readr)
library(survival)
library(Epi)
library(emmeans)
library(ggplot2)
##library(ckbplotr)
```

```{r}
dat <- fread("prepped_steps_before_exclude.csv", data.table = FALSE)
```

```{r}
tab_exc <- data.frame("Exclusion" = "Starting cohort", "Number_excluded" = NA, "Number_remaining" = nrow(dat))
```

We do the accelerometer data quality exclusions:

-   Exclude participants without step data:

```{r}
nb <- nrow(dat)
dat <- dat[!is.na(dat$med_steps), ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "No step data",
      "Number_excluded" = nb - nrow(dat),
      "Number_remaining" = nrow(dat)
    )
  )
```

-   Exclude participants whose device could not be calibrated:

```{r}
nb <- nrow(dat)
dat <- dat[dat$quality_good_calibration == "Yes", ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Poor calibration",
      "Number_excluded" = nb - nrow(dat),
      "Number_remaining" = nrow(dat)
    )
  )
```

-   Exclude participants for whom \>1% of values were clipped (fell outside the sensor's range) before or after calibration:

```{r}
nb <- nrow(dat)
dat <- dat[(dat$clips_before_cal < 0.01*dat$total_reads) & (dat$clips_after_cal < 0.01*dat$total_reads) , ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Too many clips",
      "Number_excluded" = nb - nrow(dat),
      "Number_remaining" = nrow(dat)
    )
  )
```

-   Exclude participants who had \<3 days wear or did not have wear in each hour of the 24 hour day:

```{r}
nb <- nrow(dat)
dat <- dat[dat$quality_good_wear_time == "Yes", ] # Note that this has already been calculated in UKB, 
# we don't need to manually calculate it: https://biobank.ndph.ox.ac.uk/showcase/field.cgi?id=90015
# But we might actually use the values from the new data processing

# 2023_01_12 - Now using quality.goodWearTime, which is calcualted from the new data processing
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Poor wear time",
      "Number_excluded" = nb - nrow(dat),
      "Number_remaining" = nrow(dat)
    )
  )
```

-   Exclude participants with unrealistically high overall activity values:

```{r}
nb <- nrow(dat)
dat <- dat[dat$overall_activity < 100, ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Very high overall activity",
      "Number_excluded" = nb - nrow(dat),
      "Number_remaining" = nrow(dat)
    )
  )
```

-   Exclude people lost to follow up before accelerometer wear:

```{r}
nb <- nrow(dat)
dat <- dat[!(dat$date_cens < dat$date_end_accel), ]
tab_exc <- rbind(
  tab_exc,
  data.frame(
    "Exclusion" = "Lost to linked health record follow-up before accelerometer study entry",
    "Number_excluded" = nb - nrow(dat),
    "Number_remaining" = nrow(dat)
  )
)
```


Missing data in adjustment variables:
```{r}
for (
  cov in c(
    "age_entry_years",
    "sex",
    "BMI",
    "ethnicity",
    "tdi_raw",
    "qualif",
    "smoking",
    "alcohol",
    "fresh_fruit",
    "processed_meat",
    "oily_fish",
    "added_salt",
    "PRS_RA"
  )
){
  nb <- nrow(dat)
  print(cov)
  missing_cov <- is.na(dat[, cov])|(as.character(dat[, cov]) == "") |(as.character(dat[, cov]) == "Missing") # for safety coerce to character for second check as can return NA on some classes e.g. Date
  dat <- dat[!missing_cov,]
  tab_exc <- rbind(
    tab_exc,
    data.frame(
      "Exclusion" = paste0("Missing ", cov),
      "Number_excluded" = nb - nrow(dat),
      "Number_remaining" = nrow(dat)
    )
  )
}
```
