---
title: "RiskAnalysis"
format: html
editor: visual
---

## Continue environment setup

## Load packages

Load packages for use in subsequent scripts:

```{r}
library(data.table)
library(plyr)
library(dplyr)
library(magrittr)
library(lubridate)
library(readr)
library(survival)
library(Epi)
library(emmeans)
library(ggplot2)
#library(ckbplotr)
library(rms)
library(timeROC)
library(riskRegression)
library(mfp)
library(pec)
##library(ckbplotr)
```

```{r}
source("rounding_functions.R")
source("cut_by_quantile.R")
source("format_for_table.R")
```

```{r}
dat <- fread("prepped_steps_before_exclude.csv", data.table = FALSE)
```

Sandbox - create backup DatRisk - first will be known lifestyle risk factors to get maximum events 
then a DatRisk2 - which will be as above but also contain PRS and Rheumatoid Factor as continuous

DatRisk - including PRS and RF
```{r}
DatRisk <- dat %>%
          select(eid,age_entry_years, 
                 RA_incident, RA_incident_date, fu_time,date_cens,date_end_accel,
                 #known risk factors
                 smoking, BMI, sex, 
                 #exposures in risk model
            med_steps, CadencePeak1Adjusted.steps.min., 
                 CadencePeak30Adjusted.steps.min.,
            MVPA_overall_average,sedentary_overall_average,sleep_overall_average,
            overall_activity,
            #RF and PRS
            RF_cont, PRS_RA,
            #Prevalent RA so that we can remove them
            total_prevalent_RA,
            #quality assurance in med_steps
            quality_good_wear_time,clips_before_cal,clips_after_cal,
            total_reads,quality_good_calibration )
```

```{r}
tab_exc <- data.frame("Exclusion" = "Starting cohort", "Number_excluded" = NA, "Number_remaining" = nrow(dat))
```

We do the accelerometer data quality exclusions:

-   Exclude participants without step data:

```{r}
nb <- nrow(DatRisk)
DatRisk <- DatRisk[!is.na(DatRisk$med_steps), ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "No step data",
      "Number_excluded" = nb - nrow(DatRisk),
      "Number_remaining" = nrow(DatRisk)
    )
  )
```

-   Exclude participants whose device could not be calibrated:

```{r}
nb <- nrow(DatRisk)
DatRisk <- DatRisk[DatRisk$quality_good_calibration == "Yes", ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Poor calibration",
      "Number_excluded" = nb - nrow(DatRisk),
      "Number_remaining" = nrow(DatRisk)
    )
  )
```

-   Exclude participants for whom \>1% of values were clipped (fell outside the sensor's range) before or after calibration:

```{r}
nb <- nrow(DatRisk)
DatRisk <- DatRisk[(DatRisk$clips_before_cal < 0.01*DatRisk$total_reads) & (DatRisk$clips_after_cal < 0.01*DatRisk$total_reads) , ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Too many clips",
      "Number_excluded" = nb - nrow(DatRisk),
      "Number_remaining" = nrow(DatRisk)
    )
  )
```

-   Exclude participants who had \<3 days wear or did not have wear in each hour of the 24 hour day:

```{r}
nb <- nrow(DatRisk)
DatRisk <- DatRisk[DatRisk$quality_good_wear_time == "Yes", ] # Note that this has already been calculated in UKB, 
# we don't need to manually calculate it: https://biobank.ndph.ox.ac.uk/showcase/field.cgi?id=90015
# But we might actually use the values from the new data processing

# 2023_01_12 - Now using quality.goodWearTime, which is calcualted from the new data processing
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Poor wear time",
      "Number_excluded" = nb - nrow(DatRisk),
      "Number_remaining" = nrow(DatRisk)
    )
  )
```

-   Exclude participants with unrealistically high overall activity values:

```{r}
nb <- nrow(DatRisk)
DatRisk <- DatRisk[DatRisk$overall_activity < 100, ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Very high overall activity",
      "Number_excluded" = nb - nrow(DatRisk),
      "Number_remaining" = nrow(DatRisk)
    )
  )
```

-   Exclude people lost to follow up before accelerometer wear:

```{r}
nb <- nrow(DatRisk)
DatRisk <- DatRisk[!(DatRisk$date_cens < DatRisk$date_end_accel), ]
tab_exc <- rbind(
  tab_exc,
  data.frame(
    "Exclusion" = "Lost to linked health record follow-up before accelerometer study entry",
    "Number_excluded" = nb - nrow(DatRisk),
    "Number_remaining" = nrow(DatRisk)
  )
)
```


Missing data in adjustment variables:
```{r}
for (
  cov in c(
    "age_entry_years",
    "sex",
    "BMI",
    "smoking"
  )
){
  nb <- nrow(DatRisk)
  print(cov)
  missing_cov <- is.na(DatRisk[, cov])|(as.character(DatRisk[, cov]) == "") |(as.character(DatRisk[, cov]) == "Missing") # for safety coerce to character for second check as can return NA on some classes e.g. Date
  DatRisk <- DatRisk[!missing_cov,]
  tab_exc <- rbind(
    tab_exc,
    data.frame(
      "Exclusion" = paste0("Missing ", cov),
      "Number_excluded" = nb - nrow(DatRisk),
      "Number_remaining" = nrow(DatRisk)
    )
  )
}
```

Remove prevalent RA 

```{r}
nb <- nrow(DatRisk)

DatRisk <- DatRisk[!(DatRisk$total_prevalent_RA), ]
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Prevalent RA disease at time of accelerometer wear",
      "Number_excluded" = nb - nrow(DatRisk),
      "Number_remaining" = nrow(DatRisk)
    )
  )
```

Create further dataset which is missing PRS 

```{r}
nb <- nrow(DatRisk)



DatRiskPRS <- DatRisk %>%
  filter(!is.na(PRS_RA))
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Missing Polygenic Risk Score",
      "Number_excluded" = nb - nrow(DatRiskPRS),
      "Number_remaining" = nrow(DatRiskPRS)
    )
  )

```

Missing rheumatoid factor data 
```{r}
nb <- nrow(DatRisk)


#rheumatoid factor exclusions
DatRisk_RheumFactor <- DatRiskPRS %>%
  filter(!is.na(RF_cont))
tab_exc <-
  rbind(
    tab_exc,
    data.frame(
      "Exclusion" = "Missing Rheumatoid Factor data",
      "Number_excluded" = nb - nrow(DatRisk_RheumFactor),
      "Number_remaining" = nrow(DatRisk_RheumFactor)
    )
  )

```


Write out csv files 
```{r}
write.csv(DatRisk, "prepped_risk.csv")
write.csv(DatRiskPRS, "prepped_risk_minusPRS.csv")
write.csv(DatRisk_RheumFactor, "prepped_risk_minusRheumFactor.csv")
```

```{r}
print(tab_exc)
```

```{r}
write.csv(tab_exc, "tab_exc.csv")
```
